### 2020-11-13 finalPaper/paper.tex
Initial commit from latex-example

% 
% Sample latex source for a project report
%
% This is intended as an introduction to latex, but includes some
% advice about what you might want for the content of your paper
%
% Jim Teresco
% Williams College, Mount Holyoke College, Siena College, The College
% of Saint Rose
%
% Last modified: Mon Dec  4 19:03:24 EST 2017
%
% First, as you may have guessed, % is the way to define a latex 
% comment
%
% The first line tells latex you want for a default font size and that
% you want to create an ``article''
\documentclass[12pt]{article}
% extra packages to bring in
\usepackage{latexsym}
\usepackage{graphicx}      % extended graphics package
\usepackage{epsfig}        % wrapper for graphicx package
\usepackage{times}
\usepackage{url}
% set some margins, these can be defined as in, cm, pt
\setlength{\topmargin}{-0.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}

% a few macros that might be useful -- any time we type \eg it expands
% to the italicized version defined here
\newcommand{\etal}{{\it et al}.$\:$}
\newcommand{\eg}{{\it e}.{\it g}.$\:$}
\newcommand{\cf}{{\it cf}.$\:$}
\newcommand{\ie}{{\it i}.{\it e}.$\:$}

%% to remove page numbers, uncomment this:
%% \pagestyle{empty}

%% Define single-space command
\newcommand{\singlespace}{
  \protect\renewcommand\baselinestretch{1.0}
  \protect\normalsize
}
% use this instead if you want to disable it completely:
%%\newcommand{\singlespace}{}

%% Define double-space command (really more like 1.5 spacing)
%% This is essential for rough drafts, and not a bad idea even for
%%  final submissions
\newcommand{\doublespace}{
  \protect\renewcommand\baselinestretch{1.5}
  \protect\normalsize
}
% use this instead if you want to disable it completely:
%%\newcommand{\doublespace}{}

% This tells latex we're done defining the preamble stuff and we're
% ready to start writing the document
\begin{document}

% this removes the date that is automatically placed in the title.
% comment it out if you want the date
\date{}

% the next few items define things to go onto the title section, like,
% well, the title, and the list of authors
\title{A Sample Project Report
\footnote{This work was completed in partial fulfillment of the final
project requirement for Computer Science 400 at Siena College, Fall 2017.}}

\author{J.~D.~Teresco\footnote{Corresponding author, e-mail:
jteresco@siena.edu} and O.-W.~Kenobi\\
Department of Computer Science\\
Siena College\\
Loudonville, NY  12211
}

% this tells latex that you're done setting up title stuff and that it
% should go ahead and generate the title here
\maketitle
% leave the page number off but for this page only
\thispagestyle{empty}

% Abstract!
\begin{abstract}
It's usually a good idea to include an abstract.  This should be a
brief summary of your paper.  Highlight your important results.  A
half a page is about right.  Maybe less.  It should be independent of
your subsequent introduction/overview section.
\end{abstract}

% turn on double spacing
\doublespace

% Now we write the text of the paper, hopefully breaking it up into
% nice sections and subsections, using figures and tables as
% appropriate, and referring to those sections using labels instead of
% trying to number things by hand
\section{Overview}
\label{sec:overview}

Begin with an introduction to your topic, specifically emphasizing
what you have done.  Some historical background and references to
related projects are appropriate.  You should also outline the paper
in the overview section.  It might go something like this:

In Section~\ref{sec:vectarray} we discuss how to build a vector from
an array.  Section~\ref{sec:subsections} has a lot of examples of
things you might want \LaTeX~\cite{lamport86} to do.  In
Section~\ref{sec:writing}, we discuss how writing is hard and offer
some suggestions on how to do a better job at it.
Section~\ref{sec:building} describes how to use \LaTeX{} to compile
this document into printable postscript.  Finally,
Section~\ref{sec:conclusions} summarizes the results and presents
suggestions for future work.

Notice how references are used instead of actual section numbers.
This allows you to move, add, or remove sections and have the section
numbers stay correct.  The names are specified below with label commands.

\section{Building a Vector from an Array}
\label{sec:vectarray}

This section is here to give examples of math mode, font changes, the
verbatim environment, and importing a postscript figure.

This experiment measures the time taken to copy an array of $n$ {\tt
int} values to a {\tt Vector} of $n$ {\tt Integer} objects.  We will
use the {\tt add()} method of {\tt class Vector} to append new
values to the end of the vector.  Accessing each item in the array is
a constant time ($O(1)$) operation, as are the construction of the
{\tt Integer} object which will hold the value in the vector and each
{\tt add()} call.  We do this for each of $n$ items, thus, we
expect this operation to take $O(n)$ time.  We create our vector to be
initially of size $n$ to avoid the need to resize dynamically during
the experiments.  We use the {\tt Vector} implementation found in the
{\tt structure} package.

{\it Now, we describe how the experiment is set up, how you run it, and
present and discuss the results.  This sentence is mainly inserted to
show an example of italics.}

A simple Java program was written to copy of an array into the
vector.  A method called {\tt copy()} takes the value of $n$ as its
parameter, creates and fills an array of size $n$, creates an empty
{\tt Vector}, starts a timer, copies the values from the array to the
vector, stops the timer, and prints out the time taken.  The body of
this method is shown in Figure~\ref{fig:code}.

{\singlespace % we don't want to doublespace our code snippet
\begin{figure}[htb]
\begin{verbatim}
    int[] array = new int[n];
    int i;
    for (i=0; i<n; i++) array[i]=i;

    Vector v = new Vector(n);
    long start=System.currentTimeMillis();
    for (i=0; i<n; i++) {
        v.add(new Integer(array[i]));
    }
    long end = System.currentTimeMillis();
    System.out.println("For n="+n+", time="+(end-start));
\end{verbatim}
\caption{Java source to copy an array to a Vector.}
\label{fig:code}
\end{figure}
}

\begin{figure}[htb]
\begin{center}
\includegraphics[angle=-90,width=5in]{times.eps}
\end{center}
\caption{Times in milliseconds to build a Vector from various array
sizes. using the code in Figure~\ref{fig:code}}
\label{fig:times}
\end{figure}

Your figures should have a caption that describes exactly what is in
the figure.  A discussion of the figure should be included in your
main text, referring to the figure by number, like so:

The method in Figure~\ref{fig:code} is executed 10 times for each
value of $n$, and the minimum time is taken.  The runs are all done on
a PC with a 700 MHz Intel Pentium III processor, running FreeBSD 4.3
and the Java Development Kit version 1.3.0.  We choose array and
vector sizes starting at $n=512$ and double them until our final size
$n=262,144$.  These bounds, all powers of 2, were chosen because the
512-element case is the first which takes at least one millisecond,
and the 262,144-element is the largest which can execute without
generating an out of memory exception from the Java run-time
environment.  The graph in Figure~\ref{fig:times} shows the actual
execution times for this method.  These times grow linearly with the
size of the array.  This agrees with the expected $O(n)$ behavior.

\section{Writing is Hard}
\label{sec:writing}

As you undoubtedly know, writing is hard.  Turns out, technical
writing is extra hard.  In addition to all of the good things we
should do with spelling, grammar, punctuation, and everything else, we
need to focus on writing for a technical audience.  There is no room
for a wordy writing style.  Our goals are to be concise and precise.
If you can say it in one short sentence instead of a rambling
paragraph, do it!

Here's an example of a before and after.  Before editing:

The experiment was actually set up first.  After we set up the
experiment, we ran the experiment on all of our test cases.  These
test cases are very interesting, as we carefully selected them to be
just that.  Running the experiment allowed us to gather the timings we
present below.  The timings show a trend that looks like a linear
relationship between the input parameter and time taken.  This matches
the theoretical expectation, which is good.

After:

We set up and run our experiment on a set of carefully selected test
cases.  Results, shown below, suggest a linear relationship between
the input parameter and time taken, matching theoretical expectations.

It is perfectly reasonable to write the first paragraph initially.
However, subsequent editing, ideally by multiple people, should lead
to the much more concise and precise second paragraph.

Technical writing is a learned skill, so do not be surprised if your
early versions come under heavy editing and constructive criticism.
That said, even the most experienced technical writers need to make
multiple editing passes.

Here are a few more tips to keep in mind.

\begin{itemize}

\item Define all terms, abbreviations, and acronyms the first time you
  use them.  You should not assume your reader is aware of them, and
  it is annoying, as a reader, to feel the need to stop reading to
  look up a term or acronym.

\item Avoid contractions.

\item Carefully consider the voice to use and be consistent.
  Descriptions of work done is often done using ``we'' even in the
  case of a single author.  Avoid using too much of a storytelling
  approach.

\end{itemize}

\section{Subsections, Math, Tables, Citations}
\label{sec:subsections}

Now back to more of what \LaTeX{} can do and how to present some
common components of a technical paper.  I have subsections here.  I
can label those and refer to them as Sections~\ref{sec:subsection1}
and~\ref{sec:subsection2}.

\subsection{Part 1}
\label{sec:subsection1}

Subsections are fun.  So is math mode: $f(x) = x^4 + a_0 + x^{2k}$ is
nice, but how about $\frac{\alpha}{2} + \log x^y$?  Tip: see the file
\url{isoent-ref.pdf} for 30 pages
of special characters and how to make \LaTeX print them.

\subsection{Part 2}
\label{sec:subsection2}

Did you like the equations in Section~\ref{sec:subsection1}?  How
about a table?

{\singlespace  % since the table looks funny when double-spacing is
	       % left on
\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|l|c|c|l|}
\hline
{\bf Architecture} & {\bf IPC vehicle} & {\bf Procs.} & {\bf
Computers} & {\bf Key}
\\ \hline \hline
IBM SP2 & switch & 2--32 & 2--32 & SP2s \\ \hline
IBM SP2 & Ethernet & 2--32 & 2--32 & SP2e \\ \hline
Sun Ultra 2/2200 & shared memory & 1--2 & 1 & SUNsh \\ \hline
Sun Ultra 2/2200 & local p4 & 1--2 & 1 & SUNlp4 \\ \hline
Sun Ultra 2/2200 & Ethernet p4 & 1--6 & 1--3 & SUNp4(10) \\ \hline
Sun Ultra 2/2200 & fast Ethernet p4 & 1--6 & 1--3 & SUNp4(100) \\ \hline
SGI Onyx II & shared memory & 1--8 & 1 & SGIsh \\ \hline
SGI Onyx II & local p4 & 1--8 & 1 & SGIlp4 \\ \hline
\end{tabular}
\end{center}
\caption{Parallel environments for network performance
comparisons.  Variations in architecture, interprocess communication
vehicles, numbers of processes and computers are examined.}
\label{tab:machines}
\end{table}
}

I made Table~\ref{tab:machines}.  Isn't it nice?  It also has a
caption that describes exactly what the table is supposed to show.
This is an example of how figures and tables can float around.  This
is usually a good thing, and \LaTeX will do something reasonable.
However, it is another reason that captions must describe figures and
tables well, and that the text should refer to them as ``Table X''
rather than ``the following table.''

It is essential that you include all appropriate references.  Maybe
you used something by Barnard and Simon~\cite{barnard94a}.  Or perhaps
you want to say that several papers discuss a point you are making in
more detail~\cite{beall98a,einstein}.  I'll include a few more so you
can see examples of more types of
citations~\cite{gzip,binks99,sagan94a}.  Generally, it is better to
cite a published work rather than a URL.  If you use information from
a web page, try to find books, journal articles, conference
proceedings, or technical reports that contain the same information.
When citing web pages, try to include an accurate title and author.
Tip: open the bibtex file in emacs, and the ``Entry-Types'' menu will
give you a bunch of other templates you can use.

Lists of items are also fun.  Maybe you want bullets:

\begin{itemize}

\item Item 1.

\item Item 2.

  \begin{itemize}
   \item subitem of 2!
   \item another subitem of 2.
  \end{itemize}

\end{itemize}

Maybe you want them numbered:

\begin{enumerate}

\item I bet this gets 1.

\item And this gets 2.

\end{enumerate}

\noindent
Hey!  This paragraph isn't indented.  Sometimes that is what you
want.  And you can do it.

\section{Compiling}
\label{sec:building}

To compile this file into a device independent ({\tt .dvi}) file:

\begin{verbatim}
latex paper.tex
\end{verbatim}

Most of the time, you want PDF:

\begin{verbatim}
pdflatex paper.tex
\end{verbatim}

The procedure above will not include any of your citations.  To get
those, also run
\begin{verbatim}
bibtex paper
\end{verbatim}
after you have run latex once, then rerun the latex command, at least
twice.  Since latex is a one-pass compiler, it relies on information
it stores from previous runs to generate labels and citations and
other information.  If things seem weird, just rerun latex twice and
most of the time, things will be fine.

The included Makefile should do all of this automatically, as long as
you use GNU make ({\tt gmake}).

\section{Conclusions}
\label{sec:conclusions}

Summary, conclusions, discussion of the results, and ideas for future
work should be here.

Again, remember that writing is hard and writing well is very hard.  It
requires a long cycle of writing, reading, editing, rearranging,
rewriting, and so on.  If you are working in a group, have all group
members read and edit each others' sections.  Scientific writing
should be clear and concise.  If find a paragraph that says what can
be said in a sentence, cut it down to one sentence!

% it's nice to acknowledge any person or organization that helped out
% along the way.  The * after \section means this will not be assigned
% a section number
\section*{Acknowledgments}

The authors would like to thank all of our system administrators, past
and present, for installing the tools we needed on our systems and for
keeping those systems working smoothly while we completed this
project.

% Finally, the citations.  These lines tell latex what style to use
% (abbrv means to use first and middle initials instead of full names,
% among other things), and that your bibtex file is called
% references.bib
%
% bibtex alone should be enough to make anyone abandon any wysiwyg
% word processor and use latex for everything!
% 
% we'll also turn off double spacing for the citations
\singlespace
\bibliographystyle{abbrv}
\bibliography{references}

% tell latex we're done.  Anything beyond this line will be ignored.
\end{document}

This doesn't show up, because it came after the \end{document}

### 2020-11-13 finalPaper/probset.tex
Initial commit from latex-example

--- 2020-11-13 finalPaper/paper.tex
@@
 % 
-% Sample latex source for a project report
-%
-% This is intended as an introduction to latex, but includes some
-% advice about what you might want for the content of your paper
+% Sample latex source for a problem set
 %
 % Jim Teresco
-% Williams College, Mount Holyoke College, Siena College, The College
-% of Saint Rose
+% Williams College, Mount Holyoke College, Siena College
 %
-% Last modified: Mon Dec  4 19:03:24 EST 2017
+% Last modified: Thu Jan 27 14:56:48 EST 2011
 %
@@
 \usepackage{epsfig}        % wrapper for graphicx package
-\usepackage{times}
-\usepackage{url}
+\usepackage{times}         % nicer fonts
+\usepackage{url}           % better URL formatting
 % set some margins, these can be defined as in, cm, pt
@@
 \setlength{\textwidth}{6.5in}
+\setlength{\parindent}{0in}
 
@@
 
-%% to remove page numbers, uncomment this:
-%% \pagestyle{empty}
-
-%% Define single-space command
-\newcommand{\singlespace}{
-  \protect\renewcommand\baselinestretch{1.0}
-  \protect\normalsize
-}
-% use this instead if you want to disable it completely:
-%%\newcommand{\singlespace}{}
-
-%% Define double-space command (really more like 1.5 spacing)
-%% This is essential for rough drafts, and not a bad idea even for
-%%  final submissions
-\newcommand{\doublespace}{
-  \protect\renewcommand\baselinestretch{1.5}
-  \protect\normalsize
-}
-% use this instead if you want to disable it completely:
-%%\newcommand{\doublespace}{}
-
 % This tells latex we're done defining the preamble stuff and we're
@@
 % this removes the date that is automatically placed in the title.
-% comment it out if you want the date
-\date{}
+% uncomment it if you want the date
+%\date{}
 
-% the next few items define things to go onto the title section, like,
-% well, the title, and the list of authors
-\title{A Sample Project Report
-\footnote{This work was completed in partial fulfillment of the final
-project requirement for Computer Science 400 at Siena College, Fall 2017.}}
-
-\author{J.~D.~Teresco\footnote{Corresponding author, e-mail:
-jteresco@siena.edu} and O.-W.~Kenobi\\
-Department of Computer Science\\
-Siena College\\
-Loudonville, NY  12211
-}
-
-% this tells latex that you're done setting up title stuff and that it
-% should go ahead and generate the title here
-\maketitle
 % leave the page number off but for this page only
@@
 
-% Abstract!
-\begin{abstract}
-It's usually a good idea to include an abstract.  This should be a
-brief summary of your paper.  Highlight your important results.  A
-half a page is about right.  Maybe less.  It should be independent of
-your subsequent introduction/overview section.
-\end{abstract}
+% For a problem set, it is not necessary to have a fancy title, so
+% we just list the important information.  Note that latex will decide
+% where line breaks are appropriate but we can force a line break with
+% a \\ at the end of a line
 
-% turn on double spacing
-\doublespace
+Jim Teresco, Obi-Wan Kenobi, and Albus Dumbledore\\
+Problem Set 0\\
+Computer Science 385\\
+January 11, 2011
 
-% Now we write the text of the paper, hopefully breaking it up into
-% nice sections and subsections, using figures and tables as
-% appropriate, and referring to those sections using labels instead of
-% trying to number things by hand
-\section{Overview}
-\label{sec:overview}
+\bigskip
+\hrule
+\bigskip
 
-Begin with an introduction to your topic, specifically emphasizing
-what you have done.  Some historical background and references to
-related projects are appropriate.  You should also outline the paper
-in the overview section.  It might go something like this:
+\noindent
+1.  A simple Java program was written to copy of an array into the
+arraylist.  A method called {\tt copy()} takes the value of $n$ as its
+parameter, creates and fills an array of size $n$, creates an empty
+{\tt ArrayList}, starts a timer, copies the values from the array to
+the vector, stops the timer, and prints out the time taken.
 
-In Section~\ref{sec:vectarray} we discuss how to build a vector from
-an array.  Section~\ref{sec:subsections} has a lot of examples of
-things you might want \LaTeX~\cite{lamport86} to do.  In
-Section~\ref{sec:writing}, we discuss how writing is hard and offer
-some suggestions on how to do a better job at it.
-Section~\ref{sec:building} describes how to use \LaTeX{} to compile
-this document into printable postscript.  Finally,
-Section~\ref{sec:conclusions} summarizes the results and presents
-suggestions for future work.
-
-Notice how references are used instead of actual section numbers.
-This allows you to move, add, or remove sections and have the section
-numbers stay correct.  The names are specified below with label commands.
-
-\section{Building a Vector from an Array}
-\label{sec:vectarray}
-
-This section is here to give examples of math mode, font changes, the
-verbatim environment, and importing a postscript figure.
-
-This experiment measures the time taken to copy an array of $n$ {\tt
-int} values to a {\tt Vector} of $n$ {\tt Integer} objects.  We will
-use the {\tt add()} method of {\tt class Vector} to append new
-values to the end of the vector.  Accessing each item in the array is
-a constant time ($O(1)$) operation, as are the construction of the
-{\tt Integer} object which will hold the value in the vector and each
-{\tt add()} call.  We do this for each of $n$ items, thus, we
-expect this operation to take $O(n)$ time.  We create our vector to be
-initially of size $n$ to avoid the need to resize dynamically during
-the experiments.  We use the {\tt Vector} implementation found in the
-{\tt structure} package.
-
-{\it Now, we describe how the experiment is set up, how you run it, and
-present and discuss the results.  This sentence is mainly inserted to
-show an example of italics.}
-
-A simple Java program was written to copy of an array into the
-vector.  A method called {\tt copy()} takes the value of $n$ as its
-parameter, creates and fills an array of size $n$, creates an empty
-{\tt Vector}, starts a timer, copies the values from the array to the
-vector, stops the timer, and prints out the time taken.  The body of
-this method is shown in Figure~\ref{fig:code}.
-
-{\singlespace % we don't want to doublespace our code snippet
-\begin{figure}[htb]
+% the verbatim environment uses our formatting and a fixed-width font -- very
+% good for code snips
 \begin{verbatim}
@@
 
-    Vector v = new Vector(n);
+    ArrayList v = new ArrayList(n);
     long start=System.currentTimeMillis();
@@
 \end{verbatim}
-\caption{Java source to copy an array to a Vector.}
-\label{fig:code}
-\end{figure}
-}
 
-\begin{figure}[htb]
+The following graph shows the times in milliseconds to build as {\tt
+  ArrayList} from various array sizes. using the code above:
+
 \begin{center}
-\includegraphics[angle=-90,width=5in]{times.eps}
+\psfig{file=times.ps,angle=-90,width=5in}
 \end{center}
-\caption{Times in milliseconds to build a Vector from various array
-sizes. using the code in Figure~\ref{fig:code}}
-\label{fig:times}
-\end{figure}
 
-Your figures should have a caption that describes exactly what is in
-the figure.  A discussion of the figure should be included in your
-main text, referring to the figure by number, like so:
+The method is executed 10 times for each value of $n$, and the minimum
+time is taken.  The runs are all done on a PC with a 700 MHz Intel
+Pentium III processor, running FreeBSD 4.3 and the Java Development
+Kit version 1.3.0.  We choose array and vector sizes starting at
+$n=512$ and double them until our final size $n=262,144$.  These
+bounds, all powers of 2, were chosen because the 512-element case is
+the first which takes at least one millisecond, and the
+262,144-element is the largest which can execute without generating an
+out of memory exception from the Java run-time environment.  The graph
+shows the actual execution times for this method.  These times grow
+linearly with the size of the array.  This agrees with the expected
+$O(n)$ behavior.
 
-The method in Figure~\ref{fig:code} is executed 10 times for each
-value of $n$, and the minimum time is taken.  The runs are all done on
-a PC with a 700 MHz Intel Pentium III processor, running FreeBSD 4.3
-and the Java Development Kit version 1.3.0.  We choose array and
-vector sizes starting at $n=512$ and double them until our final size
-$n=262,144$.  These bounds, all powers of 2, were chosen because the
-512-element case is the first which takes at least one millisecond,
-and the 262,144-element is the largest which can execute without
-generating an out of memory exception from the Java run-time
-environment.  The graph in Figure~\ref{fig:times} shows the actual
-execution times for this method.  These times grow linearly with the
-size of the array.  This agrees with the expected $O(n)$ behavior.
+\bigskip
 
-\section{Writing is Hard}
-\label{sec:writing}
+2.  This answer involves some math mode: The equation $f(x) = x^4 +
+a_0 + x^{2k}$ is nice, but how about $\frac{\alpha}{2} + \log x^y$?
+Tip: see the file \url{~jteresco/shared/latexexample/isoent-ref.pdf}
+for 30 pages of special characters and how to make \LaTeX{} print
+them.
 
-As you undoubtedly know, writing is hard.  Turns out, technical
-writing is extra hard.  In addition to all of the good things we
-should do with spelling, grammar, punctuation, and everything else, we
-need to focus on writing for a technical audience.  There is no room
-for a wordy writing style.  Our goals are to be concise and precise.
-If you can say it in one short sentence instead of a rambling
-paragraph, do it!
+Here's some more math, just for fun.
 
-Here's an example of a before and after.  Before editing:
+\begin{enumerate}
+  \item $N \in \Omega(\log N)$
+  \item $5N + 6 \in \Theta(N)$
+  \item $N \log N \in \Omega(N)$
+  \item $N \in \Theta(N – N^{\frac12})$
+\end{enumerate}
 
-The experiment was actually set up first.  After we set up the
-experiment, we ran the experiment on all of our test cases.  These
-test cases are very interesting, as we carefully selected them to be
-just that.  Running the experiment allowed us to gather the timings we
-present below.  The timings show a trend that looks like a linear
-relationship between the input parameter and time taken.  This matches
-the theoretical expectation, which is good.
-
-After:
-
-We set up and run our experiment on a set of carefully selected test
-cases.  Results, shown below, suggest a linear relationship between
-the input parameter and time taken, matching theoretical expectations.
-
-It is perfectly reasonable to write the first paragraph initially.
-However, subsequent editing, ideally by multiple people, should lead
-to the much more concise and precise second paragraph.
-
-Technical writing is a learned skill, so do not be surprised if your
-early versions come under heavy editing and constructive criticism.
-That said, even the most experienced technical writers need to make
-multiple editing passes.
-
-Here are a few more tips to keep in mind.
+And some others.
 
@@
 
-\item Define all terms, abbreviations, and acronyms the first time you
-  use them.  You should not assume your reader is aware of them, and
-  it is annoying, as a reader, to feel the need to stop reading to
-  look up a term or acronym.
+\item $f(n) \in O(f(n))$
 
-\item Avoid contractions.
+\item $f(n) \in O(g(n))$ iff $g(n) \in \Omega(f(n))$
 
-\item Carefully consider the voice to use and be consistent.
-  Descriptions of work done is often done using ``we'' even in the
-  case of a single author.  Avoid using too much of a storytelling
-  approach.
+\item If $f(n) \in O(g(n))$ and $g(n) \in O(h(n))$, then $f(n) \in
+  O(h(n))$
+
+\item If $f_1(n) \in O(g_1(n))$ and $f_2(n) \in O(g_2(n))$, then
+  $f_1(n) + f_2(n) \in O(\max\{g_1(n),g_2(n)\})$
 
@@
 
-\section{Subsections, Math, Tables, Citations}
-\label{sec:subsections}
+We can do limits.  This equation will be larger and centered.
 
-Now back to more of what \LaTeX{} can do and how to present some
-common components of a technical paper.  I have subsections here.  I
-can label those and refer to them as Sections~\ref{sec:subsection1}
-and~\ref{sec:subsection2}.
+$$
+\lim_{n\to\infty}\frac{f(n)}{g(n)}
+$$
 
-\subsection{Part 1}
-\label{sec:subsection1}
+We can compute summations.
 
-Subsections are fun.  So is math mode: $f(x) = x^4 + a_0 + x^{2k}$ is
-nice, but how about $\frac{\alpha}{2} + \log x^y$?  Tip: see the file
-\url{isoent-ref.pdf} for 30 pages
-of special characters and how to make \LaTeX print them.
+$$
+\sum_{i=1}^n i = \frac{n(n-1)}{2}
+$$
 
-\subsection{Part 2}
-\label{sec:subsection2}
+Next, let's work toward a more general result:
+$$
+an^2 + bn + d \in O(n^2)
+$$
+for positive constants $a$, $b$, $d$.
 
-Did you like the equations in Section~\ref{sec:subsection1}?  How
-about a table?
+We proceed by noting that
+$$
+an^2 + bn + d \leq an^2 + bn + n
+$$
+for $n>d$, and
+$$
+an^2 + bn + n = an^2 + (b+1)n \leq an^2 + n^2
+$$
+for $n>b+1$, and
+$$
+an^2 + n^2 = (a+1)n^2
+$$
+which leads us to constants of $c=a+1$ and $n_0=\max(d,b+1)$.
 
-{\singlespace  % since the table looks funny when double-spacing is
-	       % left on
-\begin{table}[htb]
+\bigskip
+
+3.  Parallel environments for network performance comparisons.
+Variations in architecture, interprocess communication vehicles,
+numbers of processes and computers are examined.
+
 \begin{center}
@@
 \end{center}
-\caption{Parallel environments for network performance
-comparisons.  Variations in architecture, interprocess communication
-vehicles, numbers of processes and computers are examined.}
-\label{tab:machines}
-\end{table}
-}
 
-I made Table~\ref{tab:machines}.  Isn't it nice?  It also has a
-caption that describes exactly what the table is supposed to show.
-This is an example of how figures and tables can float around.  This
-is usually a good thing, and \LaTeX will do something reasonable.
-However, it is another reason that captions must describe figures and
-tables well, and that the text should refer to them as ``Table X''
-rather than ``the following table.''
+\bigskip
 
-It is essential that you include all appropriate references.  Maybe
-you used something by Barnard and Simon~\cite{barnard94a}.  Or perhaps
-you want to say that several papers discuss a point you are making in
-more detail~\cite{beall98a,einstein}.  I'll include a few more so you
-can see examples of more types of
-citations~\cite{gzip,binks99,sagan94a}.  Generally, it is better to
-cite a published work rather than a URL.  If you use information from
-a web page, try to find books, journal articles, conference
-proceedings, or technical reports that contain the same information.
-When citing web pages, try to include an accurate title and author.
-Tip: open the bibtex file in emacs, and the ``Entry-Types'' menu will
-give you a bunch of other templates you can use.
-
-Lists of items are also fun.  Maybe you want bullets:
+4.  This question's answer involves bulleted lists for some reason.
 
@@
 
-\noindent
-Hey!  This paragraph isn't indented.  Sometimes that is what you
-want.  And you can do it.
+\bigskip
 
-\section{Compiling}
-\label{sec:building}
+5. How to compile your \LaTeX{} source into something readable and printable.
 
@@
 
-Most of the time, you want PDF:
+Then convert to postscript:
 
 \begin{verbatim}
-pdflatex paper.tex
+dvips -P pdf -t letter paper.dvi -o
 \end{verbatim}
 
-The procedure above will not include any of your citations.  To get
-those, also run
+You can actually leave off the ``-P pdf -t letter'' if all you want is
+postscript.  Those options are there to allow conversion to a better
+PDF file when you run
+
 \begin{verbatim}
-bibtex paper
+ps2pdf paper.ps
 \end{verbatim}
-after you have run latex once, then rerun the latex command, at least
-twice.  Since latex is a one-pass compiler, it relies on information
-it stores from previous runs to generate labels and citations and
-other information.  If things seem weird, just rerun latex twice and
-most of the time, things will be fine.
 
-The included Makefile should do all of this automatically, as long as
-you use GNU make ({\tt gmake}).
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-Summary, conclusions, discussion of the results, and ideas for future
-work should be here.
-
-Again, remember that writing is hard and writing well is very hard.  It
-requires a long cycle of writing, reading, editing, rearranging,
-rewriting, and so on.  If you are working in a group, have all group
-members read and edit each others' sections.  Scientific writing
-should be clear and concise.  If find a paragraph that says what can
-be said in a sentence, cut it down to one sentence!
-
-% it's nice to acknowledge any person or organization that helped out
-% along the way.  The * after \section means this will not be assigned
-% a section number
-\section*{Acknowledgments}
-
-The authors would like to thank all of our system administrators, past
-and present, for installing the tools we needed on our systems and for
-keeping those systems working smoothly while we completed this
-project.
-
-% Finally, the citations.  These lines tell latex what style to use
-% (abbrv means to use first and middle initials instead of full names,
-% among other things), and that your bibtex file is called
-% references.bib
-%
-% bibtex alone should be enough to make anyone abandon any wysiwyg
-% word processor and use latex for everything!
-% 
-% we'll also turn off double spacing for the citations
-\singlespace
-\bibliographystyle{abbrv}
-\bibliography{references}
+You can view your postscript files with {\tt ghostview}, or if you
+are lucky, with {\tt gv}.
 

### 2020-11-13 finalPaper/paper.tex
Started editing template to include names and new title

@@
 % Last modified: Mon Dec  4 19:03:24 EST 2017
+%
+%
+% Modified by: Pat Baumgardner, Adam Dachenhausen, Shah Syed
+%
 %
@@
 % well, the title, and the list of authors
-\title{A Sample Project Report
-\footnote{This work was completed in partial fulfillment of the final
-project requirement for Computer Science 400 at Siena College, Fall 2017.}}
+\title{Java Based Memory Latency Simulator}
 
-\author{J.~D.~Teresco\footnote{Corresponding author, e-mail:
-jteresco@siena.edu} and O.-W.~Kenobi\\
+\author{Pat Baumgardner, Adam Dachenhausen, Shah Syed\\
 Department of Computer Science\\

### 2020-11-13 memSim.java
Added more user propmts, and finished constructor

@@ -11,7 +11,7 @@ import java.util.Scanner;
 
 class memSim{
     //An array to hold all the hard drives in the system
-    HDD[] hardDrive;
+    HDD[] hardDrives;
 
     //An array to hold all the ram sticks in the system
     RAM[] rams;
@@ -33,32 +33,61 @@ class memSim{
     /**
        A constructor for the memSim obect
      **/
-    public memSim(){
-	
+    public memSim(int numHDD, int hDDSize, int numHDDPlatters, int numRam, int rAMSize, int cacheSize){
+	//Let's first initialize the number of harddrives, and their size
+	hardDrives = new hardDrives[numHDD];
+
+	for(int i = 0; i < numHDD; i++){
+	    hardDrives[i] = new HDD(hDDSize, numHDDPlatters);
+	}
+
+	//Now let's initialize our RAM
+	rams = new RAM[numRam];
+
+	for(int i = 0; i < numRam; i++){
+	    rams[i] = new RAM(rAMSize);
+	}
+
+	//Now let's initialize our cacheSize
+	cpu = new CPU(1, cacheSize);
     }
     
     /**
        Prompts the user for specifics about the system.
-       Then initializes the system
+       Then initializes the system and returns it
 
        @return a memSim object specified by the user's requests
      **/
     private static memSim promptUser(){
-	memSim sim = new memSim();
 	Scanner scnr = new Scanner(System.in);
-	
+
+	//Let's ask the user for system parameters
 	System.out.println("How many Hard Drives would you like?");
 	String input = scnr.nextLine();
+	int numHDD = Integer.parseInt(input);
+
+	System.out.println("How many platters should each hard drive have?");
+	input = scnr.nextLine();
+	int numPlatters = Integer.parseInt(input);
 
 	System.out.println("And how big (in bytes) should each one be?");
 	input = scnr.nextLine();
+	int hDDSize = Integer.parseInt(input);
 
 	System.out.println("How many RAM sticks would you like?");
 	input = scnr.nextLine();
+	int numRAM = Integer.parseInt(input);
 
 	System.out.println("And how big (in bytes) should each one be?");
 	input = scnr.nextLine();
+	int rAMSize = Integer.parseInt(input);
 
+	System.out.println("Finally, how big (in bytes) would you like your CPU cache to be?");
+	input = scnr.nextLine();
+	int cacheSize = Integer.parseInt(input);
+
[…]

### 2020-11-13 HDD.java
Changed initializeHDD to constructor for class

@@ -19,7 +19,7 @@ class HDD{
        @param size, the size in bytes of this
        @param numPlatters, the number of single sided platters that this contains.
      **/
-    public void initializeHDD(int size, int numPlatters){
+    public HDD(int size, int numPlatters){
 	this.size=size;
 	this.numPlatters=numPlatters;
 

### 2020-11-13 CPU.java
Changed cores in constructor to be equal to param, not 0

@@ -12,7 +12,7 @@ class CPU{
     //THe size of the first cache
     int cache1Size;
     public CPU(int numCores, int cacheSize){
-     cores = 0;
+     cores = numCores;
      cache1Size = cacheSize;
     }
     

### 2020-11-14 finalPaper/paper.tex
Paper is basically setup to be filled in

@@
 
-Begin with an introduction to your topic, specifically emphasizing
-what you have done.  Some historical background and references to
-related projects are appropriate.  You should also outline the paper
-in the overview section.  It might go something like this:
+The goal of this project was to create a computer memory latency simulator. 
+For its ease of use, Java was chosen to program this project in. This 
+simulator is designed to simulate data transfer from one system component 
+to another. For example, from the hard drive disk, to the main memory.
 
-In Section~\ref{sec:vectarray} we discuss how to build a vector from
-an array.  Section~\ref{sec:subsections} has a lot of examples of
-things you might want \LaTeX~\cite{lamport86} to do.  In
-Section~\ref{sec:writing}, we discuss how writing is hard and offer
-some suggestions on how to do a better job at it.
-Section~\ref{sec:building} describes how to use \LaTeX{} to compile
-this document into printable postscript.  Finally,
-Section~\ref{sec:conclusions} summarizes the results and presents
-suggestions for future work.
+In Section~\ref{sec:memlate} we discuss abstractly how the memory latency 
+simulator works. Section ~\ref{sec:build} we describe how to build and run the 
+simulator. In section ~\ref{sec:expstats} we explain all the types of 
+information collected from the simulator. Then, in ~\ref{sec:data} we 
+present the actual data collected from the simulator, as well as actual 
+real world data. Section ~\ref{sec:disc} goes on to explain any discrepencies 
+found between our data and the real world collected data. Finally, in
+~\ref{sec:conclusions} we summarize our results, as well as report what
+we have learned, and suggest how this simulator could be better
+improved or used.
 
-Notice how references are used instead of actual section numbers.
-This allows you to move, add, or remove sections and have the section
-numbers stay correct.  The names are specified below with label commands.
+\section{Simulating memory latency}
+\label{sec:memlate}
 
-\section{Building a Vector from an Array}
-\label{sec:vectarray}
+Insert detailed description of the code here. 
+Also describe components and terms as you go.
 
-This section is here to give examples of math mode, font changes, the
-verbatim environment, and importing a postscript figure.
+\section{Building and Running the Simulator}
+\label{sec:build}
 
-This experiment measures the time taken to copy an array of $n$ {\tt
-int} values to a {\tt Vector} of $n$ {\tt Integer} objects.  We will
-use the {\tt add()} method of {\tt class Vector} to append new
-values to the end of the vector.  Accessing each item in the array is
-a constant time ($O(1)$) operation, as are the construction of the
-{\tt Integer} object which will hold the value in the vector and each
-{\tt add()} call.  We do this for each of $n$ items, thus, we
-expect this operation to take $O(n)$ time.  We create our vector to be
-initially of size $n$ to avoid the need to resize dynamically during
-the experiments.  We use the {\tt Vector} implementation found in the
-{\tt structure} package.
+Insert instructions on how to run and build here.
 
-{\it Now, we describe how the experiment is set up, how you run it, and
-present and discuss the results.  This sentence is mainly inserted to
-show an example of italics.}
+\section{Explanation of Statistics Gathered}
+\label{sec:expstats}
 
-A simple Java program was written to copy of an array into the
-vector.  A method called {\tt copy()} takes the value of $n$ as its
-parameter, creates and fills an array of size $n$, creates an empty
-{\tt Vector}, starts a timer, copies the values from the array to the
-vector, stops the timer, and prints out the time taken.  The body of
-this method is shown in Figure~\ref{fig:code}.
+Explain every statistic. 
+Also describe components and terms as you go.
 
-{\singlespace % we don't want to doublespace our code snippet
-\begin{figure}[htb]
-\begin{verbatim}
-    int[] array = new int[n];
-    int i;
-    for (i=0; i<n; i++) array[i]=i;
+\section{Data Gathered}
+\label{sec:data}
 
-    Vector v = new Vector(n);
-    long start=System.currentTimeMillis();
-    for (i=0; i<n; i++) {
-        v.add(new Integer(array[i]));
-    }
-    long end = System.currentTimeMillis();
-    System.out.println("For n="+n+", time="+(end-start));
-\end{verbatim}
-\caption{Java source to copy an array to a Vector.}
-\label{fig:code}
-\end{figure}
-}
+Insert all the data gathered by the simulator and from sources here.
+Explain any terms as needed (only if not already).
 
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[angle=-90,width=5in]{times.eps}
-\end{center}
-\caption{Times in milliseconds to build a Vector from various array
-sizes. using the code in Figure~\ref{fig:code}}
-\label{fig:times}
-\end{figure}
+\section{disc}
+\label{sec:disc}
 
-Your figures should have a caption that describes exactly what is in
-the figure.  A discussion of the figure should be included in your
-main text, referring to the figure by number, like so:
-
-The method in Figure~\ref{fig:code} is executed 10 times for each
-value of $n$, and the minimum time is taken.  The runs are all done on
-a PC with a 700 MHz Intel Pentium III processor, running FreeBSD 4.3
-and the Java Development Kit version 1.3.0.  We choose array and
-vector sizes starting at $n=512$ and double them until our final size
-$n=262,144$.  These bounds, all powers of 2, were chosen because the
-512-element case is the first which takes at least one millisecond,
-and the 262,144-element is the largest which can execute without
-generating an out of memory exception from the Java run-time
-environment.  The graph in Figure~\ref{fig:times} shows the actual
-execution times for this method.  These times grow linearly with the
-size of the array.  This agrees with the expected $O(n)$ behavior.
-
-\section{Writing is Hard}
-\label{sec:writing}
-
-As you undoubtedly know, writing is hard.  Turns out, technical
-writing is extra hard.  In addition to all of the good things we
-should do with spelling, grammar, punctuation, and everything else, we
-need to focus on writing for a technical audience.  There is no room
-for a wordy writing style.  Our goals are to be concise and precise.
-If you can say it in one short sentence instead of a rambling
-paragraph, do it!
-
-Here's an example of a before and after.  Before editing:
-
-The experiment was actually set up first.  After we set up the
-experiment, we ran the experiment on all of our test cases.  These
-test cases are very interesting, as we carefully selected them to be
-just that.  Running the experiment allowed us to gather the timings we
-present below.  The timings show a trend that looks like a linear
-relationship between the input parameter and time taken.  This matches
-the theoretical expectation, which is good.
-
-After:
-
-We set up and run our experiment on a set of carefully selected test
-cases.  Results, shown below, suggest a linear relationship between
-the input parameter and time taken, matching theoretical expectations.
-
-It is perfectly reasonable to write the first paragraph initially.
-However, subsequent editing, ideally by multiple people, should lead
-to the much more concise and precise second paragraph.
-
-Technical writing is a learned skill, so do not be surprised if your
-early versions come under heavy editing and constructive criticism.
-That said, even the most experienced technical writers need to make
-multiple editing passes.
-
-Here are a few more tips to keep in mind.
-
-\begin{itemize}
-
-\item Define all terms, abbreviations, and acronyms the first time you
-  use them.  You should not assume your reader is aware of them, and
-  it is annoying, as a reader, to feel the need to stop reading to
-  look up a term or acronym.
-
-\item Avoid contractions.
-
-\item Carefully consider the voice to use and be consistent.
-  Descriptions of work done is often done using ``we'' even in the
-  case of a single author.  Avoid using too much of a storytelling
-  approach.
-
-\end{itemize}
-
-\section{Subsections, Math, Tables, Citations}
-\label{sec:subsections}
-
-Now back to more of what \LaTeX{} can do and how to present some
-common components of a technical paper.  I have subsections here.  I
-can label those and refer to them as Sections~\ref{sec:subsection1}
-and~\ref{sec:subsection2}.
-
-\subsection{Part 1}
-\label{sec:subsection1}
-
-Subsections are fun.  So is math mode: $f(x) = x^4 + a_0 + x^{2k}$ is
-nice, but how about $\frac{\alpha}{2} + \log x^y$?  Tip: see the file
-\url{isoent-ref.pdf} for 30 pages
-of special characters and how to make \LaTeX print them.
-
-\subsection{Part 2}
-\label{sec:subsection2}
-
-Did you like the equations in Section~\ref{sec:subsection1}?  How
-about a table?
-
-{\singlespace  % since the table looks funny when double-spacing is
-	       % left on
-\begin{table}[htb]
-\begin{center}
-\begin{tabular}{|l|l|c|c|l|}
-\hline
-{\bf Architecture} & {\bf IPC vehicle} & {\bf Procs.} & {\bf
-Computers} & {\bf Key}
-\\ \hline \hline
-IBM SP2 & switch & 2--32 & 2--32 & SP2s \\ \hline
-IBM SP2 & Ethernet & 2--32 & 2--32 & SP2e \\ \hline
-Sun Ultra 2/2200 & shared memory & 1--2 & 1 & SUNsh \\ \hline
-Sun Ultra 2/2200 & local p4 & 1--2 & 1 & SUNlp4 \\ \hline
-Sun Ultra 2/2200 & Ethernet p4 & 1--6 & 1--3 & SUNp4(10) \\ \hline
-Sun Ultra 2/2200 & fast Ethernet p4 & 1--6 & 1--3 & SUNp4(100) \\ \hline
-SGI Onyx II & shared memory & 1--8 & 1 & SGIsh \\ \hline
-SGI Onyx II & local p4 & 1--8 & 1 & SGIlp4 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Parallel environments for network performance
-comparisons.  Variations in architecture, interprocess communication
-vehicles, numbers of processes and computers are examined.}
-\label{tab:machines}
-\end{table}
-}
-
-I made Table~\ref{tab:machines}.  Isn't it nice?  It also has a
-caption that describes exactly what the table is supposed to show.
-This is an example of how figures and tables can float around.  This
-is usually a good thing, and \LaTeX will do something reasonable.
-However, it is another reason that captions must describe figures and
-tables well, and that the text should refer to them as ``Table X''
-rather than ``the following table.''
-
-It is essential that you include all appropriate references.  Maybe
-you used something by Barnard and Simon~\cite{barnard94a}.  Or perhaps
-you want to say that several papers discuss a point you are making in
-more detail~\cite{beall98a,einstein}.  I'll include a few more so you
-can see examples of more types of
-citations~\cite{gzip,binks99,sagan94a}.  Generally, it is better to
-cite a published work rather than a URL.  If you use information from
-a web page, try to find books, journal articles, conference
-proceedings, or technical reports that contain the same information.
-When citing web pages, try to include an accurate title and author.
-Tip: open the bibtex file in emacs, and the ``Entry-Types'' menu will
-give you a bunch of other templates you can use.
-
-Lists of items are also fun.  Maybe you want bullets:
-
-\begin{itemize}
-
-\item Item 1.
-
-\item Item 2.
-
-  \begin{itemize}
-   \item subitem of 2!
-   \item another subitem of 2.
-  \end{itemize}
-
-\end{itemize}
-
-Maybe you want them numbered:
-
-\begin{enumerate}
-
-\item I bet this gets 1.
-
-\item And this gets 2.
-
-\end{enumerate}
-
-\noindent
-Hey!  This paragraph isn't indented.  Sometimes that is what you
-want.  And you can do it.
-
-\section{Compiling}
-\label{sec:building}
-
-To compile this file into a device independent ({\tt .dvi}) file:
-
-\begin{verbatim}
-latex paper.tex
-\end{verbatim}
-
-Most of the time, you want PDF:
-
-\begin{verbatim}
-pdflatex paper.tex
-\end{verbatim}
-
-The procedure above will not include any of your citations.  To get
-those, also run
-\begin{verbatim}
-bibtex paper
-\end{verbatim}
-after you have run latex once, then rerun the latex command, at least
-twice.  Since latex is a one-pass compiler, it relies on information
-it stores from previous runs to generate labels and citations and
-other information.  If things seem weird, just rerun latex twice and
-most of the time, things will be fine.
-
-The included Makefile should do all of this automatically, as long as
-you use GNU make ({\tt gmake}).
+Explain any discrepencies found here.
 
@@
 \end{document}
-
-This doesn't show up, because it came after the \end{document}

### 2020-11-14 finalPaper/paper.tex
Added hyperef package

@@
 \usepackage{url}
+\usepackage{hyperef}
 % set some margins, these can be defined as in, cm, pt
@@
 \begin{abstract}
-It's usually a good idea to include an abstract.  This should be a
-brief summary of your paper.  Highlight your important results.  A
-half a page is about right.  Maybe less.  It should be independent of
-your subsequent introduction/overview section.
+
+Insert abstract here.
+
 \end{abstract}
@@
 
-Insert instructions on how to run and build here.
+The simulator is written in Java, and therefore, you will need to have
+Java installed to run it. See \href{https://www.java.com/en/download/}
+{Java Download} for more.
 

### 2020-11-14 finalPaper/paper.tex
Building and Running Simulator section done

@@
 \usepackage{url}
-\usepackage{hyperef}
+\usepackage{hyperref}
 % set some margins, these can be defined as in, cm, pt
@@
 
+\subsection{Java}
 The simulator is written in Java, and therefore, you will need to have
-Java installed to run it. See \href{https://www.java.com/en/download/}
-{Java Download} for more.
+Java installed to run it. See \url{https://www.java.com/en/download/}
+ for more.
+
+\subsection{Aquiring Source Code}
+To download the simulator code, go to \url{osfinal.dachenhausen.org} 
+or
+\url{https://github.com/adamdachenhausen/cs330FinalProject} 
+and clone the repository. See \url{https://docs.github.com/en/free-pro-team@latest/github/creating-cloning-and-archiving-repositories/cloning-a-repository}
+for more.
+
+\subsection{Compiling}
+\begin{itemize}
+
+\item \section*{Command Line}
+  The source code includes two ways to compile the simulator.
+  \begin{itemize}
+  \item
+    \section*{Make}
+    The simulator source code includes a Makefile, so if Make
+    is installed, the command \begin{verbatim} make \end{verbatim}
+    will compile the simulator.
+
+  \item
+    \section*{Default}
+    In the absence of Make, the default way to compile the simulator
+    is to run
+\begin{verbatim}
+    javac *.java
+\end{verbatim}
+which will compile all of the Java files so they could be run.
+
+  \end{itemize}
+
+\item \section*{Using an IDE}
+  Given the multitude of IDEs that are available, please see your
+  specific IDE's manual for how to compile and run the simulator.
+
+\end{itemize}
+
+\subsection{Licensing}
+Before running the simulator, we would like to remind you that this
+project is protected under the MIT License, so proceed at your own risk.
+
+\subsection{Running the Simulator}
+\begin{itemize}
+\item Command Line
+  Either way that the simulator was compiled, the command
+\begin{verbatim}
+  java memSim
+\end{verbatim}
+  will start the simulator
+\item IDE
+  Given the multitude of IDEs that are available, please see your
+  specific IDE's manual for how to compile and run the simulator.
+\end{itemize}
 
@@
 
-\section{disc}
+\section{Disc}
 \label{sec:disc}

### 2020-11-14 CPU.java
Added tread run method

@@ -15,5 +15,12 @@ class CPU{
      cores = numCores;
      cache1Size = cacheSize;
     }
-    
+
+    /**
+
+     **/
+    @Override
+    public void run(){
+	
+    }
 }

### 2020-11-14 HDD.java

@@ -30,4 +30,16 @@ class HDD{
 	    platters[i] = new platter(size/numPlatters);
 	}
     }
+    /**
+       
+     **/
+    @Override
+    public void run(){
+	//First we wait a bit to emulate the hard drive getting up to speed
+
+	//Then we wait for instructions
+	while(){
+
+	}
+    }
 }

### 2020-11-14 CPU.java
Removed @Override, it presented compile issues

@@ -19,7 +19,7 @@ class CPU{
     /**
 
      **/
-    @Override
+    //@Override
     public void run(){
 	
     }

### 2020-11-14 HDD.java

@@ -33,12 +33,11 @@ class HDD{
     /**
        
      **/
-    @Override
     public void run(){
 	//First we wait a bit to emulate the hard drive getting up to speed
 
 	//Then we wait for instructions
-	while(){
+	while(!memSim.done){
 
 	}
     }

### 2020-11-14 RAM.java

@@ -11,4 +11,11 @@ class RAM{
         this.size = size;
 	data = new byte[size];
     }
+    /**
+       
+     **/
+    //@Override
+    public void run(){
+	
+    }
 }

### 2020-11-14 memSim.java
Added user option tree and added a done flag to kill simulation

@@ -10,6 +10,9 @@ import java.util.Scanner;
  */
 
 class memSim{
+    //A flag to tell our components if the system is done running
+    public static boolean done;
+    
     //An array to hold all the hard drives in the system
     HDD[] hardDrives;
 
@@ -26,16 +29,21 @@ class memSim{
      */
     public static void main(String args[]){
 	//Here we will prompt the user for data
-	memSim sim = promptUser();
-
+	memSim sim = initialUserPrompt();
 	
+	while(!done){
+	   sim.promptTree();
+	}
     }
     /**
        A constructor for the memSim obect
      **/
     public memSim(int numHDD, int hDDSize, int numHDDPlatters, int numRam, int rAMSize, int cacheSize){
-	//Let's first initialize the number of harddrives, and their size
-	hardDrives = new hardDrives[numHDD];
+	//First, we need to allow everything to know that we are starting up, aka not done
+	done = false;
+	
+	//Let's initialize the number of harddrives, and their size
+	hardDrives = new HDD[numHDD];
 
 	for(int i = 0; i < numHDD; i++){
 	    hardDrives[i] = new HDD(hDDSize, numHDDPlatters);
@@ -58,7 +66,7 @@ class memSim{
 
        @return a memSim object specified by the user's requests
      **/
-    private static memSim promptUser(){
+    private static memSim initialUserPrompt(){
 	Scanner scnr = new Scanner(System.in);
 
 	//Let's ask the user for system parameters
@@ -86,9 +94,46 @@ class memSim{
 	input = scnr.nextLine();
 	int cacheSize = Integer.parseInt(input);
 
-	//And finally initialize the systemx
-	memSim sim = new memSim(numHDD,hDDSize,numPlatters,numRam,rAMSize,cacheSize);
+	//And finally initialize the system
+	return new memSim(numHDD,hDDSize,numPlatters,numRAM,rAMSize,cacheSize);
+    }
+
+    /**
+       Gives the user a list of options to choose their next move with the simulator,
+       and acts accordingly.
+
+       
+     **/
+    private void promptTree(){
+	//First print the user's options
+	System.out.println("What would you like to do next?");
+	System.out.println("1.\tMove Data");
+	System.out.println("2.\tRead Data");
+	System.out.println("3.\tWrite Data");
+	System.out.println("4.\tExit");
+	System.out.print("Please enter the number of your selection: ");
 	
-	return sim;
+	//Then wait for their response
[…]

### 2020-11-14 memSim.java
Added help method

@@ -110,13 +110,16 @@ class memSim{
 	System.out.println("1.\tMove Data");
 	System.out.println("2.\tRead Data");
 	System.out.println("3.\tWrite Data");
-	System.out.println("4.\tExit");
+	System.out.println("4.\tHelp");
+	System.out.println("5.\tExit");
 	System.out.print("Please enter the number of your selection: ");
 	
 	//Then wait for their response
 	Scanner scnr = new Scanner(System.in);
-	int input = Integer.parseInt(scnr.nextLine());
+	String userInput = scnr.nextLine();
+	int input = Integer.parseInt(userInput);
 
+	//Process the response
 	if(input == 1){
 
 	}
@@ -127,13 +130,57 @@ class memSim{
 
 	}
 	else if(input == 4){
+	    displayHelp("");
+	}
+	else if(input == 5){
 	    done = true;
 	}
 	else{
 	    //We didn't understand the instruction, so let's just tell the user, and we will return
 	    //Then our parent will just recall us.
 	    System.out.println("Your choice was not understood. Please only enter a number");
+	    displayHelp("");
 	}
+	return;
+    }
+    /**
+       A helper method to display a user help menu
+     **/
+    private static void displayHelp(String input){
+	System.out.println();
+	System.out.println();
+	System.out.println("Java Based Memory Latency Simulator");
+	System.out.println("Developed by Pat Baumgardner, Adam Dachenhausen, Shah Syed");
+	System.out.println("Supported commands: ");
+	System.out.println("move");
+	System.out.println("read");
+	System.out.println("write");
+	System.out.println("help");
+	System.out.println("exit");
+	System.out.println("For help with a specific command type 'help [command]'");
+	if(input.equalsIgnoreCase("")){
+	    return;
+	    }
+	else if(input.equalsIgnoreCase("move")){
+	    System.out.println("Moves data from one specified location to another");
+	}
+	else if(input.equalsIgnoreCase("read")){
+	    System.out.println("Reads data from specifed location");
+	}
+	else if(input.equalsIgnoreCase("write")){
+	    System.out.println("Writes data to specified location");
+	}
+	else if(input.equalsIgnoreCase("help")){
+	    System.out.println("Displays the help message");
+	}
+	else if(input.equalsIgnoreCase("exit")){
+	    System.out.println("Exits the simulator");
+	}
+	else{
+	    System.out.println("Command not found");
+	}
+	System.out.println();
+	System.out.println();
 	return;
     }
 }

### 2020-11-14 memSim.java
Added tokenizer to promptTree, which also accepts numerical values and string equivalents

@@ -114,25 +114,26 @@ class memSim{
 	System.out.println("5.\tExit");
 	System.out.print("Please enter the number of your selection: ");
 	
-	//Then wait for their response
+	//Then wait for their response and tokenize it
 	Scanner scnr = new Scanner(System.in);
-	String userInput = scnr.nextLine();
-	int input = Integer.parseInt(userInput);
-
+	String input = scnr.nextLine();
+	String[] tokens = input.split(" ");
+	String keystone = tokens[0];
+	
 	//Process the response
-	if(input == 1){
+	if(keystone.equalsIgnoreCase("1")|keystone.equalsIgnoreCase("move")){
 
 	}
-	else if(input == 2){
+	else if(keystone.equalsIgnoreCase("2")|keystone.equalsIgnoreCase("read")){
 
 	}
-	else if(input == 3){
+	else if(keystone.equalsIgnoreCase("3")|keystone.equalsIgnoreCase("write")){
 
 	}
-	else if(input == 4){
-	    displayHelp("");
+	else if(keystone.equalsIgnoreCase("4")|keystone.equalsIgnoreCase("help")){
+	    displayHelp(tokens[1]);
 	}
-	else if(input == 5){
+	else if(keystone.equalsIgnoreCase("5")|keystone.equalsIgnoreCase("exit")){
 	    done = true;
 	}
 	else{
@@ -158,7 +159,7 @@ class memSim{
 	System.out.println("help");
 	System.out.println("exit");
 	System.out.println("For help with a specific command type 'help [command]'");
-	if(input.equalsIgnoreCase("")){
+	if(input == null){
 	    return;
 	    }
 	else if(input.equalsIgnoreCase("move")){

### 2020-11-14 CPU.java
Added thread start to constructor

@@ -14,6 +14,7 @@ class CPU{
     public CPU(int numCores, int cacheSize){
      cores = numCores;
      cache1Size = cacheSize;
+     this.run();
     }
 
     /**

### 2020-11-14 HDD.java

@@ -29,6 +29,7 @@ class HDD{
 	for(int i=0; i<numPlatters; i++){
 	    platters[i] = new platter(size/numPlatters);
 	}
+	this.run();
     }
     /**
        

### 2020-11-14 RAM.java

@@ -10,6 +10,7 @@ class RAM{
     public RAM(int size){
         this.size = size;
 	data = new byte[size];
+	this.run();
     }
     /**
        

### 2020-11-14 finalPaper/paper.tex
Fixed errors and only use 1 sub-list in compiling and running sections

@@
-% 
-% Sample latex source for a project report
-%
-% This is intended as an introduction to latex, but includes some
-% advice about what you might want for the content of your paper
-%
 % Jim Teresco
@@
 %
-%
-% First, as you may have guessed, % is the way to define a latex 
-% comment
-%
-% The first line tells latex you want for a default font size and that
-% you want to create an ``article''
 \documentclass[12pt]{article}
@@
 
-\item \section*{Command Line}
+\item \textbf{Command Line}
   The source code includes two ways to compile the simulator.
   \begin{itemize}
-  \item
-    \section*{Make}
-    The simulator source code includes a Makefile, so if Make
-    is installed, the command \begin{verbatim} make \end{verbatim}
+  \item \textbf{Make}
+  The simulator source code includes a Makefile, so if Make is installed, the command
+  \begin{verbatim}make\end{verbatim}
   will compile the simulator.
-
-  \item
-    \section*{Default}
-    In the absence of Make, the default way to compile the simulator
-    is to run
-\begin{verbatim}
-    javac *.java
-\end{verbatim}
+  \item \textbf{Default}
+  In the absence of Make, the default way to compile the simulator is to run  
+  \begin{verbatim}javac *.java\end{verbatim}
   which will compile all of the Java files so they could be run.
-
   \end{itemize}
-
-\item \section*{Using an IDE}
+\item \textbf{Using an IDE}
   Given the multitude of IDEs that are available, please see your
@@
 \begin{itemize}
-\item Command Line
+\item \textbf{Command Line}
   Either way that the simulator was compiled, the command
@@
   will start the simulator
-\item IDE
+\item \textbf{IDE}
   Given the multitude of IDEs that are available, please see your
@@
 \end{itemize}
+Upon running the simulator, you will be prompted for a variety of 
+simulator parameters. Each of these is crucial, and cannot be left blank.
+\begin{verbatim}
+How many Hard Drives would you like?
+1
+How many platters should each hard drive have?
+4
+And how big (in bytes) should each one be?
+1024
+How many RAM sticks would you like?
+2
+And how big (in bytes) should each one be?
+128
+Finally, how big (in bytes) would you like your CPU cache to be?
+32
+\end{verbatim}
+After setting up the simulator, the components will each start up, and
+you will be prompted to choose an option from the menu. You can choose to either
+select based on the menu number, or the name.
+\begin{verbatim}
+Java Based Memory Latency Simulator
+Developed by Pat Baumgardner, Adam Dachenhausen, Shah Syed
+Supported commands:
+move
+read
+write
+help
+exit
+For help with a specific command type 'help [command]'
+\end{verbatim}
 

### 2020-11-14 finalPaper/paper.tex
Finished removing filler content, so we can just drop our content in

@@
 simulator parameters. Each of these is crucial, and cannot be left blank.
+{\singlespace
 \begin{verbatim}
@@
 \end{verbatim}
+}
 After setting up the simulator, the components will each start up, and
@@
 select based on the menu number, or the name.
+{\singlespace
 \begin{verbatim}
@@
 \end{verbatim}
+}
 
-\section{Explanation of Statistics Gathered}
+\section{Explanation of Data Gathered}
 \label{sec:expstats}
@@
 
-\section{Disc}
+\section{Discrepencies}
 \label{sec:disc}
@@
 
-Summary, conclusions, discussion of the results, and ideas for future
-work should be here.
+Insert conclusions here.
 
-Again, remember that writing is hard and writing well is very hard.  It
-requires a long cycle of writing, reading, editing, rearranging,
-rewriting, and so on.  If you are working in a group, have all group
-members read and edit each others' sections.  Scientific writing
-should be clear and concise.  If find a paragraph that says what can
-be said in a sentence, cut it down to one sentence!
-
-% it's nice to acknowledge any person or organization that helped out
-% along the way.  The * after \section means this will not be assigned
-% a section number
-\section*{Acknowledgments}
-
-The authors would like to thank all of our system administrators, past
-and present, for installing the tools we needed on our systems and for
-keeping those systems working smoothly while we completed this
-project.
-
-% Finally, the citations.  These lines tell latex what style to use
-% (abbrv means to use first and middle initials instead of full names,
-% among other things), and that your bibtex file is called
-% references.bib
-%
-% bibtex alone should be enough to make anyone abandon any wysiwyg
-% word processor and use latex for everything!
-% 
-% we'll also turn off double spacing for the citations
 \singlespace

### 2020-11-15 CPU.java
Now implements Runnable. Shah and Pat on Zoom

@@ -1,7 +1,7 @@
 /*
   This simulator of a Central Processing Unit with Cache
 */
-class CPU{
+class CPU implements Runnable{
 
     //The number of cores in this
     int cores;
@@ -20,7 +20,7 @@ class CPU{
     /**
 
      **/
-    //@Override
+    @Override
     public void run(){
 	
     }

### 2020-11-15 HDD.java

@@ -1,7 +1,7 @@
 /*
   This is a simulation of a Hard Drive Disk
  */
-class HDD{
+class HDD implements Runnable{
     //The overall size of this
     int size;
 
@@ -34,9 +34,10 @@ class HDD{
     /**
        
      **/
+    @Override
     public void run(){
 	//First we wait a bit to emulate the hard drive getting up to speed
-
+	
 	//Then we wait for instructions
 	while(!memSim.done){
 

### 2020-11-15 RAM.java

@@ -1,7 +1,7 @@
 /*
   This is a simulation of Main System Memory aka Random Access Memory
  */
-class RAM{
+class RAM implements Runnable{
     //The size in bytes of this
     int size;
     //The array to hold our data
@@ -15,7 +15,7 @@ class RAM{
     /**
        
      **/
-    //@Override
+    @Override
     public void run(){
 	
     }

### 2020-11-15 CPU.java
Rollback

@@ -22,6 +22,6 @@ class CPU implements Runnable{
      **/
     @Override
     public void run(){
-	
+	System.out.println("Hello");
     }
 }

### 2020-11-15 HDD.java

@@ -1,3 +1,4 @@
+
 /*
   This is a simulation of a Hard Drive Disk
  */
@@ -20,27 +21,28 @@ class HDD implements Runnable{
        @param numPlatters, the number of single sided platters that this contains.
      **/
     public HDD(int size, int numPlatters){
-	this.size=size;
-	this.numPlatters=numPlatters;
+    this.size=size;
+    this.numPlatters=numPlatters;
 
-	// Now we are going to make the virtual hard drive
-	platters = new platter[numPlatters];
-	
-	for(int i=0; i<numPlatters; i++){
-	    platters[i] = new platter(size/numPlatters);
-	}
-	this.run();
+    // Now we are going to make the virtual hard drive
+    platters = new platter[numPlatters];
+    
+    for(int i=0; i<numPlatters; i++){
+        platters[i] = new platter(size/numPlatters);
+    }
+    this.run();
     }
     /**
        
      **/
     @Override
     public void run(){
-	//First we wait a bit to emulate the hard drive getting up to speed
-	
-	//Then we wait for instructions
-	while(!memSim.done){
-
-	}
+        //First we wait a bit to emulate the hard drive getting up to speed
+    
+    //Then we wait for instructions
+    while(!memSim.done){
+        System.out.println("Hello");     
+        
+    }
     }
 }

### 2020-11-15 RAM.java

@@ -17,6 +17,6 @@ class RAM implements Runnable{
      **/
     @Override
     public void run(){
-	
+	System.out.println("Hello");
     }
 }

### 2020-11-15 memSim.java

@@ -1,6 +1,5 @@
 import java.util.ArrayList;
 import java.util.Scanner;
-
 /*
   This is the main file for the memSim project.
 
@@ -9,7 +8,7 @@ import java.util.Scanner;
   Authors: Pat Baumgardner, Adam Dachenhausen, Shah Syed
  */
 
-class memSim{
+class memSim extends Thread{
     //A flag to tell our components if the system is done running
     public static boolean done;
     
@@ -21,7 +20,7 @@ class memSim{
     
     ArrayList processes = new ArrayList<process>();
     CPU cpu;
-
+    
     /**
        Starts the simulation of our memory management.
        First prompts the user for data
@@ -35,8 +34,9 @@ class memSim{
 	   sim.promptTree();
 	}
     }
+    
     /**
-       A constructor for the memSim obect
+       A constructor for the memSim object
      **/
     public memSim(int numHDD, int hDDSize, int numHDDPlatters, int numRam, int rAMSize, int cacheSize){
 	//First, we need to allow everything to know that we are starting up, aka not done

### 2020-11-15 CPU.java
Now extends Thread, not implements Runnable

@@ -1,7 +1,7 @@
 /*
   This simulator of a Central Processing Unit with Cache
 */
-class CPU implements Runnable{
+class CPU extends Thread{
 
     //The number of cores in this
     int cores;
@@ -14,7 +14,6 @@ class CPU implements Runnable{
     public CPU(int numCores, int cacheSize){
      cores = numCores;
      cache1Size = cacheSize;
-     this.run();
     }
 
     /**
@@ -22,6 +21,8 @@ class CPU implements Runnable{
      **/
     @Override
     public void run(){
-	System.out.println("Hello");
+	while(!memSim.done){
+
+	}
     }
 }

### 2020-11-15 HDD.java

@@ -2,7 +2,7 @@
 /*
   This is a simulation of a Hard Drive Disk
  */
-class HDD implements Runnable{
+class HDD extends Thread{
     //The overall size of this
     int size;
 
@@ -30,7 +30,6 @@ class HDD implements Runnable{
     for(int i=0; i<numPlatters; i++){
         platters[i] = new platter(size/numPlatters);
     }
-    this.run();
     }
     /**
        
@@ -41,7 +40,7 @@ class HDD implements Runnable{
     
     //Then we wait for instructions
     while(!memSim.done){
-        System.out.println("Hello");     
+        
         
     }
     }

### 2020-11-15 RAM.java

@@ -1,7 +1,7 @@
 /*
   This is a simulation of Main System Memory aka Random Access Memory
  */
-class RAM implements Runnable{
+class RAM extends Thread{
     //The size in bytes of this
     int size;
     //The array to hold our data
@@ -17,6 +17,8 @@ class RAM implements Runnable{
      **/
     @Override
     public void run(){
-	System.out.println("Hello");
+	while(!memSim.done){
+
+	}
     }
 }

### 2020-11-15 memSim.java
Now .starts() all components here

@@ -47,6 +47,7 @@ class memSim extends Thread{
 
 	for(int i = 0; i < numHDD; i++){
 	    hardDrives[i] = new HDD(hDDSize, numHDDPlatters);
+	    hardDrives[i].start();
 	}
 
 	//Now let's initialize our RAM
@@ -54,10 +55,12 @@ class memSim extends Thread{
 
 	for(int i = 0; i < numRam; i++){
 	    rams[i] = new RAM(rAMSize);
+	    rams[i].start();
 	}
 
 	//Now let's initialize our cacheSize
 	cpu = new CPU(1, cacheSize);
+	cpu.start();
     }
     
     /**

### 2020-11-15 CPU.java
All components are started from memSim and Terminal is returned to user. Closes #8

@@ -21,8 +21,8 @@ class CPU extends Thread{
      **/
     @Override
     public void run(){
-	while(!memSim.done){
-
-	}
+        while(!memSim.done){
+            
+    }
     }
 }

### 2020-11-15 HDD.java

@@ -40,7 +40,7 @@ class HDD extends Thread{
     
     //Then we wait for instructions
     while(!memSim.done){
-        
+        //System.out.println("In HDD");
         
     }
     }

### 2020-11-15 RAM.java

@@ -9,16 +9,15 @@ class RAM extends Thread{
     
     public RAM(int size){
         this.size = size;
-	data = new byte[size];
-	this.run();
+    data = new byte[size];
     }
     /**
        
      **/
     @Override
     public void run(){
-	while(!memSim.done){
-
-	}
+        while(!memSim.done){
+            //System.out.print("In Ram");
+    }
     }
 }

### 2020-11-15 memSim.java
Overhauled the input to take only ints, not whole strings to navigate. Still crashes if given non-int #7

@@ -119,30 +119,83 @@ class memSim extends Thread{
 	
 	//Then wait for their response and tokenize it
 	Scanner scnr = new Scanner(System.in);
-	String input = scnr.nextLine();
-	String[] tokens = input.split(" ");
-	String keystone = tokens[0];
+	int input = scnr.nextInt();
+
+	//Used for determining how long a command took to run
+	long start;
+	long end;
+	long diff;
 	
 	//Process the response
-	if(keystone.equalsIgnoreCase("1")|keystone.equalsIgnoreCase("move")){
+	if(input == 1){
+	    start = System.currentTimeMillis();
+	    //Here we are going to process where the user wants to move the data from and to
 
-	}
-	else if(keystone.equalsIgnoreCase("2")|keystone.equalsIgnoreCase("read")){
+	    
+	    //Then we can report the statistics
+	    end = System.currentTimeMillis();
 
+	    diff = end - start;
+	    System.out.println("The time it took to do this move operation was: " + diff + " milliseconds");
 	}
-	else if(keystone.equalsIgnoreCase("3")|keystone.equalsIgnoreCase("write")){
+	else if(input == 2){
+	    //Now we are going to ask the user where they want to read the data from
+	    System.out.println("Where would you like to read the data from?");
+	    System.out.println("1.\tHard Drive");
+	    System.out.println("2.\tRAM");
+	    System.out.println("3.\tCPU Cache");
+	    System.out.print("Please enter the number of your selection: ");
+	    
+	    int userSubSelect = scnr.nextInt();
+	    
+	    if(userSubSelect == 1){
+		System.out.println("Please enter a number from 1 - " + hardDrives.length);
+		System.out.print("Please enter the number of your selection: ");
+		userSubSelect = 2; 
+	    }
+	    else if(userSubSelect == 2){
+		System.out.println("Please enter a number from 1 - " + rams.length);
+		System.out.print("Please enter the number of your selection: ");
+	    }
+	    else if(userSubSelect == 3){
 
+	    }
+	    else{
+		System.out.println("Your choice was not understood. Please only enter a number in the valid range");
+	    }
+	    
+	    
+	    start = System.currentTimeMillis();
+	    
+	    //Then we can report the statistics
+	    end = System.currentTimeMillis();
+
+	    diff = end - start;
+	    System.out.println("The time it took to do this read operation was: " + diff + " milliseconds");
 	}
[…]

### 2020-11-15 memSim.java
Added input checking for initialUserPrompt and prompTree methods

@@ -74,34 +74,44 @@ class memSim extends Thread{
 
 	//Let's ask the user for system parameters
 	System.out.println("How many Hard Drives would you like?");
-	String input = scnr.nextLine();
-	int numHDD = Integer.parseInt(input);
+	int numHDD = getCheckedIntVal(scnr);
 
 	System.out.println("How many platters should each hard drive have?");
-	input = scnr.nextLine();
-	int numPlatters = Integer.parseInt(input);
+	int numPlatters = getCheckedIntVal(scnr);
 
 	System.out.println("And how big (in bytes) should each one be?");
-	input = scnr.nextLine();
-	int hDDSize = Integer.parseInt(input);
+	int hDDSize = getCheckedIntVal(scnr);
 
 	System.out.println("How many RAM sticks would you like?");
-	input = scnr.nextLine();
-	int numRAM = Integer.parseInt(input);
+	int numRAM = getCheckedIntVal(scnr);
 
 	System.out.println("And how big (in bytes) should each one be?");
-	input = scnr.nextLine();
-	int rAMSize = Integer.parseInt(input);
+	int rAMSize = getCheckedIntVal(scnr);
 
 	System.out.println("Finally, how big (in bytes) would you like your CPU cache to be?");
-	input = scnr.nextLine();
-	int cacheSize = Integer.parseInt(input);
+	int cacheSize = getCheckedIntVal(scnr);
 
 	//And finally initialize the system
 	return new memSim(numHDD,hDDSize,numPlatters,numRAM,rAMSize,cacheSize);
     }
 
-    /**
+    //Method to get int value
+	private static int getCheckedIntVal(Scanner scnr) {
+
+    	Integer ret = null;
+    	do {
+			String input = scnr.nextLine();
+			try {
+				ret = Integer.parseInt(input);
+			} catch (Exception e) {
+				System.out.println("Only numbers are allowed! Please provide a valid input:");
+			}
+		}while(ret == null);
+
+    	return ret;
+	}
+
+	/**
        Gives the user a list of options to choose their next move with the simulator,
        and acts accordingly.
 
@@ -122,30 +132,31 @@ class memSim extends Thread{
 	String input = scnr.nextLine();
 	String[] tokens = input.split(" ");
 	String keystone = tokens[0];
-	
-	//Process the response
-	if(keystone.equalsIgnoreCase("1")|keystone.equalsIgnoreCase("move")){
 
-	}
[…]

### 2020-11-15 memSim.java
Added more UI

@@ -149,29 +149,61 @@ class memSim extends Thread{
 	    int userSubSelect = scnr.nextInt();
 	    
 	    if(userSubSelect == 1){
-		System.out.println("Please enter a number from 1 - " + hardDrives.length);
+		System.out.println("Please choose a drive from 1 - " + hardDrives.length);
 		System.out.print("Please enter the number of your selection: ");
-		userSubSelect = 2; 
+		int hddSelect = scnr.nextInt();
+
+		
 	    }
 	    else if(userSubSelect == 2){
-		System.out.println("Please enter a number from 1 - " + rams.length);
+		System.out.println("Please choose a stick of ram from 1 - " + rams.length);
 		System.out.print("Please enter the number of your selection: ");
+		int ramSelect = scnr.nextInt();
+
+		
 	    }
 	    else if(userSubSelect == 3){
-
+		//Now we are going to ask the user where they want to write the data to
+		System.out.println("Where would you like to write the data from?");
+		System.out.println("1.\tHard Drive");
+		System.out.println("2.\tRAM");
+		System.out.println("3.\tCPU Cache");
+		System.out.print("Please enter the number of your selection: ");
+	    
+		int userSubSelect = scnr.nextInt();
+		
+		if(userSubSelect == 1){
+		    System.out.println("Please choose a drive from 1 - " + hardDrives.length);
+		    System.out.print("Please enter the number of your selection: ");
+		    int hddSelect = scnr.nextInt();
+		    
+		    
+		}
+		else if(userSubSelect == 2){
+		    System.out.println("Please choose a stick of ram from 1 - " + rams.length);
+		    System.out.print("Please enter the number of your selection: ");
+		    int ramSelect = scnr.nextInt();
+		    
+		    
+		}
+		else if(userSubSelect == 3){    
+		    start = System.currentTimeMillis();
+		    
+		    //cpu.read();
+		    
+		    //Then we can report the statistics
+		    end = System.currentTimeMillis();
+		    
+		    diff = end - start;
+		    System.out.println("The time it took to do this read operation of the CPU Cache was: " + diff + " milliseconds");
+		    
+		    
+		    
+		}
+		else{
+		    System.out.println("Your choice was not understood. Please only enter a number in the valid range");
+		}
 	    }
-	    else{
-		System.out.println("Your choice was not understood. Please only enter a number in the valid range");
-	    }
-	    
-	    
-	    start = System.currentTimeMillis();
-	    
[…]

### 2020-11-15 memSim.java
Fixed up bracket errors

@@ -1,277 +1,268 @@
 import java.util.ArrayList;
 import java.util.Scanner;
 /*
-  This is the main file for the memSim project.
+This is the main file for the memSim project.
 
-  It starts up the CPU, RAM, HDD, and any other accessories specified by the user
+It starts up the CPU, RAM, HDD, and any other accessories specified by the user
 
-  Authors: Pat Baumgardner, Adam Dachenhausen, Shah Syed
+Authors: Pat Baumgardner, Adam Dachenhausen, Shah Syed
  */
 
 class memSim extends Thread{
     //A flag to tell our components if the system is done running
     public static boolean done;
-    
+
     //An array to hold all the hard drives in the system
     HDD[] hardDrives;
 
     //An array to hold all the ram sticks in the system
     RAM[] rams;
-    
+
     ArrayList processes = new ArrayList<process>();
     CPU cpu;
-    
+
     /**
-       Starts the simulation of our memory management.
-       First prompts the user for data
-       
+    Starts the simulation of our memory management.
+    First prompts the user for data
+
      */
     public static void main(String args[]){
-	//Here we will prompt the user for data
-	memSim sim = initialUserPrompt();
-	
-	while(!done){
-	   sim.promptTree();
-	}
+        //Here we will prompt the user for data
+        memSim sim = initialUserPrompt();
+
+        while(!done){
+            sim.promptTree();
+        }
     }
-    
+
     /**
-       A constructor for the memSim object
+    A constructor for the memSim object
      **/
     public memSim(int numHDD, int hDDSize, int numHDDPlatters, int numRam, int rAMSize, int cacheSize){
-	//First, we need to allow everything to know that we are starting up, aka not done
-	done = false;
-	
-	//Let's initialize the number of harddrives, and their size
-	hardDrives = new HDD[numHDD];
+        //First, we need to allow everything to know that we are starting up, aka not done
+        done = false;
 
-	for(int i = 0; i < numHDD; i++){
-	    hardDrives[i] = new HDD(hDDSize, numHDDPlatters);
-	    hardDrives[i].start();
-	}
+        //Let's initialize the number of harddrives, and their size
+        hardDrives = new HDD[numHDD];
 
-	//Now let's initialize our RAM
-	rams = new RAM[numRam];
+        for(int i = 0; i < numHDD; i++){
+            hardDrives[i] = new HDD(hDDSize, numHDDPlatters);
[…]

### 2020-11-15 memSim.java
Fixed invalid input handling

@@ -135,12 +135,12 @@ class memSim extends Thread{
         long diff;
 
 		try {
-			int inputInt = Integer.parseInt(input);
+			int inputInt = getCheckedIntVal(scnr);
 			if(inputInt < 1 || inputInt > 5){
 				throw new Exception();
 			}
 			//Process the response
-			if (input == 1) {
+			if (inputInt == 1) {
             start = System.currentTimeMillis();
             //Here we are going to process where the user wants to move the data from and to
 
@@ -150,7 +150,7 @@ class memSim extends Thread{
             diff = end - start;
             System.out.println("The time it took to do this move operation was: " + diff + " milliseconds");
         }
-        else if(input == 2){
+        else if(inputInt == 2){
             //Now we are going to ask the user where they want to read the data from
             System.out.println("Where would you like to read the data from?");
             System.out.println("1.\tHard Drive");
@@ -158,7 +158,7 @@ class memSim extends Thread{
             System.out.println("3.\tCPU Cache");
             System.out.print("Please enter the number of your selection: ");
 
-            int userSubSelect = scnr.nextInt();
+            int userSubSelect = getCheckedIntVal(scnr);
 
             if(userSubSelect == 1){
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
@@ -180,7 +180,7 @@ class memSim extends Thread{
 
 			diff = end - start;
                 System.out.println("The time it took to do this read operation of the CPU Cache was: " + diff + " milliseconds");
-            }} else if (input == 3){
+            }} else if (inputInt == 3){
 
             //Now we are going to ask the user where they want to write the data to
             System.out.println("Where would you like to write the data from?");
@@ -189,7 +189,7 @@ class memSim extends Thread{
             System.out.println("3.\tCPU Cache");
             System.out.print("Please enter the number of your selection: ");
 
-            int userSubSelect = scnr.nextInt();
+            int userSubSelect = getCheckedIntVal(scnr);
 
             if(userSubSelect == 1){
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
@@ -219,8 +219,8 @@ class memSim extends Thread{
                 System.out.println("Your choice was not understood. Please only enter a number in the valid range");
             }
         }
-        else if(input == 4){
-			} else if (input == 5) {
+        else if(inputInt == 4){
+			} else if (inputInt == 5) {
 				done = true;
 			}
 		} catch (Exception e) {

### 2020-11-15 memSim.java
Switched to else from exception, because of the new format

@@ -134,11 +134,7 @@ class memSim extends Thread{
         long end;
         long diff;
 
-		try {
 			int inputInt = getCheckedIntVal(scnr);
-			if(inputInt < 1 || inputInt > 5){
-				throw new Exception();
-			}
 			//Process the response
 			if (inputInt == 1) {
             start = System.currentTimeMillis();
@@ -222,8 +218,7 @@ class memSim extends Thread{
         else if(inputInt == 4){
 			} else if (inputInt == 5) {
 				done = true;
-			}
-		} catch (Exception e) {
+		} else {
 			//We didn't understand the instruction, so let's just tell the user, and we will return
 			//Then our parent will just recall us.
 			System.out.println("\nYour choice was not understood. Please only enter a number in the valid range");

### 2020-11-15 CPU.java
Added read and write method

@@ -7,13 +7,37 @@ class CPU extends Thread{
     int cores;
 
     //The first cache in this
-    byte[] cache1;
+    private byte[] cache1;
 
     //THe size of the first cache
     int cache1Size;
+    /**
+       Constructor for the CPU class
+       @param numCores the number of cores that this should have (currently not supported)
+       @param cacheSize the size of the cache of this
+     **/
     public CPU(int numCores, int cacheSize){
      cores = numCores;
      cache1Size = cacheSize;
+    }
+    /**
+       Reads the data in this' cache
+       (Essentially just returns cache1)
+       there should be no delay as this is stored on RAM, and the cache
+       in a real system will be much faster
+     **/
+    public byte[] read(){
+	return cache1
+    }
+    /**
+       Writes input into the cache.
+       If input is bigger, then we will recursively write to the cache,
+       and the cache1 will hold the last cache1Size bytes
+
+       @param input the data to be written to the cache
+     **/
+    public void write(byte[] input){
+	
     }
 
     /**

### 2020-11-16 CPU.java
Added debug print line to help in CPU

@@ -9,6 +9,9 @@ class CPU extends Thread{
     //The first cache in this
     private byte[] cache1;
 
+    //The index of the last element added
+    private int cache1Last;
+    
     //THe size of the first cache
     int cache1Size;
     /**
@@ -19,6 +22,7 @@ class CPU extends Thread{
     public CPU(int numCores, int cacheSize){
      cores = numCores;
      cache1Size = cacheSize;
+     cache1Last = -1; //We hold the index of the last item in the array
     }
     /**
        Reads the data in this' cache
@@ -27,17 +31,30 @@ class CPU extends Thread{
        in a real system will be much faster
      **/
     public byte[] read(){
-	return cache1
+	return cache1;
     }
     /**
        Writes input into the cache.
-       If input is bigger, then we will recursively write to the cache,
-       and the cache1 will hold the last cache1Size bytes
+       If input is bigger, then we will iteratively write to the cache,
+       and the cache1 will hold the last cache1Size bytes.
+
+       There will also not be any delay (other than overhead) as these are the
+       fastest memory locations.
 
        @param input the data to be written to the cache
+       @return 1 if the method is done, 0 if error occurs
      **/
-    public void write(byte[] input){
-	
+    public int write(byte[] input){
+	for(int i = 0; i<input.length;i++){
+	    add(input[i]);
+	}
+	return 1;
+    }
+    private void add(byte input){
+	cache1Last++;
+	System.out.println(cache1Last+" "+cache1Size);
+	if(cache1Last == cache1Size){cache1Last = 0;}
+	cache1[cache1Last] = input;
     }
 
     /**

### 2020-11-16 memSim.java
Uncommented cpu.read/write

@@ -161,7 +161,7 @@ class memSim extends Thread{
             else if(userSubSelect == 3){
                 start = System.currentTimeMillis();
 
-                //cpu.read();
+                cpu.read();
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
@@ -196,7 +196,8 @@ class memSim extends Thread{
             else if(userSubSelect == 3){    
                 start = System.currentTimeMillis();
 
-                //cpu.write();
+		byte[] a = {1,1,1};
+                cpu.write(a);
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();

### 2020-11-16 CPU.java
Fixed null pointer error in cpu class

@@ -21,6 +21,7 @@ class CPU extends Thread{
      **/
     public CPU(int numCores, int cacheSize){
      cores = numCores;
+     cache1 = new byte[cacheSize];
      cache1Size = cacheSize;
      cache1Last = -1; //We hold the index of the last item in the array
     }

### 2020-11-16 CPU.java
Removed extraneous print lines, now accepts any int for number of bytes to write

@@ -53,7 +53,7 @@ class CPU extends Thread{
     }
     private void add(byte input){
 	cache1Last++;
-	System.out.println(cache1Last+" "+cache1Size);
+	
 	if(cache1Last == cache1Size){cache1Last = 0;}
 	cache1[cache1Last] = input;
     }

### 2020-11-16 memSim.java

@@ -196,9 +196,11 @@ class memSim extends Thread{
             else if(userSubSelect == 3){    
                 start = System.currentTimeMillis();
 
-		byte[] a = {1,1,1};
-                cpu.write(a);
+		System.out.println("How many bytes would you like to write?");
+		int in = scnr.nextInt();
 
+		cpu.write(new byte[in]);
+		
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
 

### 2020-11-16 HDD.java
All components have a read and write method. memSim's procTree is fully complete to this point, with timings

@@ -31,6 +31,13 @@ class HDD extends Thread{
         platters[i] = new platter(size/numPlatters);
     }
     }
+    public void write(byte[] input)
+    {
+	
+    }
+    public byte[] read(){
+	return null;
+    }
     /**
        
      **/

### 2020-11-16 RAM.java

@@ -11,6 +11,12 @@ class RAM extends Thread{
         this.size = size;
     data = new byte[size];
     }
+    public void write(byte[] input){
+	
+    }
+    public byte[] read(){
+	return null;
+    }
     /**
        
      **/

### 2020-11-16 memSim.java

@@ -150,12 +150,34 @@ class memSim extends Thread{
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
                 System.out.print("Please enter the number of your selection: ");
                 int hddSelect = scnr.nextInt();
+		hddSelect--;
+		start = System.currentTimeMillis();
+
+                hardDrives[hddSelect].read();
+
+                //Then we can report the statistics
+                end = System.currentTimeMillis();
+
+                diff = end - start;
+		hddSelect++;
+                System.out.println("The time it took to do this read operation of Hard Drive #"+hddSelect+" was: " + diff + " milliseconds");
 
             }
             else if(userSubSelect == 2){
                 System.out.println("Please choose a stick of ram from 1 - " + rams.length);
                 System.out.print("Please enter the number of your selection: ");
                 int ramSelect = scnr.nextInt();
+		ramSelect--;
+		start = System.currentTimeMillis();
+
+                rams[ramSelect].read();
+
+                //Then we can report the statistics
+                end = System.currentTimeMillis();
+
+                diff = end - start;
+		ramSelect--;
+                System.out.println("The time it took to do this read operation of RAM["+ramSelect+"] " + diff + " milliseconds");
 
             }
             else if(userSubSelect == 3){
@@ -185,12 +207,41 @@ class memSim extends Thread{
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
                 System.out.print("Please enter the number of your selection: ");
                 int hddSelect = scnr.nextInt();
+		hddSelect--;
+
+		start = System.currentTimeMillis();
+
+		System.out.println("How many bytes would you like to write?");
+		int in = scnr.nextInt();
+
+		hardDrives[hddSelect].write(new byte[in]);
+		
+                //Then we can report the statistics
+                end = System.currentTimeMillis();
+
+                diff = end - start;
+		hddSelect++;
+                System.out.println("The time it took to do this write operation of the Hard Drive #"+hddSelect+" was: " + diff + " milliseconds");
 
             }
             else if(userSubSelect == 2){
                 System.out.println("Please choose a stick of ram from 1 - " + rams.length);
                 System.out.print("Please enter the number of your selection: ");
                 int ramSelect = scnr.nextInt();
+		ramSelect--;
+		start = System.currentTimeMillis();
+
+		System.out.println("How many bytes would you like to write?");
+		int in = scnr.nextInt();
+
+		rams[ramSelect].write(new byte[in]);
+		
[…]

### 2020-11-16 HDD.java
All components have working read and write methods

@@ -9,34 +9,37 @@ class HDD extends Thread{
     //The spaceLeft on this
     int spaceLeft;
 
-    //The number of platters that equally divide the size
-    int numPlatters;
+    //Our array of data
+    private byte[] data;
 
-    //Our array of platters
-    platter[] platters;
+    //The index of the last element added
+    private int dataLast;
+    
     /**
        Sets up a hard drive to be added to the system.
 
        @param size, the size in bytes of this
-       @param numPlatters, the number of single sided platters that this contains.
      **/
-    public HDD(int size, int numPlatters){
-    this.size=size;
-    this.numPlatters=numPlatters;
-
-    // Now we are going to make the virtual hard drive
-    platters = new platter[numPlatters];
-    
-    for(int i=0; i<numPlatters; i++){
-        platters[i] = new platter(size/numPlatters);
+    public HDD(int size){
+	this.size=size;
+	data = new byte[size];
+	dataLast=-1;
     }
-    }
-    public void write(byte[] input)
+    public int write(byte[] input)
     {
-	
+	for(int i = 0; i<input.length;i++){
+	    add(input[i]);
+	}
+	return 1;
     }
     public byte[] read(){
-	return null;
+	return data;
+    }
+    private void add(byte input){
+	dataLast++;
+
+	if(dataLast == size){dataLast = 0;}
+	data[dataLast] = input;
     }
     /**
        

### 2020-11-16 RAM.java

@@ -6,16 +6,29 @@ class RAM extends Thread{
     int size;
     //The array to hold our data
     byte[] data;
+
+    //The index of the last element added
+    private int ramLast;
     
     public RAM(int size){
         this.size = size;
-    data = new byte[size];
+	data = new byte[size];
+	ramLast=-1;
     }
-    public void write(byte[] input){
-	
+    public int write(byte[] input){
+	for(int i = 0; i<input.length;i++){
+	    add(input[i]);
+	}
+	return 1;
     }
     public byte[] read(){
-	return null;
+	return data;
+    }
+    private void add(byte input){
+	ramLast++;
+
+	if(ramLast == size){ramLast = 0;}
+	data[ramLast] = input;
     }
     /**
        

### 2020-11-16 memSim.java

@@ -38,7 +38,7 @@ class memSim extends Thread{
     /**
     A constructor for the memSim object
      **/
-    public memSim(int numHDD, int hDDSize, int numHDDPlatters, int numRam, int rAMSize, int cacheSize){
+    public memSim(int numHDD, int hDDSize, int numRam, int rAMSize, int cacheSize){
         //First, we need to allow everything to know that we are starting up, aka not done
         done = false;
 
@@ -46,7 +46,7 @@ class memSim extends Thread{
         hardDrives = new HDD[numHDD];
 
         for(int i = 0; i < numHDD; i++){
-            hardDrives[i] = new HDD(hDDSize, numHDDPlatters);
+            hardDrives[i] = new HDD(hDDSize);
             hardDrives[i].start();
         }
 
@@ -77,10 +77,6 @@ class memSim extends Thread{
         String input = scnr.nextLine();
         int numHDD = Integer.parseInt(input);
 
-        System.out.println("How many platters should each hard drive have?");
-        input = scnr.nextLine();
-        int numPlatters = Integer.parseInt(input);
-
         System.out.println("And how big (in bytes) should each one be?");
         input = scnr.nextLine();
         int hDDSize = Integer.parseInt(input);
@@ -98,7 +94,7 @@ class memSim extends Thread{
         int cacheSize = Integer.parseInt(input);
 
         //And finally initialize the system
-        return new memSim(numHDD,hDDSize,numPlatters,numRAM,rAMSize,cacheSize);
+        return new memSim(numHDD,hDDSize,numRAM,rAMSize,cacheSize);
     }
 
     /**
@@ -176,7 +172,7 @@ class memSim extends Thread{
                 end = System.currentTimeMillis();
 
                 diff = end - start;
-		ramSelect--;
+		ramSelect++;
                 System.out.println("The time it took to do this read operation of RAM["+ramSelect+"] " + diff + " milliseconds");
 
             }

### 2020-11-16 finalPaper/paper.tex
Fixed some formatting errors. Also completed Sec. 2

@@
 In Section~\ref{sec:memlate} we discuss abstractly how the memory latency 
-simulator works. Section ~\ref{sec:build} we describe how to build and run the 
-simulator. In section ~\ref{sec:expstats} we explain all the types of 
-information collected from the simulator. Then, in ~\ref{sec:data} we 
-present the actual data collected from the simulator, as well as actual 
-real world data. Section ~\ref{sec:disc} goes on to explain any discrepencies 
-found between our data and the real world collected data. Finally, in
-~\ref{sec:conclusions} we summarize our results, as well as report what
+simulator works, and why we built it. Section~\ref{sec:build} describes
+how to build and run the simulator. In section~\ref{sec:expstats} we explain
+all the types of  information collected from the simulator. Then, in section~\ref{sec:data}
+we present the actual data collected from the simulator, as well as actual 
+real world data. Section~\ref{sec:disc} goes on to explain any discrepencies 
+found between our data and the real world collected data. Finally, in~\ref{sec:conclusions}
+we summarize our results, as well as report what
 we have learned, and suggest how this simulator could be better
@@
 
-Insert detailed description of the code here. 
-Also describe components and terms as you go.
+\subsection{Abstract Definition}
+The goal of this project was to create a Java based simulator that will allow
+semi-low access to artificial data so the user could gather statistics and trends
+from moving this data around the artificial system. The simulator consists of four
+primary parts: memSim, which is essentially the contoller; a CPU with one level of cache;
+one or more sticks or RAM; and one or more hard drive disks.
+
+Each component's (CPU, RAM, HDD), storage is represented with a byte array.
+The user then has options to handle data on any component via the memSim terminal interface.
+
+\subsection{Primary Audience}
+Today, there are many levels of abstraction between the high-level user and their hardware.
+We wanted to give a way to collect data at this low level of hardware, but not have
+to be a electrical engineer working on the circuitry.
+
+In our research, there is also very little public access on this type of data, so
+it could be useful for researchers and students alike to use this simulator to collect
+data for a variety of projects.
 

### 2020-11-17 finalPaper/paper.tex
Updated stats gathering section

@@
 
-Explain every statistic. 
+Explain every statistics.
+
+As we will be simulating the momeory simulator, we will be working on the virtual
+components. These components are those linked directly with the memory usage, these will
+include the central processing unit (CPU), the Hard Drive and the RAM sticks.
+Throughout the simulation the CPU will be accessing the RAM and the Hard Drive to process
+the data exxisting in both of these components. We are going to be timing the transfer
+of data from both the Hard Drive and RAM.
+
+The movement, reading and writing of the data to any of the components will be timed, which
+will help us collect the major statistics we need here. The latency which is going to be the
+major finding of the simulation is going to be calculted throught a stopwatch implemented
+inside our program. As we keep clocking, the time is going to be a projected ratio of the time that will be taking
+place on the basic level. Thus, the time will be reduced to the ratio fully suited for the
+level of details being dealt with in the simulator.
+
+The Memory, which is the core component of the simulator will be represented by array of bytes and
+will be
+
+The variable we are going to have in the simulation are going to be the number of Cores in the CPU,
+the number of hard drives, the size of hard drive in bytes, the number of RAM sticks, the size of
+each of the RAM stick and the size of the CPU Cache. The values for each of these variables are
+going to be input by the user of the simulator, before the program starts simulating the events.
+The simulator is programmed in a way that different values for any these variables are going to effect
+the statistics from a minor to a very significant way, depending upon the value of change and the
+effectiveness of the change in the study.
+
+CPU is going to have an option to be variable in number of cores, but for the path of the study being
+conducted we are going to mantain the number of cores to just 1 throughout. Apart from the number
+of cores, CPU will have a
+
 Also describe components and terms as you go.

### 2020-11-17 finalPaper/paper.tex
Updated stats gathering section

--- 2020-11-16 finalPaper/paper.tex
@@
 
-Explain every statistic. 
+Explain every statistics.
+
+The primary path of study through the simulator was to gather the statistics about the
+events, latencies and delays that take place in the very low level of memory movement.
+We gathered the latency time, which is the time delay taking place in data transfer,reading and
+writing at the lowest level of memory management.
+
+To reperesent this level of memory management the simulator was designed to hold some key components.
+These components are those linked directly with the memory usage,including the
+central processing unit (CPU), the Hard Drive and the RAM sticks.
+Throughout the simulation the CPU was accessing the RAM and the Hard Drive to process
+the data existing. We timed the transfer of data from both the Hard Drive and RAM.
+
+The movement, reading and writing of the data to any of the components was timed, which helped
+us to  collect the major statistics we needed  here. The latency which was  going to be the
+major finding of the simulation was calculated through a stopwatch implemented
+inside our program. As we kept clocking, the time was a projected ratio of the time that was supposed to
+take place on the basic level. Thus, the time gathered was reduced to the ratio fully suited for the
+level of details being dealt with in the simulator.
+
+The Memory, which was the core component of the simulator and the primary part of each component
+was represented by arrays of byte type. As the simulation went on, these arrays in each of the components
+were modified depending upon the event tanking place. This modification gave us the data for how the size
+of data acts proportional to the latency and the delay in the transfer, writing and reading of data. The size
+of data was one of the major field of the statistics being gathered. We studied upon the size of the
+data as how much of variance does it brings to the clock.
+
+As the variable we had in the simulation were the number of Cores in the CPU,
+the number of hard drives, the size of hard drive in bytes, the number of RAM sticks, the size of
+each of the RAM stick and the size of the CPU Cache, we kept a record of each combination. However,
+even with the CPU having a variable number of cores, for the path of study being conducted, we mantained
+it to be 1 throughout the study. The simulator was structured to take in a value for each of the variables,
+except the number of cores. This led to new fields of data for statistics being gathered.
+The size of cache in the CPU, which is a small form of memory and the quickest one out of RAM and HDD, was
+varied to yield a different result, which went into the into the table of statistics too.
+The same was done with the RAM, which relatively larger form of memory. We gathered how the data access
+and transfer on RAM vary in the aspect of Time as with different sizes, and how slower or faster was the case
+with the accessing, editing and transfering from and to the Hard drives.
+
+The number of RAM sticks and the number of hard drive was also an important variable in deciding the latency,
+the access time and the writing time, given that increasing the number of these components meant more
+size but a different access time.
+
 Also describe components and terms as you go.

### 2020-11-18 CPU.java
Added some comments and delete method header

@@ -30,6 +30,8 @@ class CPU extends Thread{
        (Essentially just returns cache1)
        there should be no delay as this is stored on RAM, and the cache
        in a real system will be much faster
+
+       @return the data read in a byte array
      **/
     public byte[] read(){
 	return cache1;
@@ -51,13 +53,31 @@ class CPU extends Thread{
 	}
 	return 1;
     }
+    /**
+       Writes one byte to this, at a current pointer so we will always
+       have the x most current bytes
+     **/
     private void add(byte input){
 	cache1Last++;
 	
 	if(cache1Last == cache1Size){cache1Last = 0;}
 	cache1[cache1Last] = input;
     }
+    /**
+       Deletes data from this, by moving the tail pointer forward
 
+       @param numBytes the number of bytes to be deleted, if greater than 
+       cache1Size, then only cache1Size bytes are deleted
+     **/
+    public void delete(int numBytes){
+	//First some error checking
+	if(numBytes>cache1Size){numBytes=cache1Size;}
+	if(numBytes<0){numBytes=0;}
+
+	for(int i=0; i<numBytes;i++){
+	    cache1Last++;
+	}
+    }
     /**
 
      **/

### 2020-11-18 HDD.java

@@ -25,6 +25,13 @@ class HDD extends Thread{
 	data = new byte[size];
 	dataLast=-1;
     }
+    /**
+       Writes a given set of bytes to this.
+
+       @param input, an array of bytes to write to this
+
+       @return 1 if the write was successful
+     **/
     public int write(byte[] input)
     {
 	for(int i = 0; i<input.length;i++){
@@ -32,14 +39,31 @@ class HDD extends Thread{
 	}
 	return 1;
     }
+    /**
+       Reads data from this
+
+       @return a byte array of the data read from this
+     **/
     public byte[] read(){
 	return data;
     }
+    /**
+       Adds one byte to this at the current pointer
+       @param input the byte to add to this
+    **/
     private void add(byte input){
 	dataLast++;
 
 	if(dataLast == size){dataLast = 0;}
 	data[dataLast] = input;
+    }
+    /**
+       Deletes data from this
+
+       
+     **/
+    public void delete(){
+
     }
     /**
        

### 2020-11-19 dataBlock.java
Initial commit

@@ -1,0 +1,17 @@
+/*
+  Represents a block of data on a HDD.
+  
+ */
+class dataBlock{
+    //The size of this block of data
+    int size;
+
+    
+    /**
+       Constructor for the dataBlock class
+       
+     **/
+    public dataBlock{
+	
+    }
+}

### 2020-11-19 memSim.java
Added system confing printout

@@ -98,6 +98,16 @@ class memSim extends Thread{
         input = scnr.nextLine();
         int cacheSize = Integer.parseInt(input);
 
+	//Let's print the system config
+	System.out.println();
+	System.out.println("***********System Config**********");
+	System.out.println(numHDD+" number of hard drives, each "+hddSize+" bytes big.");
+	System.out.println(numRAM+" sticks of RAM, each "+rAMSize+" bytes big.");
+	System.out.println("CPU cache is " + cacheSize + " bytes big.");
+	System.out.println("**********************************");
+	System.out.println();
+
+	    
         //And finally initialize the system
         return new memSim(numHDD,hDDSize,numRAM,rAMSize,cacheSize);
     }

### 2020-11-19 dataBlock.java
Updated comments, and added size var

@@ -3,15 +3,30 @@
   
  */
 class dataBlock{
-    //The size of this block of data
-    int size;
 
+    //The size of this block of data in bytes
+    public static int WIDTH = 1024;
+
+    //A flag if this is the start of a chunk (many blocks) of data
+    boolean start;
+
+    //The size of the chunk (multiple blocks) of data that this starts
+    int size;
     
     /**
-       Constructor for the dataBlock class
+       Constructor for a generic dataBlock
        
+    **/
+    public dataBlock(){
+	start = false;
+    }
+    /**
+       Constructor for a starter dataBlock for a chunk of data
+       @param start, set to true if this is the head of a series of blocks
+       @param size the size of this chunk that this heads (including this)
      **/
-    public dataBlock{
-	
+    public dataBlock(boolean start, int size){
+	this.start=start;
+	this.size=size;
     }
 }

### 2020-11-19 memSim.java
Moved memSim over to accept new data format

@@ -164,9 +164,16 @@ class memSim extends Thread{
                 System.out.print("Please enter the number of your selection: ");
                 int hddSelect = scnr.nextInt();
 		hddSelect--;
+
+		//Now we have to prompt the user to pick a chunk of data
+		hardDrives[hddSelect].print();
+		System.out.println("Please select one of the above partitions");
+		int dataSelect = scnr.nextInt();
+
+		//Then do the action
 		start = System.currentTimeMillis();
 
-                hardDrives[hddSelect].read();
+                hardDrives[hddSelect].read(dataSelect);
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
@@ -183,6 +190,12 @@ class memSim extends Thread{
 		ramSelect--;
 		start = System.currentTimeMillis();
 
+		//Now we have to prompt the user to pick a chunk of data
+		rams[ramSelect].print();
+		System.out.println("Please select one of the above partitions");
+		int dataSelect = scnr.nextInt();
+
+		//Then do the action
                 rams[ramSelect].read();
 
                 //Then we can report the statistics
@@ -294,7 +307,14 @@ class memSim extends Thread{
 		hddSelect--;
 		start = System.currentTimeMillis();
 
-                hardDrives[hddSelect].delete();
+
+		//Now we have to prompt the user to pick a chunk of data
+		hardDrives[hddSelect].print();
+		System.out.println("Please select one of the above partitions");
+		int dataSelect = scnr.nextInt();
+
+		//Then do the action		
+                hardDrives[hddSelect].delete(dataSelect);
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
@@ -309,9 +329,16 @@ class memSim extends Thread{
                 System.out.print("Please enter the number of your selection: ");
                 int ramSelect = scnr.nextInt();
 		ramSelect--;
+		
+		//Now we have to prompt the user to pick a chunk of data
+		rams[ramSelect].print();
+		System.out.println("Please select one of the above partitions");
+		int dataSelect = scnr.nextInt();
+
+		//Then do the action
 		start = System.currentTimeMillis();
 
-                rams[ramSelect].delete();
+                rams[ramSelect].delete(dataSelect);
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();

### 2020-11-19 HDD.java
Updated to use new dataBlocks, instead of byte[]. Also added new delete method

@@ -10,11 +10,8 @@ class HDD extends Thread{
     int spaceLeft;
 
     //Our array of data
-    private byte[] data;
+    private dataBlock[] data;
 
-    //The index of the last element added
-    private int dataLast;
-    
     /**
        Sets up a hard drive to be added to the system.
 
@@ -22,48 +19,130 @@ class HDD extends Thread{
      **/
     public HDD(int size){
 	this.size=size;
-	data = new byte[size];
-	dataLast=-1;
+	this.spaceLeft=size;
+	data = new dataBlock[size/dataBlock.WIDTH];
     }
     /**
        Writes a given set of bytes to this.
 
-       @param input, an array of bytes to write to this
+       @param numBytes, the number of bytes to write to this
 
-       @return 1 if the write was successful
+       @return 1 if the write was successful, 0 if not (ie not enough space)
      **/
-    public int write(byte[] input)
+    public int write(int numBytes)
     {
-	for(int i = 0; i<input.length;i++){
-	    add(input[i]);
+	//First let's check if we have enough raw space
+	if(numBytes > spaceLeft){return 0;}
+
+	//Now let's reserve our space and then write our data
+	spaceLeft =- numBytes;
+
+	//Now we call our helper method
+	int totalBlocks;
+	if(numBytes == 0){totalBlocks=0;}
+	else if(numBytes<dataBlock.WIDTH){totalBlocks = 1;}
+	else{
+	    totalBlocks = numBytes/dataBlock.WIDTH;
+	    if(numBytes%dataBlock.WIDTH!=0){totalBlocks++;}
 	}
-	return 1;
+	return add(totalBlocks);
     }
+    /**
+       FIFO add method. Finds a chunk of memory at least numBlocks big, and stores
+       numBlocks there.
+
+       @param numBlocks the number of blocks that need to be added to this
+       @return 1, if successful add, otherwise 0
+     **/
+    private int add(int numBlocks){
+	int count = 0;
+
+	//First we want to be able to loop through the entire data array
+	for(int i=0;i<data.length;i++){
+	    //Then we check if each one is open
+	    if(data[i] == null){
+		count = 1;
+		//Then we see if we have a hole big enough
+		while(data[count] == null && count != numBlocks){
+		    count++;    
+		}
+		//Then we see if the hole is big enough
+		if(count == numBlocks){
+		    //Then we can add blocks starting at i, which has to be special
[…]

### 2020-11-19 RAM.java

@@ -4,50 +4,124 @@
 class RAM extends Thread{
     //The size in bytes of this
     int size;
-    //The array to hold our data
-    byte[] data;
 
-    //The index of the last element added
-    private int ramLast;
+    //The spaceLeft in this
+    int spaceLeft;
+    
+    //The array to hold our data
+    dataBlock[] data;
     
     public RAM(int size){
         this.size = size;
-	data = new byte[size];
-	ramLast=-1;
+	this.spaceLeft = size;
+	
+	data = new dataBlock[size/dataBlock.WIDTH];
     }
     /**
        Writes and input byte array to this
-       @param input a byte array to write to this
-       @return 1, if write is successful
+       @param numBytes the number of bytes to write
+       @return 1, if write is successful, 0 if not (ie not enough space left)
      **/
-    public int write(byte[] input){
-	for(int i = 0; i<input.length;i++){
-	    add(input[i]);
+    public int write(int numBytes){
+	//First let's check if we have enough raw space
+	if(numBytes > spaceLeft){return 0;}
+
+	//Now let's reserve our space and then write our data
+	spaceLeft =- numBytes;
+
+	//Now we call our helper method
+	int totalBlocks;
+	if(numBytes == 0){totalBlocks=0;}
+	else if(numBytes<dataBlock.WIDTH){totalBlocks = 1;}
+	else{
+	    totalBlocks = numBytes/dataBlock.WIDTH;
+	    if(numBytes%dataBlock.WIDTH!=0){totalBlocks++;}
 	}
-	return 1;
+	return add(totalBlocks);
+	
+    }
+    private int add(int numBlocks){
+	int count = 0;
+
+	//First we want to be able to loop through the entire data array
+	for(int i=0;i<data.length;i++){
+	    //Then we check if each one is open
+	    if(data[i] == null){
+		count = 1;
+		//Then we see if we have a hole big enough
+		while(data[count] == null && count != numBlocks){
+		    count++;
+		}
+		//Then we see if the hole is big enough
+		if(count == numBlocks){
+		    //Then we can add blocks starting at i, which has to be special
+		    data[i] = new dataBlock(true,numBlocks);
+		    count--;
+		    i++;
+		    while(count!=0){
+			data[i] = new dataBlock();
+			count--;
+			i++;
+		    }
+		    //Since we aded the chunk, we can just return
+		    return 1;
+		}
[…]

### 2020-11-19 memSim.java

@@ -101,7 +101,7 @@ class memSim extends Thread{
 	//Let's print the system config
 	System.out.println();
 	System.out.println("***********System Config**********");
-	System.out.println(numHDD+" number of hard drives, each "+hddSize+" bytes big.");
+	System.out.println(numHDD+" number of hard drives, each "+hDDSize+" bytes big.");
 	System.out.println(numRAM+" sticks of RAM, each "+rAMSize+" bytes big.");
 	System.out.println("CPU cache is " + cacheSize + " bytes big.");
 	System.out.println("**********************************");
@@ -196,7 +196,7 @@ class memSim extends Thread{
 		int dataSelect = scnr.nextInt();
 
 		//Then do the action
-                rams[ramSelect].read();
+                rams[ramSelect].read(dataSelect);
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
@@ -221,7 +221,7 @@ class memSim extends Thread{
         else if(input == 3){
 	    //We write data
             //Now we are going to ask the user where they want to write the data to
-            System.out.println("Where would you like to write the data from?");
+            System.out.println("Where would you like to write the data to?");
             System.out.println("1.\tHard Drive");
             System.out.println("2.\tRAM");
             System.out.println("3.\tCPU Cache");
@@ -240,7 +240,7 @@ class memSim extends Thread{
 		System.out.println("How many bytes would you like to write?");
 		int in = scnr.nextInt();
 
-		hardDrives[hddSelect].write(new byte[in]);
+		hardDrives[hddSelect].write(in);
 		
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
@@ -260,7 +260,7 @@ class memSim extends Thread{
 		System.out.println("How many bytes would you like to write?");
 		int in = scnr.nextInt();
 
-		rams[ramSelect].write(new byte[in]);
+		rams[ramSelect].write(in);
 		
                 //Then we can report the statistics
                 end = System.currentTimeMillis();

### 2020-11-19 HDD.java
Implemented some runtime hotfixes

@@ -20,7 +20,14 @@ class HDD extends Thread{
     public HDD(int size){
 	this.size=size;
 	this.spaceLeft=size;
-	data = new dataBlock[size/dataBlock.WIDTH];
+	int totalBlocks;
+	if(size == 0){totalBlocks=0;}
+	else if(size<dataBlock.WIDTH){totalBlocks = 1;}
+	else{
+	    totalBlocks = size/dataBlock.WIDTH;
+	    if(size%dataBlock.WIDTH!=0){totalBlocks++;}
+	}
+	data = new dataBlock[totalBlocks];
     }
     /**
        Writes a given set of bytes to this.
@@ -61,9 +68,8 @@ class HDD extends Thread{
 	for(int i=0;i<data.length;i++){
 	    //Then we check if each one is open
 	    if(data[i] == null){
-		count = 1;
 		//Then we see if we have a hole big enough
-		while(data[count] == null && count != numBlocks){
+		while(count != numBlocks && data[count] == null ){
 		    count++;    
 		}
 		//Then we see if the hole is big enough
@@ -104,7 +110,7 @@ class HDD extends Thread{
 	//Loop through the data until we hit a null block or
 	//If we have two blocks back-to-back, then the start flag
 	//will be high, so we can stop reading our dataBlock
-	while(data[index]!=null || data[index].start==true){
+	while(index<data.length || data[index]!=null || data[index].start==true){
 	    index++;   
 	    
 	}

### 2020-11-19 HDD.java
Implemented some runtime hotfixes

@@ -110,7 +110,7 @@ class HDD extends Thread{
 	//Loop through the data until we hit a null block or
 	//If we have two blocks back-to-back, then the start flag
 	//will be high, so we can stop reading our dataBlock
-	while(index<data.length || data[index]!=null || data[index].start==true){
+	while(index<data.length && (data[index]!=null || data[index].start==true)){
 	    index++;   
 	    
 	}
@@ -124,7 +124,7 @@ class HDD extends Thread{
 	//Loop through the data until we hit a null block or
 	//If we have two blocks back-to-back, then the start flag
 	//will be high, so we can stop reading our dataBlock
-	while(data[index]!=null || data[index].start==true){
+	while(index<data.length && (data[index]!=null || data[index].start==true)){
 	    //Delete the current block
 	    data[index] = null;
 

### 2020-11-19 RAM.java

@@ -15,7 +15,15 @@ class RAM extends Thread{
         this.size = size;
 	this.spaceLeft = size;
 	
-	data = new dataBlock[size/dataBlock.WIDTH];
+	int totalBlocks;
+	if(size == 0){totalBlocks=0;}
+	else if(size<dataBlock.WIDTH){totalBlocks = 1;}
+	else{
+	    totalBlocks = size/dataBlock.WIDTH;
+	    if(size%dataBlock.WIDTH!=0){totalBlocks++;}
+	}
+	
+	data = new dataBlock[totalBlocks];
     }
     /**
        Writes and input byte array to this
@@ -49,7 +57,7 @@ class RAM extends Thread{
 	    if(data[i] == null){
 		count = 1;
 		//Then we see if we have a hole big enough
-		while(data[count] == null && count != numBlocks){
+		while(count != numBlocks && data[count] == null){
 		    count++;
 		}
 		//Then we see if the hole is big enough
@@ -84,8 +92,9 @@ class RAM extends Thread{
 	//Loop through the data until we hit a null block or
 	//If we have two blocks back-to-back, then the start flag
 	//will be high, so we can stop reading our dataBlock
-	while(data[index]!=null || data[index].start==true){
-	    index++;
+	while(index<data.length && (data[index]!=null || data[index].start==true)){
+	    index++;   
+	    
 	}
     }
     /**
@@ -97,7 +106,7 @@ class RAM extends Thread{
 	//Loop through the data until we hit a null block
 	//If we have two blocks back-to-back, then the start flag
 	//will be high, so we can stop reading our dataBlock
-	while(data[index]!=null || data[index].start==true){
+	while(index<data.length && (data[index]!=null || data[index].start==true)){
 	    //Delete the current block
 	    data[index]=null;
 

### 2020-11-19 memSim.java

@@ -124,8 +124,8 @@ class memSim extends Thread{
         System.out.println("2.\tRead Data");
         System.out.println("3.\tWrite Data");
 	System.out.println("4.\tDelete Data");
-        System.out.println("4.\tHelp");
-        System.out.println("5.\tExit");
+        System.out.println("5.\tHelp");
+        System.out.println("6.\tExit");
         System.out.print("Please enter the number of your selection: ");
 
         //Then wait for their response and tokenize it

### 2020-11-19 RAM.java
Added check in delete to make sure we aren't accessing a null array

@@ -15,37 +15,23 @@ class RAM extends Thread{
         this.size = size;
 	this.spaceLeft = size;
 	
-	int totalBlocks;
-	if(size == 0){totalBlocks=0;}
-	else if(size<dataBlock.WIDTH){totalBlocks = 1;}
-	else{
-	    totalBlocks = size/dataBlock.WIDTH;
-	    if(size%dataBlock.WIDTH!=0){totalBlocks++;}
-	}
-	
-	data = new dataBlock[totalBlocks];
+	data = new dataBlock[size/dataBlock.WIDTH];
     }
     /**
        Writes and input byte array to this
-       @param numBytes the number of bytes to write
+       @param input the number of bytes to write
        @return 1, if write is successful, 0 if not (ie not enough space left)
      **/
-    public int write(int numBytes){
-	//First let's check if we have enough raw space
-	if(numBytes > spaceLeft){return 0;}
+    public int write(int input){
+	//First let's check that we have enough space
+	if(input > spaceLeft){return 0;}
 
-	//Now let's reserve our space and then write our data
-	spaceLeft =- numBytes;
+	//Now let's reserve that space, and write our data
+	spaceLeft =- input;
 
 	//Now we call our helper method
-	int totalBlocks;
-	if(numBytes == 0){totalBlocks=0;}
-	else if(numBytes<dataBlock.WIDTH){totalBlocks = 1;}
-	else{
-	    totalBlocks = numBytes/dataBlock.WIDTH;
-	    if(numBytes%dataBlock.WIDTH!=0){totalBlocks++;}
-	}
-	return add(totalBlocks);
+	return add(input/dataBlock.WIDTH);
+
 	
     }
     private int add(int numBlocks){
@@ -57,7 +43,7 @@ class RAM extends Thread{
 	    if(data[i] == null){
 		count = 1;
 		//Then we see if we have a hole big enough
-		while(count != numBlocks && data[count] == null){
+		while(data[count] == null && count != numBlocks){
 		    count++;
 		}
 		//Then we see if the hole is big enough
@@ -92,9 +78,8 @@ class RAM extends Thread{
 	//Loop through the data until we hit a null block or
 	//If we have two blocks back-to-back, then the start flag
 	//will be high, so we can stop reading our dataBlock
-	while(index<data.length && (data[index]!=null || data[index].start==true)){
-	    index++;   
-	    
+	while(data[index]!=null || data[index].start==true){
+	    index++;
 	}
     }
     /**
[…]

### 2020-11-19 HDD.java
Fixed print crashing when data was empty

@@ -1,6 +1,6 @@
 
 /*
-  This is a simulation of a Hard Drive Disk
+This is a simulation of a Hard Drive Disk
  */
 class HDD extends Thread{
     //The overall size of this
@@ -13,154 +13,161 @@ class HDD extends Thread{
     private dataBlock[] data;
 
     /**
-       Sets up a hard drive to be added to the system.
+    Sets up a hard drive to be added to the system.
 
-       @param size, the size in bytes of this
+    @param size, the size in bytes of this
      **/
     public HDD(int size){
-	this.size=size;
-	this.spaceLeft=size;
-	int totalBlocks;
-	if(size == 0){totalBlocks=0;}
-	else if(size<dataBlock.WIDTH){totalBlocks = 1;}
-	else{
-	    totalBlocks = size/dataBlock.WIDTH;
-	    if(size%dataBlock.WIDTH!=0){totalBlocks++;}
-	}
-	data = new dataBlock[totalBlocks];
+        this.size=size;
+        this.spaceLeft=size;
+        int totalBlocks;
+        if(size == 0){totalBlocks=0;}
+        else if(size<dataBlock.WIDTH){totalBlocks = 1;}
+        else{
+            totalBlocks = size/dataBlock.WIDTH;
+            if(size%dataBlock.WIDTH!=0){totalBlocks++;}
+        }
+        data = new dataBlock[totalBlocks];
     }
+
     /**
-       Writes a given set of bytes to this.
+    Writes a given set of bytes to this.
 
-       @param numBytes, the number of bytes to write to this
+    @param numBytes, the number of bytes to write to this
 
-       @return 1 if the write was successful, 0 if not (ie not enough space)
+    @return 1 if the write was successful, 0 if not (ie not enough space)
      **/
     public int write(int numBytes)
     {
-	//First let's check if we have enough raw space
-	if(numBytes > spaceLeft){return 0;}
+        //First let's check if we have enough raw space
+        if(numBytes > spaceLeft){return 0;}
 
-	//Now let's reserve our space and then write our data
-	spaceLeft =- numBytes;
+        //Now let's reserve our space and then write our data
+        spaceLeft =- numBytes;
 
-	//Now we call our helper method
-	int totalBlocks;
-	if(numBytes == 0){totalBlocks=0;}
-	else if(numBytes<dataBlock.WIDTH){totalBlocks = 1;}
-	else{
[…]

### 2020-11-19 RAM.java

@@ -1,5 +1,5 @@
 /*
-  This is a simulation of Main System Memory aka Random Access Memory
+This is a simulation of Main System Memory aka Random Access Memory
  */
 class RAM extends Thread{
     //The size in bytes of this
@@ -7,125 +7,132 @@ class RAM extends Thread{
 
     //The spaceLeft in this
     int spaceLeft;
-    
+
     //The array to hold our data
     dataBlock[] data;
-    
+
     public RAM(int size){
         this.size = size;
-	this.spaceLeft = size;
-	
-	data = new dataBlock[size/dataBlock.WIDTH];
+        this.spaceLeft = size;
+
+        data = new dataBlock[size/dataBlock.WIDTH];
     }
+
     /**
-       Writes and input byte array to this
-       @param input the number of bytes to write
-       @return 1, if write is successful, 0 if not (ie not enough space left)
+    Writes and input byte array to this
+    @param input the number of bytes to write
+    @return 1, if write is successful, 0 if not (ie not enough space left)
      **/
     public int write(int input){
-	//First let's check that we have enough space
-	if(input > spaceLeft){return 0;}
+        //First let's check that we have enough space
+        if(input > spaceLeft){return 0;}
 
-	//Now let's reserve that space, and write our data
-	spaceLeft =- input;
+        //Now let's reserve that space, and write our data
+        spaceLeft =- input;
 
-	//Now we call our helper method
-	return add(input/dataBlock.WIDTH);
+        //Now we call our helper method
+        return add(input/dataBlock.WIDTH);
 
-	
     }
+
     private int add(int numBlocks){
-	int count = 0;
+        int count = 0;
 
-	//First we want to be able to loop through the entire data array
-	for(int i=0;i<data.length;i++){
-	    //Then we check if each one is open
-	    if(data[i] == null){
-		count = 1;
-		//Then we see if we have a hole big enough
-		while(data[count] == null && count != numBlocks){
-		    count++;
-		}
-		//Then we see if the hole is big enough
-		if(count == numBlocks){
-		    //Then we can add blocks starting at i, which has to be special
-		    data[i] = new dataBlock(true,numBlocks);
-		    count--;
-		    i++;
-		    while(count!=0){
-			data[i] = new dataBlock();
-			count--;
-			i++;
-		    }
-		    //Since we aded the chunk, we can just return
-		    return 1;
-		}
-		//If we got here, that means we don't have a hole big enough, so let's reset count
[…]

### 2020-11-19 memSim.java
Completed move method

@@ -139,9 +139,94 @@ class memSim extends Thread{
 
         //Process the response
         if(input == 1){
+	    //First let's prompt the user
+	    System.out.println("Where would you like to read the data from?");
+	    System.out.println("1.\tHard Drive");
+            System.out.println("2.\tRAM");
+            System.out.println("3.\tCPU Cache");
+            System.out.print("Please enter the number of your selection: ");
+
+            int userSubSelect1 = scnr.nextInt();
+	    int userSubSelect2 = userSubSelect1;
+	    int dataSelect1 = -1;
+	    int dataSelect2 = -1;
+
+	    if(userSubSelect1 == 1){
+                System.out.println("Please choose a drive from 1 - " + hardDrives.length);
+                System.out.print("Please enter the number of your selection: ");
+                int hddSelect = scnr.nextInt();
+		hddSelect--;
+
+		//Now we have to prompt the user to pick a chunk of data
+		hardDrives[hddSelect].print();
+		System.out.println("Please select one of the above partitions");
+		dataSelect1 = scnr.nextInt();
+
+            }
+            else if(userSubSelect1 == 2){
+                System.out.println("Please choose a stick of ram from 1 - " + rams.length);
+                System.out.print("Please enter the number of your selection: ");
+                int ramSelect = scnr.nextInt();
+		ramSelect--;
+		
+		//Now we have to prompt the user to pick a chunk of data
+		rams[ramSelect].print();
+		System.out.println("Please select one of the above partitions");
+		dataSelect1 = scnr.nextInt();
+
+            }
+            else if(userSubSelect1 == 3){
+		dataSelect1 = 1;
+            }
+	    
+	    //Now we prompt where the user wants to put this data
+	    while(userSubSelect1 == userSubSelect2){
+	    System.out.println("Where would you like to read the data from?");
+	    System.out.println("1.\tHard Drive");
+            System.out.println("2.\tRAM");
+            System.out.println("3.\tCPU Cache");
+            System.out.print("Please enter the number of your selection: ");
+
+	    userSubSelect2 = scnr.nextInt();
+
+	    if(userSubSelect2 == 1){
+                System.out.println("Please choose a drive from 1 - " + hardDrives.length);
+                System.out.print("Please enter the number of your selection: ");
+                int hddSelect = scnr.nextInt();
+		hddSelect--;
+
+		//Now we have to prompt the user to pick a chunk of data
+		hardDrives[hddSelect].print();
[…]

### 2020-11-19 memSim.java
Now using ANSI to make colored text!

@@ -126,7 +126,7 @@ class memSim extends Thread{
 	System.out.println("4.\tDelete Data");
         System.out.println("5.\tHelp");
         System.out.println("6.\tExit");
-        System.out.print("Please enter the number of your selection: ");
+        System.out.print("\u001B[33mPlease enter the number of your selection: \u001B[0m");
 
         //Then wait for their response and tokenize it
         Scanner scnr = new Scanner(System.in);
@@ -144,7 +144,7 @@ class memSim extends Thread{
 	    System.out.println("1.\tHard Drive");
             System.out.println("2.\tRAM");
             System.out.println("3.\tCPU Cache");
-            System.out.print("Please enter the number of your selection: ");
+            System.out.print("\u001B[33mPlease enter the number of your selection: \u001B[0m");
 
             int userSubSelect1 = scnr.nextInt();
 	    int userSubSelect2 = userSubSelect1;
@@ -153,25 +153,25 @@ class memSim extends Thread{
 
 	    if(userSubSelect1 == 1){
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
-                System.out.print("Please enter the number of your selection: ");
+                System.out.print("\u001B[33mPlease enter the number of your selection: \u001B[0m");
                 int hddSelect = scnr.nextInt();
 		hddSelect--;
 
 		//Now we have to prompt the user to pick a chunk of data
 		hardDrives[hddSelect].print();
-		System.out.println("Please select one of the above partitions");
+		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
 		dataSelect1 = scnr.nextInt();
 
             }
             else if(userSubSelect1 == 2){
                 System.out.println("Please choose a stick of ram from 1 - " + rams.length);
-                System.out.print("Please enter the number of your selection: ");
+                System.out.print("\u001B[33mPlease enter the number of your selection: \u001B[0m");
                 int ramSelect = scnr.nextInt();
 		ramSelect--;
 		
 		//Now we have to prompt the user to pick a chunk of data
 		rams[ramSelect].print();
-		System.out.println("Please select one of the above partitions");
+		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
 		dataSelect1 = scnr.nextInt();
 
             }
@@ -185,38 +185,38 @@ class memSim extends Thread{
 	    System.out.println("1.\tHard Drive");
             System.out.println("2.\tRAM");
             System.out.println("3.\tCPU Cache");
[…]

### 2020-11-19 memSim.java
Added more ANSI, and help message complete

@@ -451,7 +451,7 @@ class memSim extends Thread{
 	}
         else if(input == 5){
 	    //We display the help message
-	    
+	    displayHelp();
         }
         else if(input == 6){
 	    //We exit the simulator
@@ -462,7 +462,7 @@ class memSim extends Thread{
             //We didn't understand the instruction, so let's just tell the user, and we will return
             //Then our parent will just recall us.
             System.out.println("\u001B[41mYour choice was not understood. Please only enter a number in the valid range\u001B[0m");
-            displayHelp("");
+            displayHelp();
         }
 
         return;
@@ -471,7 +471,7 @@ class memSim extends Thread{
     /**
     A helper method to display a user help menu
      **/
-    private static void displayHelp(String input){
+    private static void displayHelp(){
         System.out.println();
         System.out.println();
         System.out.println("Java Based Memory Latency Simulator");
@@ -482,28 +482,42 @@ class memSim extends Thread{
         System.out.println("write");
         System.out.println("help");
         System.out.println("exit");
-        System.out.println("For help with a specific command type 'help [command]'");
-        if(input == null){
-            return;
-        }
-        else if(input.equalsIgnoreCase("move")){
-            System.out.println("Moves data from one specified location to another");
-        }
-        else if(input.equalsIgnoreCase("read")){
-            System.out.println("Reads data from specifed location");
-        }
-        else if(input.equalsIgnoreCase("write")){
-            System.out.println("Writes data to specified location");
-        }
-        else if(input.equalsIgnoreCase("help")){
-            System.out.println("Displays the help message");
-        }
-        else if(input.equalsIgnoreCase("exit")){
-            System.out.println("Exits the simulator");
-        }
-        else{
-            System.out.println("Command not found");
-        }
+        System.out.println("For help with a specific command type [command]");
+	System.out.println("Type 'q' to quit the help menu");
+	Scanner scnr = new Scanner(System.in);
+	String input = scnr.next();
+	while(!input.equalsIgnoreCase("q")){
+	    //First display the info requested
+	    System.out.print("\u001B[36m");
+	    if(input == null){
+		return;
+	    }
+	    else if(input.equalsIgnoreCase("move")){
+		System.out.println("Moves data from one specified location to another");
+	    }
+	    else if(input.equalsIgnoreCase("read")){
+		System.out.println("Reads data from specifed location");
+	    }
[…]

### 2020-11-19 memSim.java
Implemented a new load feature

@@ -77,39 +77,125 @@ class memSim extends Thread{
     private static memSim initialUserPrompt(){
         Scanner scnr = new Scanner(System.in);
 
-        //Let's ask the user for system parameters
-        System.out.println("How many Hard Drives would you like?");
-        String input = scnr.nextLine();
-        int numHDD = Integer.parseInt(input);
-
-        System.out.println("And how big (in bytes) should each one be?");
-        input = scnr.nextLine();
-        int hDDSize = Integer.parseInt(input);
-
-        System.out.println("How many RAM sticks would you like?");
-        input = scnr.nextLine();
-        int numRAM = Integer.parseInt(input);
-
-        System.out.println("And how big (in bytes) should each one be?");
-        input = scnr.nextLine();
-        int rAMSize = Integer.parseInt(input);
-
-        System.out.println("Finally, how big (in bytes) would you like your CPU cache to be?");
-        input = scnr.nextLine();
-        int cacheSize = Integer.parseInt(input);
-
-	//Let's print the system config
-	System.out.println();
-	System.out.println("***********System Config**********");
-	System.out.println(numHDD+" number of hard drives, each "+hDDSize+" bytes big.");
-	System.out.println(numRAM+" sticks of RAM, each "+rAMSize+" bytes big.");
-	System.out.println("CPU cache is " + cacheSize + " bytes big.");
-	System.out.println("**********************************");
-	System.out.println();
-
+	//Let's first ask the user if they want to make their own system, or load one
+	System.out.println("Would you like to load a system, or configure your own?");
+	System.out.println("Type \u001B[33mLOAD\u001B[0m or \u001B[33mNEW\u001B[0m: ");
+	String choice = scnr.next();
+	if(choice.equalsIgnoreCase("Load")){
+	    //Firt we display the choices that the user can choose from
+	    System.out.println("\u001B[35mSystem 1:\u001B[0m");
+	    System.out.println("\t2 HDDs with size: 10240 bytes each");
+	    System.out.println("\t2 Sticks of RAM with size: 5120 bytes each");
+	    System.out.println("\t CPU Cache size: 1024 bytes");
+	    System.out.println("\tThis system has 3 partitions of data written to HDD 1, each 1024 bytes,");
+	    System.out.println("\tand 2 partitions of data written to HDD 2, each 2048 bytes,");
[…]

### 2020-11-19 memSim.java
Fixed not being able to load system 1

@@ -110,7 +110,7 @@ class memSim extends Thread{
 	    int loadSelect = scnr.nextInt();
 	    if(loadSelect == 2){
 		//First initialize an empty memSim
-		memSim finger = new memSim(2,102400,2,51200,2048);
+		memSim finger = new memSim(1,102400,1,51200,2048);
 
 		//Then write the specified data
 		for(int i=0;i<10;i++){
@@ -135,7 +135,7 @@ class memSim extends Thread{
 		//or if the user just wants to load 1
 
 		//First initialize an empty memSim
-		memSim finger = new memSim(1,10240,1,51200,1024);
+		memSim finger = new memSim(2,102400,2,51200,1024);
 
 		//Then write the specified data
 		for(int i=0;i<3;i++){

### 2020-11-19 memSim.java
Added warning to system 3 load

@@ -100,7 +100,7 @@ class memSim extends Thread{
 	    System.out.println("\tand 3 partitions of 1024 bytes, and 1 of 2048 bytes written to the RAM,");
 	    System.out.println("\tand there is nothing in the cache.");
 	    
-	    System.out.println("\u001B[35mSystem 3:\u001B[0m");
+	    System.out.println("\u001B[35mSystem 3: \u001B[31mRUN AT YOUR OWN RISK\u001B[0m");
 	    System.out.println("\t1 HDD with size: 1024000 bytes");
 	    System.out.println("\t1 Stick of RAM with size: 51200 bytes");
 	    System.out.println("\t CPU Cache size: 2048 bytes");
@@ -135,7 +135,7 @@ class memSim extends Thread{
 		//or if the user just wants to load 1
 
 		//First initialize an empty memSim
-		memSim finger = new memSim(2,102400,2,51200,1024);
+		memSim finger = new memSim(2,10240,2,51200,1024);
 
 		//Then write the specified data
 		for(int i=0;i<3;i++){

### 2020-11-20 memSim.java
fixed NEW prompt

@@ -80,7 +80,7 @@ class memSim extends Thread{
 	//Let's first ask the user if they want to make their own system, or load one
 	System.out.println("Would you like to load a system, or configure your own?");
 	System.out.println("Type \u001B[33mLOAD\u001B[0m or \u001B[33mNEW\u001B[0m: ");
-	String choice = scnr.next();
+	String choice = scnr.nextLine();
 	if(choice.equalsIgnoreCase("Load")){
 	    //Firt we display the choices that the user can choose from
 	    System.out.println("\u001B[35mSystem 1:\u001B[0m");

### 2020-11-20 finalPaper/paper.tex
Updated paper

@@
 
-Explain every statistics.
+The primary path of study through the simulator was to gather the statistics about the
+events, latencies and delays that take place in the very low level of memory management.
+Memory management as the name suggests is the act of managing memory in the low level of
+a computer. Often represented by allocation of portions and freeing them once they are
+done being used.
 
-The primary path of study through the simulator was to gather the statistics about the
-events, latencies and delays that take place in the very low level of memory movement.
+If we consider memory management as the super class, then the statistics that we gathered, such
+as the latency of memory regulation are the one of the aspects of the memory management.
+The latency is the delay or the time that it takes the data to be read from memory.
+Since the movement of this data is bound to be within the speed of light,
+the medium it is travelling in, this means that even though that speed is above human physical
+comprehension but still there would be a delay, based on the size of the data.
+
 We gathered the latency time, which is the time delay taking place in data transfer,reading and
-writing at the lowest level of memory management.
+writing at the lowest level of memory management. This was the primary statistic for us to consider
+when building the simulator.
 
@@
 central processing unit (CPU), the Hard Drive and the RAM sticks.
-Throughout the simulation the CPU was accessing the RAM and the Hard Drive to process
-the data existing. We timed the transfer of data from both the Hard Drive and RAM.
+
+The CPU further consists of cache, modern day powerful CPUS consisted of 2 level of caches, L1 and L2,
+where L1 is very small compared to L2 in memory. L1 ranging from 2KB to 64KB, whereas the second level
+is in the range of 256 KB to 2 MB. Cache is used for quicker access by the CPU of the data that it
+frequently uses. Cache stores that data and thus decreasing the latency. However, in the simulator we
+implemented a single level of cache with a variable size, which is input by the user.
+
+Further down the heirarchy of memory managemnt, we have the RAM, which is relatively slower to CPU but bigger
+in size by quite some margin. The RAM, which translates to RAndom Access Memory is a volatile form of memory
+built by several memory cells. There are two types of RAMs; DRAM and SRAM, eventhought we did not distinguish
+between them in our simulator. Given that the RAM is slower compared to the CPU(cache) it takes time to read
+and move data off it. In the simulator we kept the size and the number of RAM sticks a variable, so
+the user inputs the value for each. Since the RAM is voaltile it loses all the data it has on it, once we
+take the power off it. As soon as the power is turned on, the alot of important section of OS is read from
+the hard drive and put onto the RAM so some important functions can be accessed quickly instead of reading
+it off the hard drive which is way slower. This OS data along with other make up alot the RAM usage when the
+computer is running. Running other process lead to more writing and reading off the RAM which leads to different
+delays.
+
+Finally comes the Hard Drive, which is the biggest form of permanent memory but the slowest one. It is
+extremely large compared to both RAM and cache. Almost all the software including the OS is stored on
+the HDD. HDDs have evolved in many ways over the course of times, making them quicker, but they have
+remained slower compared to RAM and CPU always. In our simulator we have built an HDD on arrays of bytes,
+just like the RAM and cache.
 
@@
 was represented by arrays of byte type. As the simulation went on, these arrays in each of the components
-were modified depending upon the event tanking place. This modification gave us the data for how the size
+were modified depending upon the event taking place. This modification gave us the data for how the size
 of data acts proportional to the latency and the delay in the transfer, writing and reading of data. The size
@@
 
-As the variable we had in the simulation were the number of Cores in the CPU,
+As the variables we had in the simulation were the number of Cores in the CPU,
 the number of hard drives, the size of hard drive in bytes, the number of RAM sticks, the size of
@@
 
-Also describe components and terms as you go.
-
 \section{Data Gathered}
@@
 Insert all the data gathered by the simulator and from sources here.
+
 Explain any terms as needed (only if not already).

### 2020-11-20 platter.java
Removed unneeded files

@@ -1,16 +1,0 @@
-/*
-  A simulation of a single sided HDD platter
- */
-class platter{
-    //An array to hold our data
-    byte[][] data;
-    
-    /**
-       Constructor for a platter
-     **/
-    public platter(int size){
-	//This will have to be changed, as we have a square array, but actually, more sectors
-	//are present farther from the center of the platter
-	data = new byte[size][size];
-    }
-}

### 2020-11-20 memSim.java
Fixed spacing issue on some user prompt lines

@@ -245,7 +245,7 @@ class memSim extends Thread{
 
 		//Now we have to prompt the user to pick a chunk of data
 		hardDrives[hddSelect].print();
-		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
+		System.out.print("\u001B[33mPlease select one of the above partitions: \u001B[0m");
 		dataSelect1 = scnr.nextInt();
 
             }
@@ -257,7 +257,7 @@ class memSim extends Thread{
 		
 		//Now we have to prompt the user to pick a chunk of data
 		rams[ramSelect].print();
-		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
+		System.out.print("\u001B[33mPlease select one of the above partitions: \u001B[0m");
 		dataSelect1 = scnr.nextInt();
 
             }
@@ -283,7 +283,7 @@ class memSim extends Thread{
 
 		//Now we have to prompt the user to pick a chunk of data
 		hardDrives[hddSelect].print();
-		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
+		System.out.print("\u001B[33mPlease select one of the above partitions: \u001B[0m");
 		dataSelect2 = scnr.nextInt();
 
             }
@@ -295,7 +295,7 @@ class memSim extends Thread{
 		
 		//Now we have to prompt the user to pick a chunk of data
 		rams[ramSelect].print();
-		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
+		System.out.print("\u001B[33mPlease select one of the above partitions: \u001B[0m");
 		dataSelect2 = scnr.nextInt();
 
             }
@@ -338,7 +338,7 @@ class memSim extends Thread{
 
 		//Now we have to prompt the user to pick a chunk of data
 		hardDrives[hddSelect].print();
-		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
+		System.out.print("\u001B[33mPlease select one of the above partitions: \u001B[0m");
 		int dataSelect = scnr.nextInt();
 
 		//Then do the action
@@ -363,7 +363,7 @@ class memSim extends Thread{
 
 		//Now we have to prompt the user to pick a chunk of data
 		rams[ramSelect].print();
-		System.out.print("\u001B[33mPlease select one of the above partitions\u001B[0m");
+		System.out.print("\u001B[33mPlease select one of the above partitions: \u001B[0m");
[…]

### 2020-11-20 HDD.java
Added sleep to read method

@@ -111,9 +111,12 @@ class HDD extends Thread{
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
-            while(index<data.length && (data[index]!=null || data[index].start==true)){
+            while(index<data.length && (data[index]!=null || !data[index].start==true)){
                 index++;   
-
+		try{
+		    sleep(1);
+		}
+		catch(Exception e){}
             }
         }
     }

### 2020-11-20 HDD.java
Fixed read taking 0 seconds, and sometimes throwing an error

@@ -104,21 +104,31 @@ class HDD extends Thread{
     Reads data from this
     Precondition: index is a dataBlock where .start is true.
     @param index The index of the array where the data block starts
+    @return 1, when the program has finished reading
      **/
-    public void read(int index){
+    public int read(int index){
 
         //Loop through the data until we hit a null block or
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
-            while(index<data.length && (data[index]!=null || !data[index].start==true)){
-                index++;   
-		try{
-		    sleep(1);
-		}
-		catch(Exception e){}
+            while(index<data.length && data[index]!=null){
+                index++;  
+                
+                try{
+                    sleep(1);
+                }
+                catch(Exception e){
+                    
+                }
+                
+                //Now we check the next position if it is a starter, and we can exit
+                if(data[index]!=null && data[index].start==true){
+                    break;
+                }
             }
         }
+        return 1;
     }
 
     /**

### 2020-11-20 RAM.java

@@ -75,14 +75,30 @@ class RAM extends Thread{
     Reads data from this, and returns it
     Precondition the index is a datablock with .start = true
     @param the index of the starting array index
+    @return 1, when the program has finished reading
      **/
-    public void read(int index){
+    public int read(int index){
         //Loop through the data until we hit a null block or
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
-        while(data[index]!=null || data[index].start==true){
-            index++;
+        if(data!=null){
+            while(index<data.length && data[index]!=null){
+                index++;  
+                
+                try{
+                    sleep(1);
+                }
+                catch(Exception e){
+                    
+                }
+                
+                //Now we check the next position if it is a starter, and we can exit
+                if(data[index]!=null && data[index].start==true){
+                    break;
+                }
+            }
         }
+        return 1;
     }
 
     /**

### 2020-11-20 memSim.java

@@ -344,7 +344,8 @@ class memSim extends Thread{
 		//Then do the action
 		start = System.currentTimeMillis();
 
-                hardDrives[hddSelect].read(dataSelect);
+		//We don't need this return value for anything, we just need memSim to wait
+                int x = hardDrives[hddSelect].read(dataSelect);
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
@@ -367,7 +368,7 @@ class memSim extends Thread{
 		int dataSelect = scnr.nextInt();
 
 		//Then do the action
-                rams[ramSelect].read(dataSelect);
+                int x = rams[ramSelect].read(dataSelect);
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();
@@ -380,7 +381,8 @@ class memSim extends Thread{
             else if(userSubSelect == 3){
                 start = System.currentTimeMillis();
 
-                cpu.read();
+                //We don't need this return value for anything, we just need memSim to wait
+                byte[] x = cpu.read();
 
                 //Then we can report the statistics
                 end = System.currentTimeMillis();

### 2020-11-20 memSim.java
Added delete to help message

@@ -568,6 +568,7 @@ class memSim extends Thread{
         System.out.println("move");
         System.out.println("read");
         System.out.println("write");
+	System.out.println("delete");
         System.out.println("help");
         System.out.println("exit");
         System.out.println("For help with a specific command type [command]");
@@ -590,7 +591,7 @@ class memSim extends Thread{
 		System.out.println("Writes data to specified location");
 	    }
 	    else if(input.equalsIgnoreCase("delete")){
-		System.out.println("Deletes data from the specified drive");
+		System.out.println("Deletes data from the specified component");
 	    }
 	    else if(input.equalsIgnoreCase("help")){
 		System.out.println("Displays the help message");

### 2020-11-20 HDD.java
Fixed printer not showing correct number of bytes

@@ -116,7 +116,7 @@ class HDD extends Thread{
                 index++;  
                 
                 try{
-                    sleep(1);
+                    sleep(2);
                 }
                 catch(Exception e){
                     
@@ -163,7 +163,7 @@ class HDD extends Thread{
             for(int i=0;i<data.length;i++){
                 //Then if we encounter a dataBlock that is a start, print it
                 if(data[i]!=null && data[i].start == true){
-                    System.out.println(i+": "+data[i].size+" bytes big");
+                    System.out.println(i+": "+data[i].size*dataBlock.WIDTH+" bytes big");
                 }
             }
         }

### 2020-11-20 RAM.java

@@ -87,10 +87,10 @@ class RAM extends Thread{
                 
                 try{
                     sleep(1);
-                }
+		}
                 catch(Exception e){
                     
-                }
+		}
                 
                 //Now we check the next position if it is a starter, and we can exit
                 if(data[index]!=null && data[index].start==true){
@@ -135,7 +135,7 @@ class RAM extends Thread{
             for(int i=0;i<data.length;i++){
                 //Then if we encounter a dataBlock that is a start, print it
                 if(data[i]!=null && data[i].start == true){
-                    System.out.println(i+": "+data[i].size+" bytes big");
+                    System.out.println(i+": "+data[i].size*dataBlock.WIDTH+" bytes big");
                 }
             }
         }

### 2020-11-20 memSim.java
Added ANSI code check, and vars to hold these, so the user can use all white text, and colored text

@@ -23,9 +23,19 @@ class memSim extends Thread{
     //An array to hold all the ram sticks in the system
     RAM[] rams;
 
-    ArrayList processes = new ArrayList<process>();
+    //Our cpu
     CPU cpu;
+    
+    //Some colored text tags
+    private static String reset;
+    private static String red;
+    private static String green;
+    private static String yellow;
+    private static String blue;
+    private static String purple;
+    private static String cyan;
 
+    
     /**
     Starts the simulation of our memory management.
     First prompts the user for data
@@ -65,7 +75,7 @@ class memSim extends Thread{
 
         //Now let's initialize our cacheSize
         cpu = new CPU(1, cacheSize);
-        cpu.start();
+       cpu.start();
     }
 
     /**
@@ -77,36 +87,66 @@ class memSim extends Thread{
     private static memSim initialUserPrompt(){
         Scanner scnr = new Scanner(System.in);
 
-	//Let's first ask the user if they want to make their own system, or load one
-	System.out.println("Would you like to load a system, or configure your own?");
-	System.out.println("Type \u001B[33mLOAD\u001B[0m or \u001B[33mNEW\u001B[0m: ");
+	//First we need to ask the user if their terminal is ANSI compatible
+	System.out.println("\u001B[34mIs this text blue?\u001B[0m Y/n");
 	String choice = scnr.nextLine();
+
+	if(choice.equalsIgnoreCase("y")){
+	    //Since they can see the blue text, we can use our ANSI codes
+
+	    reset = "\u001B[0m";
+	    red = "\u001B[31m";
+	    green = "\u001B[32m";
+	    yellow = "\u001B[33m";
+	    blue = "\u001B[34m";
+	    purple = "\u001B[35m";
+	    cyan = "\u001B[36m";
+	}
+	else{
+	    //Otherwise they can't see the blue text, or we didn't understand
+	    if(!choice.equalsIgnoreCase("n")){
+		System.out.println("We didn't understand your answer. Reverting to plain text");
+	    }
+
+	    reset = "";
+	    red = "";
+	    green = "";
+	    yellow = "";
+	    blue = "";
+	    purple = "";
+	    cyan = "";
+	}
+
+	//Let's ask the user if they want to make their own system, or load one
+	System.out.println("Would you like to load a system, or configure your own?");
+	System.out.println("Type "+yellow+"LOAD"+reset+" or "+yellow+"NEW"+reset+": ");
+	choice = scnr.nextLine();
 	if(choice.equalsIgnoreCase("Load")){
[…]

### 2020-11-20 HDD.java
Hotfix for not writing to HDD correctly

@@ -43,7 +43,7 @@ class HDD extends Thread{
         if(numBytes > spaceLeft){return 0;}
 
         //Now let's reserve our space and then write our data
-        spaceLeft =- numBytes;
+        spaceLeft = spaceLeft - numBytes;
 
         //Now we call our helper method
         int totalBlocks;
@@ -89,10 +89,8 @@ class HDD extends Thread{
                     return 1;
                 }
                 //If we got here, that means we don't have a hole big enough, so let's reset count
-                //and skip those value by adding (count - 1) + i
-                count -= 1;
-                i += count;
-                count = 0;
+                
+                count = i;
 
             }
         }

### 2020-11-20 RAM.java
Updated RAM to use same HDD hotfixes

@@ -28,7 +28,7 @@ class RAM extends Thread{
         if(input > spaceLeft){return 0;}
 
         //Now let's reserve that space, and write our data
-        spaceLeft =- input;
+        spaceLeft = spaceLeft - input;
 
         //Now we call our helper method
         return add(input/dataBlock.WIDTH);
@@ -62,9 +62,6 @@ class RAM extends Thread{
                     return 1;
                 }
                 //If we got here, that means we don't have a hole big enough, so let's reset count
-                //and skip those values by adding (count - 1) and i
-                count -= 1;
-                i += count;
                 count = 0;
             }
         }

### 2020-11-20 CPU.java
Converted to use a int param instead of byte[]

@@ -31,10 +31,10 @@ class CPU extends Thread{
        there should be no delay as this is stored on RAM, and the cache
        in a real system will be much faster
 
-       @return the data read in a byte array
+       @return the size of the cache read
      **/
-    public byte[] read(){
-	return cache1;
+    public int read(){
+	return cache1Size;
     }
     /**
        Writes input into the cache.
@@ -44,12 +44,12 @@ class CPU extends Thread{
        There will also not be any delay (other than overhead) as these are the
        fastest memory locations.
 
-       @param input the data to be written to the cache
+       @param numBytes, the number of bytes to be written to the cache
        @return 1 if the method is done, 0 if error occurs
      **/
-    public int write(byte[] input){
-	for(int i = 0; i<input.length;i++){
-	    add(input[i]);
+    public int write(int numBytes){
+	for(int i = 0; i<numBytes;i++){
+	    add((byte)0);
 	}
 	return 1;
     }

### 2020-11-20 HDD.java
Read() now returns the number of bytes read

@@ -102,7 +102,7 @@ class HDD extends Thread{
     Reads data from this
     Precondition: index is a dataBlock where .start is true.
     @param index The index of the array where the data block starts
-    @return 1, when the program has finished reading
+    @return the size of the chunk read, otherwise -1
      **/
     public int read(int index){
 
@@ -110,6 +110,8 @@ class HDD extends Thread{
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
+	    if(data[index]!=null){
+		int output = data[index].size;
             while(index<data.length && data[index]!=null){
                 index++;  
                 
@@ -125,8 +127,10 @@ class HDD extends Thread{
                     break;
                 }
             }
+	    return output;
+	    }
         }
-        return 1;
+        return -1;
     }
 
     /**

### 2020-11-20 RAM.java

@@ -72,13 +72,15 @@ class RAM extends Thread{
     Reads data from this, and returns it
     Precondition the index is a datablock with .start = true
     @param the index of the starting array index
-    @return 1, when the program has finished reading
+    @return the size of the chunk read, or -1 if error
      **/
     public int read(int index){
         //Loop through the data until we hit a null block or
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
+	    if(data[index]!=null){
+	    int output = data[index].size;
             while(index<data.length && data[index]!=null){
                 index++;  
                 
@@ -94,8 +96,10 @@ class RAM extends Thread{
                     break;
                 }
             }
+	    return output;
+	    }
         }
-        return 1;
+        return -1;
     }
 
     /**

### 2020-11-20 memSim.java
Now complies with new updates

@@ -272,19 +272,24 @@ class memSim extends Thread{
             System.out.println("3.\tCPU Cache");
             System.out.print(yellow+"Please enter the number of your selection: "+reset);
 
+	    //userSubSelects tell us what component to access
             int userSubSelect1 = scnr.nextInt();
 	    int userSubSelect2 = userSubSelect1;
+	    //subSelects tell us which number of that component type
+	    int subSelect1 = -1;
+	    int subSelect2 = -1;
+	    //dataSelects tell us what data to move on the drive
 	    int dataSelect1 = -1;
 	    int dataSelect2 = -1;
-
+	    
 	    if(userSubSelect1 == 1){
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
                 System.out.print(yellow+"Please enter the number of your selection: "+reset);
-                int hddSelect = scnr.nextInt();
-		hddSelect--;
+                subSelect1 = scnr.nextInt();
+		subSelect1--;
 
 		//Now we have to prompt the user to pick a chunk of data
-		hardDrives[hddSelect].print();
+		hardDrives[subSelect1].print();
 		System.out.print(yellow+"Please select one of the above partitions: "+reset);
 		dataSelect1 = scnr.nextInt();
 
@@ -292,17 +297,18 @@ class memSim extends Thread{
             else if(userSubSelect1 == 2){
                 System.out.println("Please choose a stick of ram from 1 - " + rams.length);
                 System.out.print(yellow+"Please enter the number of your selection: "+reset);
-                int ramSelect = scnr.nextInt();
-		ramSelect--;
+                subSelect1 = scnr.nextInt();
+		subSelect1--;
 		
 		//Now we have to prompt the user to pick a chunk of data
-		rams[ramSelect].print();
+		rams[subSelect1].print();
 		System.out.print(yellow+"Please select one of the above partitions: "+reset);
 		dataSelect1 = scnr.nextInt();
 
             }
             else if(userSubSelect1 == 3){
 		dataSelect1 = 1;
+		
             }
 	    
 	    //Now we prompt where the user wants to put this data
@@ -318,11 +324,11 @@ class memSim extends Thread{
 	    if(userSubSelect2 == 1){
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
                 System.out.print(yellow+"Please enter the number of your selection: "+reset);
-                int hddSelect = scnr.nextInt();
[…]

### 2020-11-21 HDD.java
Implemented hotfix so move is now working correctly

@@ -89,7 +89,7 @@ class HDD extends Thread{
                     return 1;
                 }
                 //If we got here, that means we don't have a hole big enough, so let's reset count
-                
+
                 count = i;
 
             }
@@ -110,25 +110,25 @@ class HDD extends Thread{
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
-	    if(data[index]!=null){
-		int output = data[index].size;
-            while(index<data.length && data[index]!=null){
-                index++;  
-                
-                try{
-                    sleep(2);
+            if(data[index]!=null){
+                int output = data[index].size;
+                while(index<data.length && data[index]!=null){
+                    index++;  
+
+                    try{
+                        sleep(2);
+                    }
+                    catch(Exception e){
+
+                    }
+
+                    //Now we check the next position if it is a starter, and we can exit
+                    if(data[index]!=null && data[index].start==true){
+                        break;
+                    }
                 }
-                catch(Exception e){
-                    
-                }
-                
-                //Now we check the next position if it is a starter, and we can exit
-                if(data[index]!=null && data[index].start==true){
-                    break;
-                }
+                return output;
             }
-	    return output;
-	    }
         }
         return -1;
     }
@@ -142,13 +142,15 @@ class HDD extends Thread{
         //Loop through the data until we hit a null block or
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
-        while(index<data.length && (data[index]!=null || data[index].start==true)){
-            //Delete the current block
-            data[index] = null;
+        if(data!=null){
+            while(index<data.length && data[index]!=null){
+                //Delete the current block
+                data[index] = null;
 
-            //Then increment and free space
-            index++;
-            spaceLeft+=dataBlock.WIDTH;
+                //Then increment and free space
+                index++;
+                spaceLeft+=dataBlock.WIDTH;
+            }
         }
     }
 

### 2020-11-21 RAM.java

@@ -112,7 +112,7 @@ class RAM extends Thread{
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
-            while(data[index]!=null || data[index].start==true){
+            while(index<data.length && data[index]!=null){
                 //Delete the current block
                 data[index]=null;
 

### 2020-11-21 memSim.java

@@ -280,7 +280,6 @@ class memSim extends Thread{
 	    int subSelect2 = -1;
 	    //dataSelects tell us what data to move on the drive
 	    int dataSelect1 = -1;
-	    int dataSelect2 = -1;
 	    
 	    if(userSubSelect1 == 1){
                 System.out.println("Please choose a drive from 1 - " + hardDrives.length);
@@ -327,30 +326,20 @@ class memSim extends Thread{
                 subSelect2 = scnr.nextInt();
 		subSelect2--;
 
-		//Now we have to prompt the user to pick a chunk of data
-		hardDrives[subSelect2].print();
-		System.out.print(yellow+"Please select one of the above partitions: "+reset);
-		dataSelect2 = scnr.nextInt();
-
             }
             else if(userSubSelect2 == 2){
                 System.out.println("Please choose a stick of ram from 1 - " + rams.length);
                 System.out.print(yellow+"Please enter the number of your selection: "+reset);
                 subSelect2= scnr.nextInt();
 		subSelect2--;
-		
-		//Now we have to prompt the user to pick a chunk of data
-		rams[subSelect2].print();
-		System.out.print(yellow+"Please select one of the above partitions: "+reset);
-		dataSelect2 = scnr.nextInt();
 
             }
             else if(userSubSelect2 == 3){
-		dataSelect2 = 1;
+		
             }
 	    
 	    //We make sure that the user didn't try to move from A to A
-	    if(userSubSelect1==userSubSelect2 && dataSelect1==dataSelect2){
+	    if(userSubSelect1==userSubSelect2){
 		System.out.println(red+"Error, you cannot move data from one place to the same place. Please try again"+reset);
 	    }
 	    }

### 2020-11-21 finalPaper/paper.tex
Added more to the running section, and some spell checking done

@@
 we present the actual data collected from the simulator, as well as actual 
-real world data. Section~\ref{sec:disc} goes on to explain any discrepencies 
+real world data. Section~\ref{sec:disc} goes on to explain any discrepancies 
 found between our data and the real world collected data. Finally, in~\ref{sec:conclusions}
@@
 from moving this data around the artificial system. The simulator consists of four
-primary parts: memSim, which is essentially the contoller; a CPU with one level of cache;
+primary parts: memSim, which is essentially the controller; a CPU with one level of cache;
 one or more sticks or RAM; and one or more hard drive disks.
@@
 
-\subsection{Aquiring Source Code}
+\subsection{Acquiring Source Code}
 To download the simulator code, go to \url{osfinal.dachenhausen.org} 
@@
 simulator parameters. Each of these is crucial, and cannot be left blank.
+
+The first thing the simulator will ask you is
+{\singlespace
+\begin{verbatim}
+Is this text blue? Y/n
+\end{verbatim}
+If your terminal supports ANSI colored text, then the question should be blue,
+and you should respond Y. This helps point out important data.
+If your terminal does not support the colored text, then you will see something like
+{\singlespace
+\begin{verbatim}
+[34mIs this text blue? [0m Y/n
+\end{verbatim}
+in which case, if you respond n, the simulator will not use any ANSI text and
+everything will print normally.
+
+The simulator will then ask you whether you would like to LOAD or create a NEW system.
+
+\subsection*{Load An Existing System}
+The simulator will then show you three system specifications and a prompt, in which
+you should choose a system by entering its system number.
+
+\subsection*{Create A New System}
+If you opt to create a new system, you will be presented with the prompt shown below,
+which has sample data included. Note: very large numbers will make the simulator lag
+the Java virtual machine, so please be careful when inputting this data.
 {\singlespace
@@
 1
-How many platters should each hard drive have?
-4
 And how big (in bytes) should each one be?
@@
 write
+delete
 help
 exit
-For help with a specific command type 'help [command]'
+For help with a specific command type [command]
+Type 'q' to quit the help menu
 \end{verbatim}
@@
 
+In all actions, the simulator starts its timing only when user input is done,
+so your data will not reflect your typing speed.
+
+\subsection*{Reading Data}
+Reading data is fairly simple, upon selecting read by entering its number into
+the prompt, the simulator will show the following menu:
+{\singlespace
+\begin{verbatim}
+Where would you like to read the data from?
+1.      Hard Drive
+2.      RAM
+3.      CPU Cache
+Please enter the number of your selection:
+\end{verbatim}
+}
+You should then enter the number corresponding to your selection.
+If you select a hard drive or a RAM stick, you will first get shown:
+{\singlespace
+\begin{verbatim}
+Please choose a drive from 1 - 2
+\end{verbatim}
+}
+Then, you will get shown that drive/stick's specific contents:
+{\singlespace
+\begin{verbatim}
+****Data on this hard drive****
+0: 1024 bytes big
+2: 1024 bytes big
+3: 1024 bytes big
+*******************************
+\end{verbatim}
+}
+You should then select the data that you wish to be read.
+
+If you select the CPU cache, the entire contents of the cache will be read.
+
+\subsection*{Writing Data}
+Writing data is similar to reading data, as you will still be shown a prompt
+of where you would like to write, and which specific drive/stick's contents to write to.
+You will then be shown:
+{\singlespace
+\begin{verbatim}
+How many bytes would you like to write?
+\end{verbatim}
+}
+In which, you should enter the number of bytes to write. Note, if you enter a number greater
+than the space remaining on the drive/stick, then the write will fail.
+
+If you select the CPU cache, there is no limiter (other than that of Java) on maximum data
+that can be written. The data structure is set up so that writing to cache is circular, so
+the last n bytes are the ones stored in the cache, where n is the size of the cache.
+
+\subsection*{Deleting Data}
+Deleting data is very similar to reading data, as they are essentially the same operation.
+In this case, every iteration through the stored data, the current index is set to null.
+
+\subsection*{Moving Data}
+A move is a combination of the three aforementioned commands. First, the simulator will prompt
+like a read from where to read, and then delete the data from. Then the simulator will prompt
+for where this data should be written. At this point the simulator will also make sure that
+you do not try to read from one location, then write to that exact same location, ie. read
+from drive 1, chunk 0, then try to write to drive 1, chunk 0. The simulator will re-prompt until
+this is not the case. 
 \section{Explanation of Data Gathered}
@@
 Since the movement of this data is bound to be within the speed of light,
-the medium it is travelling in, this means that even though that speed is above human physical
+the medium it is traveling in, this means that even though that speed is above human physical
 comprehension but still there would be a delay, based on the size of the data.
@@
 
-To reperesent this level of memory management the simulator was designed to hold some key components.
+To represent this level of memory management the simulator was designed to hold some key components.
 These components are those linked directly with the memory usage,including the
@@
 
-Further down the heirarchy of memory managemnt, we have the RAM, which is relatively slower to CPU but bigger
-in size by quite some margin. The RAM, which translates to RAndom Access Memory is a volatile form of memory
-built by several memory cells. There are two types of RAMs; DRAM and SRAM, eventhought we did not distinguish
+Further down the hierarchy of memory management, we have the RAM, which is relatively slower to CPU but bigger
+in size by quite some margin. The RAM, which translates to Random Access Memory is a volatile form of memory
+built by several memory cells. There are two types of RAMs; DRAM and SRAM, even thought we did not distinguish
 between them in our simulator. Given that the RAM is slower compared to the CPU(cache) it takes time to read
 and move data off it. In the simulator we kept the size and the number of RAM sticks a variable, so
-the user inputs the value for each. Since the RAM is voaltile it loses all the data it has on it, once we
+the user inputs the value for each. Since the RAM is volatile it loses all the data it has on it, once we
 take the power off it. As soon as the power is turned on, the alot of important section of OS is read from
@@
 each of the RAM stick and the size of the CPU Cache, we kept a record of each combination. However,
-even with the CPU having a variable number of cores, for the path of study being conducted, we mantained
+even with the CPU having a variable number of cores, for the path of study being conducted, we maintained
 it to be 1 throughout the study. The simulator was structured to take in a value for each of the variables,
@@
 and transfer on RAM vary in the aspect of Time as with different sizes, and how slower or faster was the case
-with the accessing, editing and transfering from and to the Hard drives.
+with the accessing, editing and transferring from and to the Hard drives.
 
@@
 
-\section{Discrepencies}
+\section{Discrepancies}
 \label{sec:disc}
 
-Explain any discrepencies found here.
+Explain any discrepancies found here.
 

### 2020-11-21 finalPaper/paper.tex
Abstract started

@@
 
-Insert abstract here.
+The Java Based Memory Latency Simulator was developed to provide a way to
+study how long it takes to move bytes of data around a common computer
+system containing hard drive disks, random access memory, and a central
+processing unit with cache memory. This project aims first to create a modular
+system, that could be used to simulate data movement computer components, then
+second to use that simulator and upgrade it to allow data to be collected about
+data latency. We first focused on creating data structures that represented various
+components of the computer, like a hard drive disk. Then we created a dual purpose
+structure that both holds these components and interacts with the user. For this project,
+we gathered data about the memory latency as an example of the modularity of the simulator.
 
@@
 The goal of this project was to create a computer memory latency simulator. 
-For its ease of use, Java was chosen to program this project in. This 
-simulator is designed to simulate data transfer from one system component 
+Java was chosen as the programming language for this project for its ease of use and
+it is what we, the authors are most comfortable with.
+This simulator is designed to simulate data transfer from one system component 
 to another. For example, from the hard drive disk, to the main memory.

### 2020-11-21 finalPaper/paper.tex
Updated abstract def. Closes #13 Closes #14

@@
 The goal of this project was to create a Java based simulator that will allow
-semi-low access to artificial data so the user could gather statistics and trends
-from moving this data around the artificial system. The simulator consists of four
+access to artificial data so the user could gather statistics and trends
+from moving this data around the artificial system. The simulator allows simple operations
+on blocks of data, not individual bytes. The simulator consists of four
 primary parts: memSim, which is essentially the controller; a CPU with one level of cache;
@@
 
-Each component's (CPU, RAM, HDD), storage is represented with a byte array.
-The user then has options to handle data on any component via the memSim terminal interface.
+Each component's (CPU, RAM, HDD), storage is represented with a simple data structure called
+a dataBlock, which represents 1024 bytes. The user then has options to handle data on any
+component via the memSim terminal interface.
 

### 2020-11-21 memSim.java
Added delay system

@@ -17,6 +17,9 @@ class memSim extends Thread{
     //A flag to tell our components if the system is done running
     public static boolean done;
 
+    //The value of one megabyte
+    public static final int MB = 1000000;
+
     //An array to hold all the hard drives in the system
     HDD[] hardDrives;
 
@@ -53,7 +56,7 @@ class memSim extends Thread{
     /**
     A constructor for the memSim object
      **/
-    public memSim(int numHDD, int hDDSize, int numRam, int rAMSize, int cacheSize){
+    public memSim(int numHDD, int hDDSize, int numRam, int rAMSize, int cacheSize, int cacheDelay, int ramDelay, int hddDelay){
         //First, we need to allow everything to know that we are starting up, aka not done
         done = false;
 
@@ -61,7 +64,7 @@ class memSim extends Thread{
         hardDrives = new HDD[numHDD];
 
         for(int i = 0; i < numHDD; i++){
-            hardDrives[i] = new HDD(hDDSize);
+            hardDrives[i] = new HDD(hDDSize, hddDelay);
             hardDrives[i].start();
         }
 
@@ -69,12 +72,12 @@ class memSim extends Thread{
         rams = new RAM[numRam];
 
         for(int i = 0; i < numRam; i++){
-            rams[i] = new RAM(rAMSize);
+            rams[i] = new RAM(rAMSize, ramDelay);
             rams[i].start();
         }
 
         //Now let's initialize our cacheSize
-        cpu = new CPU(1, cacheSize);
+        cpu = new CPU(1, cacheSize, cacheDelay);
        cpu.start();
     }
 
@@ -145,12 +148,20 @@ class memSim extends Thread{
 	    System.out.println("\t1 Stick of RAM with size: 51200 bytes");
 	    System.out.println("\tCPU Cache size: 2048 bytes");
 	    System.out.println("\tThere is no data written to any of the components in this system.");
+
+	    //Now we let the user know about how the delays are set up
+	    System.out.println("**** Delays ****");
+	    System.out.println("All systems are configured with the following");
+	    System.out.println("CPU cache:\tno delay");
+	    System.out.println("RAM:\tTo read/write 1 MB = 5 secs");
+	    System.out.println("HDD:\tTo read/write 1 MB = 10 secs");
+	    System.out.println("****************");
[…]

### 2020-11-21 CPU.java
Added delay system

@@ -14,15 +14,20 @@ class CPU extends Thread{
     
     //THe size of the first cache
     int cache1Size;
+
+    //The time it takes to write 1 MB
+    int delay;
     /**
        Constructor for the CPU class
        @param numCores the number of cores that this should have (currently not supported)
        @param cacheSize the size of the cache of this
+       @param cacheDelay, the time it takes for this to write 1 MB
      **/
-    public CPU(int numCores, int cacheSize){
+    public CPU(int numCores, int cacheSize, int cacheDelay){
      cores = numCores;
      cache1 = new byte[cacheSize];
      cache1Size = cacheSize;
+     delay = cacheDelay;
      cache1Last = -1; //We hold the index of the last item in the array
     }
     /**
@@ -34,6 +39,10 @@ class CPU extends Thread{
        @return the size of the cache read
      **/
     public int read(){
+	try{
+	    sleep(delay*(cache1Size/memSim.MB));
+	}
+	catch(Exception e){}
 	return cache1Size;
     }
     /**
@@ -48,6 +57,10 @@ class CPU extends Thread{
        @return 1 if the method is done, 0 if error occurs
      **/
     public int write(int numBytes){
+	try{
+	    sleep(delay*(cache1Size/memSim.MB));
+	}
+	catch(Exception e){}
 	for(int i = 0; i<numBytes;i++){
 	    add((byte)0);
 	}
@@ -77,6 +90,10 @@ class CPU extends Thread{
 	for(int i=0; i<numBytes;i++){
 	    cache1Last++;
 	}
+	try{
+	    sleep(delay*(cache1Size/memSim.MB));
+	}
+	catch(Exception e){}
     }
     /**
 

### 2020-11-21 HDD.java

@@ -12,14 +12,18 @@ class HDD extends Thread{
     //Our array of data
     private dataBlock[] data;
 
+    //
+    int delay;
     /**
     Sets up a hard drive to be added to the system.
 
     @param size, the size in bytes of this
+    @param hddDelay, the time it takes handle 1 MB of data
      **/
-    public HDD(int size){
+    public HDD(int size, int hddDelay){
         this.size=size;
         this.spaceLeft=size;
+	delay = hddDelay;
         int totalBlocks;
         if(size == 0){totalBlocks=0;}
         else if(size<dataBlock.WIDTH){totalBlocks = 1;}
@@ -53,6 +57,9 @@ class HDD extends Thread{
             totalBlocks = numBytes/dataBlock.WIDTH;
             if(numBytes%dataBlock.WIDTH!=0){totalBlocks++;}
         }
+	try{
+	    sleep(delay*(numBytes/memSim.MB));}
+	catch(Exception e){}
         return add(totalBlocks);
     }
 
@@ -105,13 +112,15 @@ class HDD extends Thread{
     @return the size of the chunk read, otherwise -1
      **/
     public int read(int index){
-
         //Loop through the data until we hit a null block or
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
             if(data[index]!=null){
                 int output = data[index].size;
+		try{
+		    sleep(delay*(output/memSim.MB));}
+		catch(Exception e){}
                 while(index<data.length && data[index]!=null){
                     index++;  
 
@@ -143,6 +152,13 @@ class HDD extends Thread{
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
+	    if(data[index]!=null && data[index].start == true){
+		int numBytes = data[index].size;
+		try{
+		    sleep(delay*(numBytes/memSim.MB));}
+		catch(Exception e){}
+		
+	    }
             while(index<data.length && data[index]!=null){
                 //Delete the current block
                 data[index] = null;
@@ -150,6 +166,9 @@ class HDD extends Thread{
                 //Then increment and free space
                 index++;
                 spaceLeft+=dataBlock.WIDTH;
+		if(data[index]!=null && data[index].start == true){
+		    break;
+		}
             }
         }
     }

### 2020-11-21 RAM.java
Added delay system and constructor comment

@@ -11,9 +11,18 @@ class RAM extends Thread{
     //The array to hold our data
     dataBlock[] data;
 
-    public RAM(int size){
+    //The time it takes to write 1 MB
+    private int delay;
+
+    /**
+       Constructor for RAM object
+       @param size the size of this
+       @param ramDelay the time it take to write 1 MB
+     **/
+    public RAM(int size, int ramDelay){
         this.size = size;
         this.spaceLeft = size;
+	delay = ramDelay;
 
         data = new dataBlock[size/dataBlock.WIDTH];
     }
@@ -30,6 +39,11 @@ class RAM extends Thread{
         //Now let's reserve that space, and write our data
         spaceLeft = spaceLeft - input;
 
+	try{
+	    sleep(delay*(input/memSim.MB));
+	}
+	catch(Exception e){}
+	
         //Now we call our helper method
         return add(input/dataBlock.WIDTH);
 
@@ -80,29 +94,26 @@ class RAM extends Thread{
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
 	    if(data[index]!=null){
-	    int output = data[index].size;
-            while(index<data.length && data[index]!=null){
-                index++;  
-                
-                try{
-                    sleep(1);
+		int output = data[index].size;
+		try{
+		    sleep(delay*(output/memSim.MB));
 		}
-                catch(Exception e){
-                    
+		catch(Exception e){}
+		while(index<data.length && data[index]!=null){
+		    index++;  
+		    
+		    //Now we check the next position if it is a starter, and we can exit
+		    if(data[index]!=null && data[index].start==true){
+			break;
+		    }
 		}
-                
-                //Now we check the next position if it is a starter, and we can exit
-                if(data[index]!=null && data[index].start==true){
-                    break;
-                }
-            }
-	    return output;
+		return output;
 	    }
-        }
-        return -1;
+	}
+	return -1;
     }
-
-    /**
+	
+	/**
     Deletes data from this
     Precondition the index is a datablock with .start = true
     @param the index of the starting point to delete from
@@ -112,6 +123,12 @@ class RAM extends Thread{
         //If we have two blocks back-to-back, then the start flag
         //will be high, so we can stop reading our dataBlock
         if(data!=null){
+	    if(data[index]!=null){
+		try{
+		    sleep(delay*(data[index].size/memSim.MB));
+		  }
+		catch(Exception e){}
+	    }
             while(index<data.length && data[index]!=null){
                 //Delete the current block
                 data[index]=null;

### 2020-11-21 memSim.java
Added questions to initialPrompt

@@ -234,15 +234,15 @@ class memSim extends Thread{
 	    input = scnr.nextLine();
 	    int cacheSize = Integer.parseInt(input);
 
-	    System.out.println("");
+	    System.out.println("How long would you like it to take in seconds, to read/write to the cache?");
 	    input = scnr.nextLine();
 	    int cacheDelay = Integer.parseInt(input);
 
-	    System.out.println("");
+	    System.out.println("How long would you like it to take in seconds, to read/write to the RAM?");
 	    input = scnr.nextLine();
 	    int ramDelay = Integer.parseInt(input);
 
-	    System.out.println("Finally, ");
+	    System.out.println("Finally, how long would you like it to take in seconds, to read/write to a HDD?");
 	    input = scnr.nextLine();
 	    int hddDelay = Integer.parseInt(input);
 	    

### 2020-11-21 memSim.java
Added system config print lines when initializing a new system

@@ -252,6 +252,9 @@ class memSim extends Thread{
 	    System.out.println(numHDD+" number of hard drives, each "+hDDSize+" bytes big.");
 	    System.out.println(numRAM+" sticks of RAM, each "+rAMSize+" bytes big.");
 	    System.out.println("CPU cache is " + cacheSize + " bytes big.");
+	    System.out.println("Cache delay is:\t"+cacheDelay+"seconds");
+	    System.out.println("RAM delay is:\t"+ramDelay+"seconds");
+	    System.out.println("HDD delay is:\t"+hddDelay+"seconds");
 	    System.out.println(purple+"**********************************"+reset);
 	    System.out.println();
 

### 2020-11-21 finalPaper/paper.tex
Updated some verbatims

@@
 How many RAM sticks would you like?
-2
+1
 And how big (in bytes) should each one be?
-128
-Finally, how big (in bytes) would you like your CPU cache to be?
+512
+How big (in bytes) would you like your CPU cache to be?
 32
+How long would you like it to take in seconds, to read/write to the cache?
+0
+How long would you like it to take in seconds, to read/write to the RAM?
+5
+Finally, how long would you like it to take in seconds, to read/write to a HDD?
+10
+
+***********System Config**********
+1 number of hard drives, each 1024 bytes big.
+1 sticks of RAM, each 512 bytes big.
+CPU cache is 32 bytes big.
+Cache delay is: 0seconds
+RAM delay is:   5seconds
+HDD delay is:   10seconds
+**********************************
 \end{verbatim}
@@
 \begin{verbatim}
-Java Based Memory Latency Simulator
-Developed by Pat Baumgardner, Adam Dachenhausen, Shah Syed
-Supported commands:
-move
-read
-write
-delete
-help
-exit
-For help with a specific command type [command]
-Type 'q' to quit the help menu
+Please enter the number of your selection: 1
+What would you like to do next?
+1.      Move Data
+2.      Read Data
+3.      Write Data
+4.      Delete Data
+5.      Help
+6.      Exit
 \end{verbatim}

### 2020-11-21 finalPaper/paper.tex
Added to final paper

--- 2020-11-21 finalPaper/paper.tex
@@
 
-The Java Based Memory Latency Simulator was developed to provide a way to
-study how long it takes to move bytes of data around a common computer
-system containing hard drive disks, random access memory, and a central
-processing unit with cache memory. This project aims first to create a modular
-system, that could be used to simulate data movement computer components, then
-second to use that simulator and upgrade it to allow data to be collected about
-data latency. We first focused on creating data structures that represented various
-components of the computer, like a hard drive disk. Then we created a dual purpose
-structure that both holds these components and interacts with the user. For this project,
-we gathered data about the memory latency as an example of the modularity of the simulator.
+Insert abstract here.
 
@@
 The goal of this project was to create a computer memory latency simulator. 
-Java was chosen as the programming language for this project for its ease of use and
-it is what we, the authors are most comfortable with.
-This simulator is designed to simulate data transfer from one system component 
+For its ease of use, Java was chosen to program this project in. This 
+simulator is designed to simulate data transfer from one system component 
 to another. For example, from the hard drive disk, to the main memory.
@@
 we present the actual data collected from the simulator, as well as actual 
-real world data. Section~\ref{sec:disc} goes on to explain any discrepancies 
+real world data. Section~\ref{sec:disc} goes on to explain any discrepencies 
 found between our data and the real world collected data. Finally, in~\ref{sec:conclusions}
@@
 from moving this data around the artificial system. The simulator consists of four
-primary parts: memSim, which is essentially the controller; a CPU with one level of cache;
+primary parts: memSim, which is essentially the contoller; a CPU with one level of cache;
 one or more sticks or RAM; and one or more hard drive disks.
@@
 
-\subsection{Acquiring Source Code}
+\subsection{Aquiring Source Code}
 To download the simulator code, go to \url{osfinal.dachenhausen.org} 
@@
 simulator parameters. Each of these is crucial, and cannot be left blank.
-
-The first thing the simulator will ask you is
-{\singlespace
-\begin{verbatim}
-Is this text blue? Y/n
-\end{verbatim}
-If your terminal supports ANSI colored text, then the question should be blue,
-and you should respond Y. This helps point out important data.
-If your terminal does not support the colored text, then you will see something like
-{\singlespace
-\begin{verbatim}
-[34mIs this text blue? [0m Y/n
-\end{verbatim}
-in which case, if you respond n, the simulator will not use any ANSI text and
-everything will print normally.
-
-The simulator will then ask you whether you would like to LOAD or create a NEW system.
-
-\subsection*{Load An Existing System}
-The simulator will then show you three system specifications and a prompt, in which
-you should choose a system by entering its system number.
-
-\subsection*{Create A New System}
-If you opt to create a new system, you will be presented with the prompt shown below,
-which has sample data included. Note: very large numbers will make the simulator lag
-the Java virtual machine, so please be careful when inputting this data.
 {\singlespace
@@
 1
+How many platters should each hard drive have?
+4
 And how big (in bytes) should each one be?
@@
 write
-delete
 help
 exit
-For help with a specific command type [command]
-Type 'q' to quit the help menu
+For help with a specific command type 'help [command]'
 \end{verbatim}
@@
 
-In all actions, the simulator starts its timing only when user input is done,
-so your data will not reflect your typing speed.
-
-\subsection*{Reading Data}
-Reading data is fairly simple, upon selecting read by entering its number into
-the prompt, the simulator will show the following menu:
-{\singlespace
-\begin{verbatim}
-Where would you like to read the data from?
-1.      Hard Drive
-2.      RAM
-3.      CPU Cache
-Please enter the number of your selection:
-\end{verbatim}
-}
-You should then enter the number corresponding to your selection.
-If you select a hard drive or a RAM stick, you will first get shown:
-{\singlespace
-\begin{verbatim}
-Please choose a drive from 1 - 2
-\end{verbatim}
-}
-Then, you will get shown that drive/stick's specific contents:
-{\singlespace
-\begin{verbatim}
-****Data on this hard drive****
-0: 1024 bytes big
-2: 1024 bytes big
-3: 1024 bytes big
-*******************************
-\end{verbatim}
-}
-You should then select the data that you wish to be read.
-
-If you select the CPU cache, the entire contents of the cache will be read.
-
-\subsection*{Writing Data}
-Writing data is similar to reading data, as you will still be shown a prompt
-of where you would like to write, and which specific drive/stick's contents to write to.
-You will then be shown:
-{\singlespace
-\begin{verbatim}
-How many bytes would you like to write?
-\end{verbatim}
-}
-In which, you should enter the number of bytes to write. Note, if you enter a number greater
-than the space remaining on the drive/stick, then the write will fail.
-
-If you select the CPU cache, there is no limiter (other than that of Java) on maximum data
-that can be written. The data structure is set up so that writing to cache is circular, so
-the last n bytes are the ones stored in the cache, where n is the size of the cache.
-
-\subsection*{Deleting Data}
-Deleting data is very similar to reading data, as they are essentially the same operation.
-In this case, every iteration through the stored data, the current index is set to null.
-
-\subsection*{Moving Data}
-A move is a combination of the three aforementioned commands. First, the simulator will prompt
-like a read from where to read, and then delete the data from. Then the simulator will prompt
-for where this data should be written. At this point the simulator will also make sure that
-you do not try to read from one location, then write to that exact same location, ie. read
-from drive 1, chunk 0, then try to write to drive 1, chunk 0. The simulator will re-prompt until
-this is not the case. 
 \section{Explanation of Data Gathered}
@@
 Since the movement of this data is bound to be within the speed of light,
-the medium it is traveling in, this means that even though that speed is above human physical
+the medium it is travelling in, this means that even though that speed is above human physical
 comprehension but still there would be a delay, based on the size of the data.
@@
 
-To represent this level of memory management the simulator was designed to hold some key components.
+To reperesent this level of memory management the simulator was designed to hold some key components.
 These components are those linked directly with the memory usage,including the
@@
 
-Further down the hierarchy of memory management, we have the RAM, which is relatively slower to CPU but bigger
-in size by quite some margin. The RAM, which translates to Random Access Memory is a volatile form of memory
-built by several memory cells. There are two types of RAMs; DRAM and SRAM, even thought we did not distinguish
+Further down the heirarchy of memory managemnt, we have the RAM, which is relatively slower to CPU but bigger
+in size by quite some margin. The RAM, which translates to RAndom Access Memory is a volatile form of memory
+built by several memory cells. There are two types of RAMs; DRAM and SRAM, eventhought we did not distinguish
 between them in our simulator. Given that the RAM is slower compared to the CPU(cache) it takes time to read
 and move data off it. In the simulator we kept the size and the number of RAM sticks a variable, so
-the user inputs the value for each. Since the RAM is volatile it loses all the data it has on it, once we
+the user inputs the value for each. Since the RAM is voaltile it loses all the data it has on it, once we
 take the power off it. As soon as the power is turned on, the alot of important section of OS is read from
@@
 each of the RAM stick and the size of the CPU Cache, we kept a record of each combination. However,
-even with the CPU having a variable number of cores, for the path of study being conducted, we maintained
+even with the CPU having a variable number of cores, for the path of study being conducted, we mantained
 it to be 1 throughout the study. The simulator was structured to take in a value for each of the variables,
@@
 and transfer on RAM vary in the aspect of Time as with different sizes, and how slower or faster was the case
-with the accessing, editing and transferring from and to the Hard drives.
+with the accessing, editing and transfering from and to the Hard drives.
 
@@
 
+Latency Time in the all 3 major components being implemented in the simulator vary accordingly to their sizes,
+that becomes the reason why cache is the fastest between RAM and the HDD.
+
+Starting with the hardrive latency time for the input and output of data as it is the largest size of all memories here.
+The latency of a hard drive is a combination of it RPM, Seek Time and Transfer time. In the simulator that we
+implemented, we didnt go into details to a point that we offer the RPM, Seek time and the Transfer time as
+a variable to the user. However we implemented the HDD on data blocks instead of platters, which acted as
+partitions.
+According to Eric Shanks from ithollow.com, the following were the Hard Drive latencies in 2013, given the different
+RPM.
+
+\begin{tabular}{|c|c|}
+  \hline
+  RPMs & Rotational Latency(ms)\\
+  \hline
+  5400 & 11 ms\\
+  \hline
+  7200 & 8 ms\\
+  \hline
+  10000 & 6 ms\\
+  \hline
+  15000 & 4 ms\\
+  \hline
+\end {tabular}
+
+Seek time; which is the time it takes for the Read-Write to move up and down the platter.
+The Seek time itself is comprised of a series of phase times. The first being the acceleration phase,
+the read write arm starts to move, then comes the coasting phase when the arm is moving at full speed. This
+is followed by the deceleration phase where the read write arm slows down. Then finally comes the
+settling phase as the read write head stops right at the data it needs.
+The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
+(source: pages.cs.wisc.edu - Operating Systems: Three Easy peices peices - Remzi Arpaci, Andrea Arpaci).
+
+The last part of total I/O latency for hard drive is the transfer time, which is the time it takes the
+data to be transfer from or to the Read-Write head. Compared to Rotational Latency and seek time, transfer
+time is insignificant, given that the mechanical movements take more time than data movement which is more of
+movement of energy packets.
+
+After all this we come to an equation:
+
+I/O Latency Time of HDD = Rotational Latency + Seek Time + Latency Time
+
+In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the Operating Systems: Three
+easy peices, two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
+data being dealt with was 100 MB. However, another controller in the comparison was whether the data read or writing
+was sequential or random. Sequential meant that large chunks of data are just being read and written which are in
+sequence, whereas the Random transfer just meant Random data was selected from a read issue some where on the disk,
+which is how the data for applications work. Mind it that the data we represented in the simulator was also sequential
+and not random, as the hardrive implemented was built of data blocks.
+
+The I/O Latency time along with other statistics mentioned above was (taken from the source):
+
+\begin{tabular}{|l|r|r|}
+  \hline
+  HDD & Cheetah 15K.5 & Barracuda\\
+  \hline
+  Capacity & 300 GB & 1 TB\\
+  \hline
+  RPM & 15,000 & 7,200\\
+  \hline
+  Average Seek & 4 ms & 9ms\\
+  \hline
+  Platters & 4 & 4\\
+  \hline
+  T I/O Random & 6 ms & 13.2 ms\\
+  \hline
+  T I/O Sequential & 800 ms & 900 ms
+  \hline
+\end {tabular}
+
+## In our simulation ##
+
+Now coming to our second component RAM and how the data tranfer works in their and what delays take place.
+RAM is a way smaller storage device compared to HDD and is alsos volatile, but that makes it quicker than
+the HDD. The Latency of data movement in RAM is comprised primarily of CAS latency which translates to
+Column Address strobe or signal latency. Since the RAM is built upon columns or cells of memory,
+CAS latency means how much time it would take the RAM to access the memory cells, read the data and make it
+available as an output.
+
 Explain any terms as needed (only if not already).
 
-\section{Discrepancies}
+\section{Discrepencies}
 \label{sec:disc}
 
-Explain any discrepancies found here.
+Explain any discrepencies found here.
 
@@
 \bibliography{references}
+www.ithollow.com/2013/11/18/disk-latency-concepts/
+pages.sc.wisc.edu/~remzi/OSFEP/file-disks.pdf
+www.tomshardware.com/reviews/cas-latency-ram-cl-timings-glossary-definition,6011.html
 

### 2020-11-21 finalPaper/paper.tex

--- 2020-11-21 finalPaper/paper.tex
@@
 
+Latency Time in the all 3 major components being implemented in the simulator vary accordingly to their sizes,
+that becomes the reason why cache is the fastest between RAM and the HDD.
+
+Starting with the hardrive latency time for the input and output of data as it is the largest size of all memories here.
+The latency of a hard drive is a combination of it RPM, Seek Time and Transfer time. In the simulator that we
+implemented, we didnt go into details to a point that we offer the RPM, Seek time and the Transfer time as
+a variable to the user. However we implemented the HDD on data blocks instead of platters, which acted as
+partitions.
+According to Eric Shanks from ithollow.com, the following were the Hard Drive latencies in 2013, given the different
+RPM.
+
+\begin{tabular}{|c|c|}
+  \hline
+  RPMs & Rotational Latency(ms)\\
+  \hline
+  5400 & 11 ms\\
+  \hline
+  7200 & 8 ms\\
+  \hline
+  10000 & 6 ms\\
+  \hline
+  15000 & 4 ms\\
+  \hline
+\end {tabular}
+
+Seek time; which is the time it takes for the Read-Write to move up and down the platter.
+The Seek time itself is comprised of a series of phase times. The first being the acceleration phase,
+the read write arm starts to move, then comes the coasting phase when the arm is moving at full speed. This
+is followed by the deceleration phase where the read write arm slows down. Then finally comes the
+settling phase as the read write head stops right at the data it needs.
+The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
+(source: pages.cs.wisc.edu - Operating Systems: Three Easy peices peices - Remzi Arpaci, Andrea Arpaci).
+
+The last part of total I/O latency for hard drive is the transfer time, which is the time it takes the
+data to be transfer from or to the Read-Write head. Compared to Rotational Latency and seek time, transfer
+time is insignificant, given that the mechanical movements take more time than data movement which is more of
+movement of energy packets.
+
+After all this we come to an equation:
+
+I/O Latency Time of HDD = Rotational Latency + Seek Time + Latency Time
+
+In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the Operating Systems: Three
+easy peices, two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
+data being dealt with was 100 MB. However, another controller in the comparison was whether the data read or writing
+was sequential or random. Sequential meant that large chunks of data are just being read and written which are in
+sequence, whereas the Random transfer just meant Random data was selected from a read issue some where on the disk,
+which is how the data for applications work. Mind it that the data we represented in the simulator was also sequential
+and not random, as the hardrive implemented was built of data blocks.
+
+The I/O Latency time along with other statistics mentioned above was (taken from the source):
+
+\begin{tabular}{|l|r|r|}
+  \hline
+  HDD & Cheetah 15K.5 & Barracuda\\
+  \hline
+  Capacity & 300 GB & 1 TB\\
+  \hline
+  RPM & 15,000 & 7,200\\
+  \hline
+  Average Seek & 4 ms & 9ms\\
+  \hline
+  Platters & 4 & 4\\
+  \hline
+  T I/O Random & 6 ms & 13.2 ms\\
+  \hline
+  T I/O Sequential & 800 ms & 900 ms
+  \hline
+\end {tabular}
+
+## In our simulation ##
+
+Now coming to our second component RAM and how the data tranfer works in their and what delays take place.
+RAM is a way smaller storage device compared to HDD and is alsos volatile, but that makes it quicker than
+the HDD. The Latency of data movement in RAM is comprised primarily of CAS latency which translates to
+Column Address strobe or signal latency. Since the RAM is built upon columns or cells of memory,
+CAS latency means how much time it would take the RAM to access the memory cells, read the data and make it
+available as an output.
+
 Explain any terms as needed (only if not already).

### 2020-11-21 finalPaper/paper.tex
Some proofreading and updated refs

@@
 To download the simulator code, go to \url{osfinal.dachenhausen.org}
-or
-\url{https://github.com/adamdachenhausen/cs330FinalProject} 
-and clone the repository. See \url{https://docs.github.com/en/free-pro-team@latest/github/creating-cloning-and-archiving-repositories/cloning-a-repository}
-for more.
+or \url{https://github.com/adamdachenhausen/cs330FinalProject} and clone the repository.
+See \url{https:docs.github.com/en/free-pro-team@latest/github/
+creating-cloning-and-archiving-repositories/cloning-a-repository} for more.
 
@@
   Either way that the simulator was compiled, the command
+{\singlespace
 \begin{verbatim}
@@
 \end{verbatim}
+}
   will start the simulator
@@
 \end{verbatim}
+}
 If your terminal supports ANSI colored text, then the question should be blue,
@@
 If your terminal does not support the colored text, then you will see something like
-{\singlespace
+
 \begin{verbatim}
@@
 32
-How long would you like it to take in seconds, to read/write to the cache?
+How long would you like it to take in seconds,
+ to read/write to the cache?
 0
-How long would you like it to take in seconds, to read/write to the RAM?
+How long would you like it to take in seconds,
+ to read/write to the RAM?
 5
-Finally, how long would you like it to take in seconds, to read/write to a HDD?
+Finally, how long would you like it to 
+take in seconds, to read/write to a HDD?
 10
-
+\end{verbatim}
+\begin{verbatim}
 ***********System Config**********
@@
   \hline
-  T I/O Sequential & 800 ms & 900 ms
+  T I/O Sequential & 800 ms & 900 ms\\
   \hline
@@
 
-## In our simulation ##
+\subsection*{In our simulation}
 

### 2020-11-21 finalPaper/paper.tex
Updated paper.tex

--- 2020-11-21 finalPaper/paper.tex
@@
 
-Insert all the data gathered by the simulator and from sources here.
-
 Latency Time in the all 3 major components being implemented in the simulator vary accordingly to their sizes,
@@
 partitions.
-According to Eric Shanks from ithollow.com, the following were the Hard Drive latencies in 2013, given the different
-RPM.
+
+Rotatational latency of a hard drive is how long does it take a platter to spin, as after all it is a mechanical
+component bounded by a certain speed.
+According to Eric Shanks from ithollow.com, the following were the Hard Drive Rotational latencies in 2013.
 
@@
 
-In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the Operating Systems: Three
-easy peices, two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
+In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the 'Operating Systems: Three
+easy peices', two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
 data being dealt with was 100 MB. However, another controller in the comparison was whether the data read or writing
@@
 
-\begin{tabular}{|l|r|r|}
+\begin{tabular}{|l|rr|}
   \hline
@@
 the HDD. The Latency of data movement in RAM is comprised primarily of CAS latency which translates to
-Column Address strobe or signal latency. Since the RAM is built upon columns or cells of memory,
+Column Address strobe latency. Since the RAM is built upon columns or cells of memory,
 CAS latency means how much time it would take the RAM to access the memory cells, read the data and make it
-available as an output.
+available as an output. And again, in the simulation we were not able to completely configure the design of
+the RAM to this extent however we did implement a RAM on data blocks.
 
-Explain any terms as needed (only if not already).
+Over the course of Time the latency of RAM has improved, but compared to the bandwidth and capacity it has
+been on the very low side. According to a research done by Kevin Chang from CMU, in his paper 'Understanding
+and Improving the Latency of D-RAM based Memory Systems', in the period of 1999 - 2017, the capacity of
+RAM has increased a gigantic 128 times, where as the Bandwidth of a RAM has increased a good enough 20 times,
+however the improvenement in latency haa only a factor of 1.3x. We can safely assume that the latency of a RAM
+has been consistent compared to the latency of Hard Drive over time or the CPU.
+
+## ##
+
+Compared to the RAM the gap between the CPU latency and the RAM latency has been growing at a high rate. As
+discussed previously the latency of a RAM hasn't had much improvement in the last 3 decades. Sangyeun Cho
+from UPittsburg, comes with the data that considering 2005, the improvement in RAM latency has been less than
+10 times than it was in 1980. However on the other hand the CPU latency has steered up to a performance of
+a whopping 10000 times better in 2005 than it was in 1980.
+
+When we talk about the CPU, we are dealing with the smallest form of major memory in the CPU yet the fastest one.
+As mentioned in the previous section, the CPU consists of caches to save the time to fetch frequent data
+again from the memory. Given that the modern CPUs have 2 levels of cache built in them, we were only able to
+implement a single level on our simulator. The CPU latency is comprised of many factors or sometimes not. That
+depend upon how effecient is cache working. If the CPU cache does not have the data required than the CPU has
+to use the buses to fetch the data from the memory, compromising the bus bandwidth and timing in the process.
+These two different situations can yield a very different latency period.
+
+As of 2012 these were some of the Latencies, that includes different operation on a low level
+and at the high level.(Source: gist/github.com/jboner/2841832)
+
+\begin{tabular}{|l|c|}
+  \hline
+  L1 cache reference & 0.5 ns\\
+  \hline
+  Branch mispredict & 5 ns\\
+  \hline
+  L2 cache reference & 7 ns\\
+  \hline
+  Mutex lock/unclock & 25 ns\\
+  \hline
+  Main memory reference & 100 ns\\
+  \hline
+  Compress 1k bytes with zippy & 3000 ns / 3 us\\
+  \hline
+  Send 1 KB over Gbps network & 10000 ns / 10 us\\
+  \hline
+  Read 4 KB randomly from memory & 150000 ns / 150 us\\
+  \hline
+  Read 1 MB sequentially from memory & 250000 ns / 250 us\\
+  \hline
+  Round trip within same data center & 500000 ns / 500 us\\
+  \hline
+  Read 1 MB sequentially from SSD & 1000 us / 1 ms\\
+  \hline
+  Disk Seek & 10000 us / 10 ms\\
+  \hline
+  Read 1 MB sequentially from the disk & 20000 us / 20 ms\\ 
+  \hline    
+\end {tabular}
+
+Not all of this information, especially concerning network is that closely relevant to our field of study, but
+we can certianly see the latencies that included reading, writing and referencing from the CPU caches, memory
+and the drives had a huge difference in between them. The low level operations that took place within the
+CPU were the quickest of all, referencing Caches and branching. Secondly referencing the RAM (7 ns) was slower than
+referencing the the caches (0.5 ns & 7 ns but alot faster that going to a and reading from disk which was almost a
+20 million times slower than refernecing the L1 Cache and upto a hundred thousand times slower than referencing
+the RAM.
 
@@
 
-Insert conclusions here.
-
 \singlespace
@@
 www.tomshardware.com/reviews/cas-latency-ram-cl-timings-glossary-definition,6011.html
+research.ece.cmu.edu/safari/thesis/kchang_dissertation.pdf
+people.cs.pitt.edu/~cho/cs1541/current/handouts/lect-memh_4up.pdf
 

### 2020-11-21 finalPaper/paper.tex
Updated final paper and ref

@@
 component bounded by a certain speed.
-According to Eric Shanks from ithollow.com, the following were the Hard Drive Rotational latencies in 2013.
+According to \cite(1) Eric Shanks from ithollow.com, the following were the Hard Drive Rotational latencies in 2013.
 
@@
 (source: pages.cs.wisc.edu - Operating Systems: Three Easy peices peices - Remzi Arpaci, Andrea Arpaci).
-
+(source: \cite }
 The last part of total I/O latency for hard drive is the transfer time, which is the time it takes the
@@
 As of 2012 these were some of the Latencies, that includes different operation on a low level
-and at the high level.(Source: gist/github.com/jboner/2841832)
+and at the high level.(Source: }
 
@@
 
+The goal of the project was to find out about the varying latencies of some of the major components of the CPU, with
+or without the simulator. By implementing a simulator good enough to differentiate and manage data we were able to
+find out more about the
+
 \singlespace
@@
 \bibliography{references}
-www.ithollow.com/2013/11/18/disk-latency-concepts/
-pages.sc.wisc.edu/~remzi/OSFEP/file-disks.pdf
-www.tomshardware.com/reviews/cas-latency-ram-cl-timings-glossary-definition,6011.html
-research.ece.cmu.edu/safari/thesis/kchang_dissertation.pdf
-people.cs.pitt.edu/~cho/cs1541/current/handouts/lect-memh_4up.pdf
 

### 2020-11-21 finalPaper/paper.tex
Updated

@@
 The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
-(source: pages.cs.wisc.edu - Operating Systems: Three Easy peices peices - Remzi Arpaci, Andrea Arpaci).
-(source: \cite }
+(source: \cite{tbree} )
+
 The last part of total I/O latency for hard drive is the transfer time, which is the time it takes the
@@
 
-In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the 'Operating Systems: Three
-easy peices', two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
-data being dealt with was 100 MB. However, another controller in the comparison was whether the data read or writing
+In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the \cite{tbree},
+two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
+data being dealt with was 100 MB. However, another controller in the comparison was whether the data reading or writing
 was sequential or random. Sequential meant that large chunks of data are just being read and written which are in
@@
 
-The I/O Latency time along with other statistics mentioned above was (taken from the source):
+The I/O Latency time along with other statistics mentioned above was (taken from the source \cite{tbree}):
 
@@
 been on the very low side. According to a research done by Kevin Chang from CMU, in his paper 'Understanding
-and Improving the Latency of D-RAM based Memory Systems', in the period of 1999 - 2017, the capacity of
-RAM has increased a gigantic 128 times, where as the Bandwidth of a RAM has increased a good enough 20 times,
-however the improvenement in latency haa only a factor of 1.3x. We can safely assume that the latency of a RAM
-has been consistent compared to the latency of Hard Drive over time or the CPU.
-
-## ##
+and Improving the Latency of D-RAM based Memory Systems'(\cite {DRAMTimings}) in the period of 1999 - 2017,
+the capacity of RAM has increased a gigantic 128 times, where as the Bandwidth of a RAM has increased a
+good enough 20 times, however the improvenement in latency had  only a factor of 1.3x. We can safely assume
+that the latency of a RAM has been consistent compared to the latency of Hard Drive over time or the CPU.
 
@@
 As of 2012 these were some of the Latencies, that includes different operation on a low level
-and at the high level.(Source: }
+and at the high level.(Source: \cite {Latencies})
 
@@
 
-Explain any discrepancies found here.
+The goal of the project was to find out about the varying latencies of some of the major components of the CPU, with
+or without the simulator. By implementing a simulator good enough to differentiate and manage data we were able to
+find out more about the memory management in an environment ran by threads. However, since there exists some complex
+and deep details and components in the real world units that we represented, we were only able to implement as much
+as the basic level of each component.
+The real world data that we represented made up most of our information in the paper, but that was along with the
+fact that the data we imported, was a combination of several different projects and researchs.
+Starting from the hard drive we did not represent the sectors, platters and other mechanical components, instead
+we implemented the HDD using data blocks, thus we were only able to gather stats for a sequential read and write.
+A similar implementation of RAM meant that we did not explicitly render the memory cells in the RAM but only the
+data blocks which act as those cells.
+For the CPU we only represented a single level of cache a L1, while almost all the modern CPUS work on dual level
+of cache.
+Being programmed in java our simulator was bounded by the high level barrier, so to represent the low
+level of memory management, we did have to implement delays ourselves. Our delays were then rationalized according
+to the time it takes in the low level. 
+A lot that was not gathered through the simulator, a deep analysis of sources did cover for that. Especially getting
+to know the latencies of new combinations of components that we did not implement in out simulator helped us to a
+more realistic approach. The research itself was based upon a latency simulator with a concentration on what makes
+up the delays in a real world programming system.
 
 \section{Conclusions}
+
 \label{sec:conclusions}
 
-The goal of the project was to find out about the varying latencies of some of the major components of the CPU, with
-or without the simulator. By implementing a simulator good enough to differentiate and manage data we were able to
-find out more about the
+Memory Management within a memory system is a complex set of organizations of several different peices, which
+themselves are a set of several different elements. So if we are to study the time delays as a superset of
+memory latencies, we cannot just discuss the general latencies of each component seperately. But one must
+realize that the components are very closely linked with each other. An example of this would be that if a
+a CPU references either  cache and it hits, that would be an ideal situation for the a real model, but if does
+not find what it needs then it has to ref the memory or eventually the Disks, resulting in so many differnt
+factors for delays, including but not only, the CAS latency of the RAM, the HDD I/O latency, which further
+includes the delays of rotation, read write head, settling time etc.
 

### 2020-11-21 finalPaper/paper.tex
Fixed some errors

@@
 The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
-(source: \cite{tbree} )
+(source: \cite{three} )
 
@@
 
-In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the \cite{tbree},
+In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the \cite{three},
 two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
@@
 
-The I/O Latency time along with other statistics mentioned above was (taken from the source \cite{tbree}):
+The I/O Latency time along with other statistics mentioned above was (taken from the source \cite{three}):
 
@@
 CPU were the quickest of all, referencing Caches and branching. Secondly referencing the RAM (7 ns) was slower than
-referencing the the caches (0.5 ns & 7 ns but alot faster that going to a and reading from disk which was almost a
+referencing the the caches (0.5 ns and 7 ns) but alot faster that going to a and reading from disk which was almost a
 20 million times slower than refernecing the L1 Cache and upto a hundred thousand times slower than referencing

### 2020-11-21 finalPaper/paper.tex
Updated final paper and ref

@@
 The simulator is written in Java, and therefore, you will need to have
-Java installed to run it. See \url{https://www.java.com/en/download/}
+Java installed to run it. See \href{https://www.java.com/en/download/}
  for more.
@@
 \subsection{Acquiring Source Code}
-To download the simulator code, go to \url{osfinal.dachenhausen.org}
-or \url{https://github.com/adamdachenhausen/cs330FinalProject} and clone the repository.
-See \url{https:docs.github.com/en/free-pro-team@latest/github/
-creating-cloning-and-archiving-repositories/cloning-a-repository} for more.
+To download the simulator code, go to \href{osfinal.dachenhausen.org}
+or \href{https://github.com/adamdachenhausen/cs330FinalProject} and clone the repository.
+See
+\href{https:docs.github.com/en/free-pro-team@latest/github/creating-cloning-and-archiving-repositories/cloning-a-repository}
+for more.
 

### 2020-11-21 finalPaper/paper.tex
Updated final paper and ref
fixed citations

--- 2020-11-21 finalPaper/paper.tex
@@
 or \url{https://github.com/adamdachenhausen/cs330FinalProject} and clone the repository.
-See \url{https:docs.github.com/en/free-pro-team@latest/github/
-creating-cloning-and-archiving-repositories/cloning-a-repository} for more.
+See
+\url{https:docs.github.com/en/free-pro-team@latest/github/creating-cloning-and-archiving-repositories/cloning-a-repository}
+for more.
 
@@
 component bounded by a certain speed.
-According to \cite(1) Eric Shanks from ithollow.com, the following were the Hard Drive Rotational latencies in 2013.
+According to \cite{diskLatency}
 

### 2020-11-21 finalPaper/paper.tex
Fixed some errors

@@
 According to \cite{diskLatency} 
-
+\\
 \begin{tabular}{|c|c|}
@@
 \end {tabular}
-
+\\
 Seek time; which is the time it takes for the Read-Write to move up and down the platter.
@@
 
+\\
 \begin{tabular}{|l|rr|}
@@
 \end {tabular}
-
+\\
 \subsection*{In our simulation}
@@
 and at the high level.(Source: \cite {Latencies})
-
+\\
 \begin{tabular}{|l|c|}
@@
 \end {tabular}
-
+\\
 Not all of this information, especially concerning network is that closely relevant to our field of study, but

### 2020-11-21 finalPaper/paper.tex
Update paper.tex

--- 2020-11-21 finalPaper/paper.tex
@@
 referencing the the caches (0.5 ns and 7 ns) but alot faster that going to a and reading from disk which was almost a
-20 million times slower than refernecing the L1 Cache and upto a hundred thousand times slower than referencing
+20 million times slower than referencing the L1 Cache and upto a hundred thousand times slower than referencing
 the RAM.
@@
 The real world data that we represented made up most of our information in the paper, but that was along with the
-fact that the data we imported, was a combination of several different projects and researchs.
+fact that the data we imported, was a combination of research and several different projects.
 Starting from the hard drive we did not represent the sectors, platters and other mechanical components, instead
@@
 data blocks which act as those cells.
-For the CPU we only represented a single level of cache a L1, while almost all the modern CPUS work on dual level
+For the CPU we only represented a single level of cache a L1, while almost all the modern CPUs work on dual level
 of cache.
@@
 
-Memory Management within a memory system is a complex set of organizations of several different peices, which
+Memory Management within a memory system is a complex set of organizations of several different pieces, which
 themselves are a set of several different elements. So if we are to study the time delays as a superset of
-memory latencies, we cannot just discuss the general latencies of each component seperately. But one must
+memory latencies, we cannot just discuss the general latencies of each component separately. But one must
 realize that the components are very closely linked with each other. An example of this would be that if a
 a CPU references either  cache and it hits, that would be an ideal situation for the a real model, but if does
-not find what it needs then it has to ref the memory or eventually the Disks, resulting in so many differnt
+not find what it needs then it has to ref the memory or eventually the Disks, resulting in so many different
 factors for delays, including but not only, the CAS latency of the RAM, the HDD I/O latency, which further

### 2020-11-21 finalPaper/paper.tex
Fixed spelling mistakes

@@
 
-Starting with the hardrive latency time for the input and output of data as it is the largest size of all memories here.
+Starting with the hard drive latency time for the input and output of data as it is the largest size of all memories here.
 The latency of a hard drive is a combination of it RPM, Seek Time and Transfer time. In the simulator that we
-implemented, we didnt go into details to a point that we offer the RPM, Seek time and the Transfer time as
+implemented, we did not go into details to a point that we offer the RPM, Seek time and the Transfer time as
 a variable to the user. However we implemented the HDD on data blocks instead of platters, which acted as
@@
 
-Rotatational latency of a hard drive is how long does it take a platter to spin, as after all it is a mechanical
+Rotational latency of a hard drive is how long does it take a platter to spin, as after all it is a mechanical
 component bounded by a certain speed.
@@
 settling phase as the read write head stops right at the data it needs.
-The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
+The average seek time of hard drives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
 (source: \cite{three} )
@@
 which is how the data for applications work. Mind it that the data we represented in the simulator was also sequential
-and not random, as the hardrive implemented was built of data blocks.
+and not random, as the hard drive implemented was built of data blocks.
 
@@
 
-Now coming to our second component RAM and how the data tranfer works in their and what delays take place.
-RAM is a way smaller storage device compared to HDD and is alsos volatile, but that makes it quicker than
+Now coming to our second component RAM and how the data transfer works in their and what delays take place.
+RAM is a way smaller storage device compared to HDD and is also volatile, but that makes it quicker than
 the HDD. The Latency of data movement in RAM is comprised primarily of CAS latency which translates to
@@
 the capacity of RAM has increased a gigantic 128 times, where as the Bandwidth of a RAM has increased a
-good enough 20 times, however the improvenement in latency had  only a factor of 1.3x. We can safely assume
+good enough 20 times, however the improvement in latency had  only a factor of 1.3x. We can safely assume
 that the latency of a RAM has been consistent compared to the latency of Hard Drive over time or the CPU.
@@
 implement a single level on our simulator. The CPU latency is comprised of many factors or sometimes not. That
-depend upon how effecient is cache working. If the CPU cache does not have the data required than the CPU has
+depend upon how efficient is cache working. If the CPU cache does not have the data required than the CPU has
 to use the buses to fetch the data from the memory, compromising the bus bandwidth and timing in the process.
@@
   \hline
-  Mutex lock/unclock & 25 ns\\
+  Mutex lock/unlock & 25 ns\\
   \hline
@@
 Not all of this information, especially concerning network is that closely relevant to our field of study, but
-we can certianly see the latencies that included reading, writing and referencing from the CPU caches, memory
+we can certainly see the latencies that included reading, writing and referencing from the CPU caches, memory
 and the drives had a huge difference in between them. The low level operations that took place within the
@@
 referencing the the caches (0.5 ns and 7 ns) but alot faster that going to a and reading from disk which was almost a
-20 million times slower than referencing the L1 Cache and upto a hundred thousand times slower than referencing
+20 million times slower than referencing the L1 Cache and up to a hundred thousand times slower than referencing
 the RAM.
@@
 The real world data that we represented made up most of our information in the paper, but that was along with the
-fact that the data we imported, was a combination of research and several different projects.
+fact that the data we imported, was a combination of several different projects and researches.
 Starting from the hard drive we did not represent the sectors, platters and other mechanical components, instead
@@
 data blocks which act as those cells.
-For the CPU we only represented a single level of cache a L1, while almost all the modern CPUs work on dual level
+For the CPU we only represented a single level of cache a L1, while almost all the modern CPUS work on dual level
 of cache.

### 2020-11-21 finalPaper/paper.tex

@@
 The real world data that we represented made up most of our information in the paper, but that was along with the
-fact that the data we imported, was a combination of several different projects and researches.
+fact that the data we imported, was a combination of research and several different projects.
 Starting from the hard drive we did not represent the sectors, platters and other mechanical components, instead

### 2020-11-21 finalPaper/paper.tex
Fixed some errors

--- 2020-11-21 finalPaper/paper.tex
@@
 According to \cite{diskLatency} 
-
+\\
+\\
 \begin{tabular}{|c|c|}
@@
 \end {tabular}
-
+\\
+\\
 Seek time; which is the time it takes for the Read-Write to move up and down the platter.
@@
 The I/O Latency time along with other statistics mentioned above was (taken from the source \cite{three}): 
-
+\\
+\\
 \begin{tabular}{|l|rr|}
@@
 and at the high level.(Source: \cite {Latencies})
-
+\\
+\\
 \begin{tabular}{|l|c|}
@@
 \end {tabular}
-
+\\
+\\
 Not all of this information, especially concerning network is that closely relevant to our field of study, but
@@
 referencing the the caches (0.5 ns and 7 ns) but alot faster that going to a and reading from disk which was almost a
-20 million times slower than referencing the L1 Cache and upto a hundred thousand times slower than referencing
+20 million times slower than refernecing the L1 Cache and upto a hundred thousand times slower than referencing
 the RAM.
@@
 The real world data that we represented made up most of our information in the paper, but that was along with the
-fact that the data we imported, was a combination of research and several different projects.
+fact that the data we imported, was a combination of several different projects and researchs.
 Starting from the hard drive we did not represent the sectors, platters and other mechanical components, instead
@@
 
-Memory Management within a memory system is a complex set of organizations of several different pieces, which
+Memory Management within a memory system is a complex set of organizations of several different peices, which
 themselves are a set of several different elements. So if we are to study the time delays as a superset of
-memory latencies, we cannot just discuss the general latencies of each component separately. But one must
+memory latencies, we cannot just discuss the general latencies of each component seperately. But one must
 realize that the components are very closely linked with each other. An example of this would be that if a
 a CPU references either  cache and it hits, that would be an ideal situation for the a real model, but if does
-not find what it needs then it has to ref the memory or eventually the Disks, resulting in so many different
+not find what it needs then it has to ref the memory or eventually the Disks, resulting in so many differnt
 factors for delays, including but not only, the CAS latency of the RAM, the HDD I/O latency, which further

### 2020-11-21 finalPaper/paper.tex
Fixed some errors

--- 2020-11-21 finalPaper/paper.tex
@@
 According to \cite{diskLatency} 
-
+\\\\
 \begin{tabular}{|c|c|}
@@
 \end {tabular}
-
+\\\\
 Seek time; which is the time it takes for the Read-Write to move up and down the platter.
@@
 The I/O Latency time along with other statistics mentioned above was (taken from the source \cite{three}): 
-
+\\\\
 \begin{tabular}{|l|rr|}
@@
 and at the high level.(Source: \cite {Latencies})
-
+\\\\
 \begin{tabular}{|l|c|}
@@
 \end {tabular}
-
+\\\\
 Not all of this information, especially concerning network is that closely relevant to our field of study, but

### 2020-11-21 finalPaper/paper.tex
Proofreading fixes

--- 2020-11-21 finalPaper/paper.tex
@@
 processing unit with cache memory. This project aims first to create a modular
-system, that could be used to simulate data movement computer components, then
+system, that could be used to simulate data movement between computer components, then
 second to use that simulator and upgrade it to allow data to be collected about
@@
 
-In Section~\ref{sec:memlate} we discuss abstractly how the memory latency 
+In section~\ref{sec:memlate} we discuss abstractly how the memory latency 
 simulator works, and why we built it. Section~\ref{sec:build} describes
@@
 real world data. Section~\ref{sec:disc} goes on to explain any discrepancies 
-found between our data and the real world collected data. Finally, in~\ref{sec:conclusions}
+found between our data and the real world collected data. Finally, in section~\ref{sec:conclusions}
 we summarize our results, as well as report what
@@
 
-Each component's (CPU, RAM, HDD), storage is represented with a simple data structure called
+Each component's (CPU, RAM, HDD), storage is represented with a simple array of a data structure called
 a dataBlock, which represents 1024 bytes. The user then has options to handle data on any
@@
 simulator parameters. Each of these is crucial, and cannot be left blank.
-
-The first thing the simulator will ask you is
+The first thing the simulator will ask you is:
 {\singlespace
@@
 
+In the CPU cache, deleting data move the last added pointer back, so you can overwrite the data
+deleted.
+
 \subsection*{Moving Data}
@@
 Memory management as the name suggests is the act of managing memory in the low level of
-a computer. Often represented by allocation of portions and freeing them once they are
+a computer system. Often represented by allocation of chunks of data and freeing them once they are
 done being used.
@@
 The latency is the delay or the time that it takes the data to be read from memory.
-Since the movement of this data is bound to be within the speed of light,
+Since the movement of this data is bound to be within the speed of light, and furthermore
 the medium it is traveling in, this means that even though that speed is above human physical
@@
 
-We gathered the latency time, which is the time delay taking place in data transfer,reading and
+We gathered the latency time, which is the time delay taking place in data transfer, reading and
 writing at the lowest level of memory management. This was the primary statistic for us to consider
@@
 To represent this level of memory management the simulator was designed to hold some key components.
-These components are those linked directly with the memory usage,including the
-central processing unit (CPU), the Hard Drive and the RAM sticks.
+These components are those linked directly with the memory usage, including the
+central processing unit (CPU), the Hard Drive Disk (HDD) and the Random Access Memory (RAM) sticks.
 
-The CPU further consists of cache, modern day powerful CPUS consisted of 2 level of caches, L1 and L2,
+The CPU further consists of cache, modern day powerful CPUs consist of 3 level of caches, L1, L2, and L3.
 where L1 is very small compared to L2 in memory. L1 ranging from 2KB to 64KB, whereas the second level
-is in the range of 256 KB to 2 MB. Cache is used for quicker access by the CPU of the data that it
+is in the range of 256 KB to 2 MB, but much slower than L1. L3 can be anywhere from 4MB to 50MB, but much slower than L2.\cite{makeuseof}
+Cache is used for quicker access by the CPU of the data that it
 frequently uses. Cache stores that data and thus decreasing the latency. However, in the simulator we
@@
 Further down the hierarchy of memory management, we have the RAM, which is relatively slower to CPU but bigger
-in size by quite some margin. The RAM, which translates to Random Access Memory is a volatile form of memory
-built by several memory cells. There are two types of RAMs; DRAM and SRAM, even thought we did not distinguish
+in size by quite some margin. The RAM is a volatile form of memory
+built by several memory cells. There are two types of RAMs; DRAM and SRAM, although we did not distinguish
 between them in our simulator. Given that the RAM is slower compared to the CPU(cache) it takes time to read
 and move data off it. In the simulator we kept the size and the number of RAM sticks a variable, so
-the user inputs the value for each. Since the RAM is volatile it loses all the data it has on it, once we
-take the power off it. As soon as the power is turned on, the alot of important section of OS is read from
+the user inputs the value for each. Since the RAM is volatile it loses all the data it has on it, once it
+stops being powered. As soon as the power is turned on, the most important section of the operating system(OS) is read from
 the hard drive and put onto the RAM so some important functions can be accessed quickly instead of reading
-it off the hard drive which is way slower. This OS data along with other make up alot the RAM usage when the
+it off the hard drive which is slower. This OS data along with other make up a lot the RAM usage when the
 computer is running. Running other process lead to more writing and reading off the RAM which leads to different
@@
 
-Finally comes the Hard Drive, which is the biggest form of permanent memory but the slowest one. It is
+Finally, the hard drive, which is the biggest form of permanent memory but the slowest one. It is
 extremely large compared to both RAM and cache. Almost all the software including the OS is stored on
 the HDD. HDDs have evolved in many ways over the course of times, making them quicker, but they have
-remained slower compared to RAM and CPU always. In our simulator we have built an HDD on arrays of bytes,
-just like the RAM and cache.
+still remained slower compared to RAM and CPU cache. In our simulator we have built an HDD on an array of dataBlocks,
+just like the RAM.
 
@@
 us to  collect the major statistics we needed here. The latency which was going to be the
-major finding of the simulation was calculated through a stopwatch implemented
-inside our program. As we kept clocking, the time was a projected ratio of the time that was supposed to
+main focus of the simulation was calculated through a stopwatch implemented
+inside our program. As we kept gathering data, the time was a projected ratio of the time that was supposed to
 take place on the basic level. Thus, the time gathered was reduced to the ratio fully suited for the
@@
 
-The Memory, which was the core component of the simulator and the primary part of each component
-was represented by arrays of byte type. As the simulation went on, these arrays in each of the components
+The memory, which was the core component of the simulator and the primary part of each component
+was represented by arrays of dataBlocks. As the simulation went on, the array in each of the components
 were modified depending upon the event taking place. This modification gave us the data for how the size
@@
 
-As the variables we had in the simulation were the number of Cores in the CPU,
+As the variables we had in the simulation were the number of cores in the CPU,
 the number of hard drives, the size of hard drive in bytes, the number of RAM sticks, the size of
-each of the RAM stick and the size of the CPU Cache, we kept a record of each combination. However,
-even with the CPU having a variable number of cores, for the path of study being conducted, we maintained
+each of the RAM stick and the size of the CPU Cache, and the time it took to read or write one megabyte to each component,
+we kept a record of each combination. However, even with the CPU having a variable number of cores,
+for the path of study being conducted, we maintained
 it to be 1 throughout the study. The simulator was structured to take in a value for each of the variables,

### 2020-11-21 finalPaper/paper.tex
Fixed some errors
Further proofreading fixes

--- 2020-11-21 finalPaper/paper.tex
@@
 
-Latency Time in the all 3 major components being implemented in the simulator vary accordingly to their sizes,
-that becomes the reason why cache is the fastest between RAM and the HDD.
+Latency time in the all 3 major components being implemented in the simulator vary accordingly to their sizes,
+that becomes the reason why cache is the fastest between the three component types.
 
-Starting with the hard drive latency time for the input and output of data as it is the largest size of all memories here.
+First, the hard drive latency time for the input and output of data as it is the largest of both memory size, and latency time.
 The latency of a hard drive is a combination of it RPM, Seek Time and Transfer time. In the simulator that we
-implemented, we did not go into details to a point that we offer the RPM, Seek time and the Transfer time as
+implemented, we did not go into details to a point that we offer the RPM, seek time and the transfer time as
 a variable to the user. However we implemented the HDD on data blocks instead of platters, which acted as
@@
 
-Seek time; which is the time it takes for the Read-Write to move up and down the platter.
-The Seek time itself is comprised of a series of phase times. The first being the acceleration phase,
-the read write arm starts to move, then comes the coasting phase when the arm is moving at full speed. This
-is followed by the deceleration phase where the read write arm slows down. Then finally comes the
+Seek time the time it takes for the read-write head to move up and down the platter.
+The seek time itself is comprised of a series of phase times. The first being the acceleration phase,
+the read-write arm starts to move, then comes the coasting phase when the head is moving at full speed. This
+is followed by the deceleration phase where the read-write head slows down. Then finally comes the
 settling phase as the read write head stops right at the data it needs.
@@
 The last part of total I/O latency for hard drive is the transfer time, which is the time it takes the
-data to be transfer from or to the Read-Write head. Compared to Rotational Latency and seek time, transfer
+data to be transfer from or to the read-write head. Compared to rotational latency and seek time, transfer
 time is insignificant, given that the mechanical movements take more time than data movement which is more of
@@
 After all this we come to an equation:
-
+{\singlespace
+\begin{verbatim}
 I/O Latency Time of HDD = Rotational Latency + Seek Time + Latency Time 
-
-In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau, in the \cite{three},
-two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time. The size of
+\end{verbatim}
+}
+In an extensive calculation done by Remzi Arpaci-Dusseau and Andrea Arpaci-Dusseau,
+two hard drives one Cheetah, and the other being Barracuda was compared for the I/O latency time.\cite{three} The size of
 data being dealt with was 100 MB. However, another controller in the comparison was whether the data reading or writing
-was sequential or random. Sequential meant that large chunks of data are just being read and written which are in
-sequence, whereas the Random transfer just meant Random data was selected from a read issue some where on the disk,
-which is how the data for applications work. Mind it that the data we represented in the simulator was also sequential
+was sequential or random. Sequential means that large chunks of data are just being read and written which are in
+sequence, whereas the random transfer just means random data was selected from a read issue somewhere on the disk,
+which is how the data for applications work. Note that the data we represented in the simulator was also sequential
 and not random, as the hard drive implemented was built of data blocks.
-
-The I/O Latency time along with other statistics mentioned above was (taken from the source \cite{three}):
 
@@
 
-Now coming to our second component RAM and how the data transfer works in their and what delays take place.
-RAM is a way smaller storage device compared to HDD and is also volatile, but that makes it quicker than
+RAM is a much smaller storage device compared to HDD and is volatile, but that makes it quicker than
 the HDD. The Latency of data movement in RAM is comprised primarily of CAS latency which translates to
-Column Address strobe latency. Since the RAM is built upon columns or cells of memory,
+Column Address Strobe Latency. Since the RAM is built upon columns or cells of memory,
 CAS latency means how much time it would take the RAM to access the memory cells, read the data and make it
@@
 
-Over the course of Time the latency of RAM has improved, but compared to the bandwidth and capacity it has
+Over the course of time the latency of RAM has improved, but compared to the bandwidth and capacity it has
 been on the very low side. According to a research done by Kevin Chang from CMU, in his paper 'Understanding
@@
 the capacity of RAM has increased a gigantic 128 times, where as the Bandwidth of a RAM has increased a
-good enough 20 times, however the improvement in latency had  only a factor of 1.3x. We can safely assume
-that the latency of a RAM has been consistent compared to the latency of Hard Drive over time or the CPU.
+measly 20 times, however the improvement in latency had  only a factor of 1.3x. We can safely assume
+that the latency of a RAM has been consistent compared to the latency of HDD or CPU over time.
 
-Compared to the RAM the gap between the CPU latency and the RAM latency has been growing at a high rate. As
-discussed previously the latency of a RAM hasn't had much improvement in the last 3 decades. Sangyeun Cho
-from UPittsburg, comes with the data that considering 2005, the improvement in RAM latency has been less than
-10 times than it was in 1980. However on the other hand the CPU latency has steered up to a performance of
-a whopping 10000 times better in 2005 than it was in 1980.
+Compared to the RAM, the gap between the CPU cache latency and the RAM latency has been growing at a high rate. As
+discussed previously the latency of a RAM has not had much improvement in the last 3 decades. Sangyeun Cho
+from the University of Pittsburg, states the improvements in RAM latency has been less than
+10 times than it was in 1980. However on the other hand the CPU latency has improved to be 10000 times better in 2005 than it was in 1980.
 
-When we talk about the CPU, we are dealing with the smallest form of major memory in the CPU yet the fastest one.
+When we talk about the CPU cache, we are dealing with the smallest form of major memory in the system yet the fastest one.
 As mentioned in the previous section, the CPU consists of caches to save the time to fetch frequent data
-again from the memory. Given that the modern CPUs have 2 levels of cache built in them, we were only able to
-implement a single level on our simulator. The CPU latency is comprised of many factors or sometimes not. That
+again from the memory. Given that the modern CPUs have 3 levels of cache built in them, we were only able to
+implement a single level on our simulator. The CPU latency is comprised of many factors that
 depend upon how efficient is cache working. If the CPU cache does not have the data required than the CPU has
@@
 Not all of this information, especially concerning network is that closely relevant to our field of study, but
-we can certainly see the latencies that included reading, writing and referencing from the CPU caches, memory
-and the drives had a huge difference in between them. The low level operations that took place within the
-CPU were the quickest of all, referencing Caches and branching. Secondly referencing the RAM (7 ns) was slower than
+we can certainly see the latencies that included reading, writing and referencing from the CPU caches, RAM
+and the hard drives had a huge difference in between them. The low level operations that took place within the
+CPU were the quickest of all, referencing caches and branching. Secondly referencing the RAM (7 ns) was slower than
 referencing the the caches (0.5 ns and 7 ns) but alot faster that going to a and reading from disk which was almost a

### 2020-11-21 finalPaper/paper.tex
Fixed some errors

--- 2020-11-21 finalPaper/paper.tex
@@
 \end {tabular}
-
-\subsection*{In our simulation}
-
+\\\\
 Now coming to our second component RAM and how the data transfer works in their and what delays take place.
@@
 
+In the simulation where we computed the actual data, we came up with the following statistics:
+Moving a data block from RAM to HDD takes 0 ms
+Moving a data block from HDD to RAM takes 1 ms
+Moving a data block from CPU cache to the RAM is 1 ms
+Moving a data block from Cache to HDD takes 0 ms
+Reading the HDD takes 10 ms
+Reading from the RAM is 2.7 s
+Reading from the cache is 0 ms
+Writing to the HDD is 1817 ms
+Writing to the RAM is 1328 ms
+Writing to Cache is 0
+
 \section{Discrepancies}
@@
 to the time it takes in the low level.
+Due to the limitations of java, we ran into some issues with our delays that caused the RAM only to be accessed at
+a very longed latency, which was the only major error in our simulator
 A lot that was not gathered through the simulator, a deep analysis of sources did cover for that. Especially getting
@@
 includes the delays of rotation, read write head, settling time etc.
+We found out that after the research that  our simulator which based upon a simple memory system, could not
+keep up with a modern memory system, which included tonnes of elements to care about.
 

### 2020-11-21 memSim.java
Updated an error in one of the user prompts

@@ -338,7 +338,7 @@ class memSim extends Thread{
 	    
 	    //Now we prompt where the user wants to put this data
 	    while(userSubSelect1 == userSubSelect2){
-	    System.out.println("Where would you like to read the data from?");
+	    System.out.println("Where would you like to write the data to?");
 	    System.out.println("1.\tHard Drive");
             System.out.println("2.\tRAM");
             System.out.println("3.\tCPU Cache");
